### 2019-03-21 code_excess/explore_maill.Rmd

---
title: "Cochrane Dataset"
output: html_notebook
---

```{r}
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'thesis/code')
```

# Read data
```{r}
source(file.path(PATH, 'prepare.R'))
```

```{r}
require(tidyverse)
require(meta)
```

###Example Reviews
Anti-D administration in pregnancy for preventing Rhesus alloimmunisation: 2 Trials, 3 Identical outcome measures, but with different subgroups.

```{r}
# arrange(filter(madata, file.nr==10) %>% select(comparison.name, outcome.name, outcome.measure, subgroup.name, study.name), study.name)
```

Comparison of

```{r}
# arrange(filter(madata, file.nr==21) %>% select(comparison.name, outcome.name, outcome.measure, subgroup.name, study.name), study.name)
```

###Frequencies of Trials in the Reviews

Frequencies of trials analyzing the same subject in a review (with the same outcome measure):

```{r}
madata %>% group_by(file.nr, outcome.nr, comparison.nr) %>% count %>% group_by(n) %>% count %>% 
  filter(n < 15) %>% 
  full_join( madata %>% group_by(file.nr, outcome.nr, comparison.nr) %>% count %>% group_by(n) %>% count %>% 
    filter(n > 14) %>% ungroup %>% summarise(n = 15, nn = sum(nn))) %>%  
  ggplot(aes(x = n, y = nn)) + geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() + labs(title = "Trials Comparing the Same Subject per Review") + ylab("Trials")
```

Frequencies of trials analyzing the same subject with the same subgroups in a review (and the same outcome measure):

```{r}
madata %>% group_by(file.nr, outcome.nr, comparison.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>% filter(n < 15) %>% 
  full_join( madata %>% group_by(file.nr, outcome.nr, comparison.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>% 
               filter(n > 14) %>% ungroup %>% summarise(n = 15, nn = sum(nn))) %>%  
  ggplot(aes(x = n, y = nn)) + geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() + labs(title = "Trials Comparing the Same Subject in the Same Subgroups per Review") +  ylab ("Trials")
```

###Frequencies of Different Outcome Subjects per Review

```{r}
# madata %>% group_by(file.nr) %>% summarise(diff.out = length(unique(outcome.name))) %>% count(diff.out) %>%  
#   ggplot(aes(x = diff.out, y = n)) + geom_col(col = "gray15", fill = "dodgerblue") +
#   theme_bw() + labs(title = "Different subjects per Review") + xlab("study subjects per review")
```

###Analysis Study Year Frequencies

```{r}
madata %>% filter(study.year > 1900 & study.year < 2019) %>% ggplot(aes(x = study.year)) + geom_histogram(col = "gray15", fill = "dodgerblue") + theme_bw() + labs(title = "Study years of analyses")
```

###Effect sizes over time

Odds ratios with initial finding > 1:

```{r}
madata %>% filter(outcome.measure == "Odds Ratio") %>% filter(!is.na(study.year)) %>% 
  group_by(file.nr, comparison.nr, subgroup.nr) %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>% 
  mutate(first.effect = effect[which.min(study.year)]) %>% 
  filter(first.effect > 1) %>% 
  mutate(scaled.effect = scale(effect), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .5, alpha = 0.2)  + 
  geom_smooth(method = lm) + theme_bw()  + 
  labs(title = "Scaled and centered OR over time", subtitle = "Initial finding > 1") + ylab("Odds Ratio")  + ylab("Time")
```

Odds ratios with initial finding < 1:

```{r}
madata %>% filter(outcome.measure == "Odds Ratio") %>% filter(!is.na(study.year)) %>% 
  group_by(file.nr, comparison.nr, subgroup.nr) %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>% 
  mutate(first.effect = effect[which.min(study.year)]) %>% 
  filter(first.effect < 1) %>% 
  mutate(scaled.effect = scale(effect), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .35, alpha = 0.1)  + 
  geom_smooth(method = lm) + theme_bw() + 
  labs(title = "Scaled and centered OR over time", subtitle = "Initial finding < 1") + ylab("Odds Ratio")  + ylab("Time")
```

Risk ratios with initial finding > 1:

```{r}
#Scaled RR over time with initial RR > 1
madata %>% filter(outcome.measure == "Risk Ratio") %>% filter(!is.na(study.year)) %>% 
  group_by(file.nr, comparison.nr, subgroup.nr) %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>% 
  mutate(first.effect = effect[which.min(study.year)]) %>% 
  filter(first.effect > 1) %>% 
  mutate(scaled.effect = scale(effect), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .25, alpha = 0.025)  + 
  geom_smooth(method = lm) + theme_bw() + 
  labs(title = "Scaled and centered RR over time", subtitle = "Initial finding > 1") + ylab("Risk Ratio")  + xlab("Time") + xlim(c(-4,4)) + ylim(c(-5,7))
```

Risk ratios with initial finding < 1:

```{r}
#Scaled RR over time with initial RR > 1
madata %>% filter(outcome.measure == "Risk Ratio") %>% filter(!is.na(study.year)) %>% 
  group_by(file.nr, comparison.nr, subgroup.nr) %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>% 
  mutate(first.effect = effect[which.min(study.year)]) %>% 
  filter(first.effect < 1) %>% 
  mutate(scaled.effect = scale(effect), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .2, alpha = 0.025)  + 
  geom_smooth(method = lm) + theme_bw() + 
  labs(title = "Scaled and centered RR over time", subtitle = "Initial finding < 1") + ylab("Risk Ratio")  + ylab("Time") + xlim(c(-5,5)) + ylim(c(-5, 10))
```

Statistical test to see if incorporation of a time - intial RR estimate improves a linear model for RR and time:

```{r}
regression.temp <- madata %>% filter(outcome.measure == "Risk Ratio") %>% filter(!is.na(study.year)) %>% 
  group_by(file.nr, comparison.nr, subgroup.nr) %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>% 
  mutate(first.effect = effect[which.min(study.year)]) %>% 
  mutate(effect.sign = ifelse(first.effect > 1, "pos", "neg"), scaled.effect = scale(effect), scaled.time = scale(study.year))
regression.temp$effect.sign <- as.factor(regression.temp$effect.sign)

m1 <- lm(scaled.effect ~ scaled.time, data = regression.temp)
m2 <- lm(scaled.effect ~ scaled.time + effect.sign, data = regression.temp)
m3 <- lm(scaled.effect ~ scaled.time * effect.sign, data = regression.temp)
anova(m1, m2, m3)
```

###Cochrane Heterogeneity test

Cochrane heterogeneity test for Odds Ratios (with meta::metabin()):

```{r}
madata %>% group_by(file.nr, outcome.nr, comparison.nr, subgroup.nr) %>% 
  filter(outcome.measure == "Odds Ratio") %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>%
  filter(events1 > 0 & events2 > 0) %>% 
  summarise(
    Q = metabin(event.e = events1, n.e = total1, event.c = events2, n.c = total2, studlab = study.name, method = "Inverse", sm = "OR")$Q,
    p.value = metabin(event.e = events1, n.e = total1, event.c = events2, n.c = total2, studlab = study.name, 
                      method = "Inverse", sm =  "OR")$pval.Q) %>% 
  ggplot(aes(x = p.value)) + geom_histogram(col = "gray15", fill = "dodgerblue") +
  theme_bw() + labs(title = "Cochrane Heterogeneity Test P-values of ORs")
```

##Transformation to fishers z-scaled correlation

Mean differences and binary outcomes can be compared by transformation to standardized mean difference to correlation and finally to fishers z-scaled correlation.

```{r}
#Transform Binary outcomes to odds ratios -> standardized mean difference -> correlation -> fisher scala correlation
madata <- madata %>% 
  filter(!is.na(outcome.measure)) %>% 
  mutate(odds.ratio = (events1/(total1 - events1)) / (events2/(total2 - events2))) %>% 
  mutate(std.mean.d = log(odds.ratio) * (sqrt(3)/pi)) %>% 
  mutate(correlation = std.mean.d/sqrt((std.mean.d^2) + ((total1 + total2)^2)/(total2*total1))) %>% 
  mutate(fishersz = 0.5 * log( (1 + correlation)/(1 - correlation) ), variance = 1/(total1 + total2 - 3))

#Transform mean difference to standardized mean difference -> correlation -> fisher scala correlation
madata[madata$outcome.measure == "Mean Difference", ] <- madata %>% filter(outcome.measure == "Mean Difference") %>% 
  mutate(std.mean.d = effect/se) %>% 
  mutate(correlation = std.mean.d/sqrt(std.mean.d^2 + ((total1 + total2)^2)/(total2*total1))) %>% 
  mutate(fishersz = 0.5 * log( (1 + correlation)/(1 - correlation) ), variance = 1/(total1 + total2 - 3))

madata[madata$outcome.measure == "Std. Mean Difference", ] <- madata %>% filter(outcome.measure == "Std. Mean Difference") %>% 
  mutate(std.mean.d = effect) %>% 
  mutate(correlation = std.mean.d/sqrt(std.mean.d^2 + ((total1 + total2)^2)/(total2*total1))) %>% 
  mutate(fishersz = 0.5 * log( (1 + correlation)/(1 - correlation) ), variance = 1/(total1 + total2 - 3))

madata[madata$outcome.measure == "Mean difference", ] <- madata %>% filter(outcome.measure == "Mean difference") %>% 
  mutate(std.mean.d = effect) %>% 
  mutate(correlation = std.mean.d/sqrt(std.mean.d^2 + ((total1 + total2)^2)/(total2*total1))) %>% 
  mutate(fishersz = 0.5 * log( (1 + correlation)/(1 - correlation) ), variance = 1/(total1 + total2 - 3))

madata[madata$outcome.measure == "mean difference", ] <- madata %>% filter(outcome.measure == "mean difference") %>% 
  mutate(std.mean.d = effect) %>% 
  mutate(correlation = std.mean.d/sqrt(std.mean.d^2 + ((total1 + total2)^2)/(total2*total1))) %>% 
  mutate(fishersz = 0.5 * log( (1 + correlation)/(1 - correlation) ), variance = 1/(total1 + total2 - 3))
```

```{r}
madata %>% filter(!is.na(fishersz)) %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>%
  ungroup %>% arrange(desc(n)) %>% mutate(trials = cumsum(n * nn)) %>% 
  filter(n < 50) %>% 
  full_join( madata %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>% 
               filter(n > 49) %>% ungroup %>% summarise(n = 50, nn = sum(nn))) %>%  
  ggplot(aes(x = n, y = trials)) + geom_point(col = "gray15") + 
  theme_bw() + labs(title = "Sum of Trials with Number of Replicate (Trials) >= n") + ylab("Number of trials") + xlab("n, number of replicates")
```

```{r}
madata %>% filter(!is.na(fishersz)) %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>%
  ungroup %>% arrange(desc(n)) %>% mutate(trials = cumsum(nn)) %>% 
  filter(n < 50) %>% 
  full_join( madata %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>% 
               filter(n > 49) %>% ungroup %>% summarise(n = 50, nn = sum(nn))) %>%  
  ggplot(aes(x = n, y = trials)) + geom_point(col = "gray15") + 
  theme_bw() + labs(title = "Total Number of Studies with Number of Replicate (Trials) >= n") + ylab("Number of Studies") + xlab("n, number of replicates")
```