### 2019-04-29 report/chapter02?.Rnw
Presentation upload

% LaTeX file for Chapter 02

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(
    fig.path='figure/ch02_fig',
    self.contained=FALSE,
    cache=TRUE
)
@

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch02_fig',
               echo=TRUE, message=FALSE,
               fig.width=8.1, fig.height=3,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

<<echo = FALSE>>=
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
PATH2 = file.path(PATH_HOME, 'PubBias')
FILE = 'cochrane_2018-06-09.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH2, 'code')
PATH_RESULTS = file.path(PATH2, 'results')
PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

file_results = "pb.RData"

source(file.path(PATH_CODE, 'PubBias_functions.R'))

data = pb.readData(path = PATH_DATA, file = FILE)
tmp = pb.clean(data)
data = tmp[[1]]
aliases = tmp[[2]]

file.bin <- "pb.bin.RData"
if (file.exists(file.path(PATH_RESULTS, file.bin))) {
  load(file.path(PATH_RESULTS, file.bin))
} else { 
  meta.bin <- meta.bin.complete(data, min.study.number = 10, sig.level = 0.05)
  save(meta.bin, file =  file.path(PATH_RESULTS, file.bin))
}

file.cont <- "pb.cont.RData"
if (file.exists(file.path(PATH_RESULTS, file.cont))) {
  load(file.path(PATH_RESULTS, file.cont))
} else { 
  meta.cont <- meta.cont.complete(data, min.study.number = 10, sig.level = 0.05)
  save(meta.cont, file =  file.path(PATH_RESULTS, file.cont))
}

meta <- pb.meta.merge(meta.bin, meta.cont)

file.cont <- "mly.cont.RData"
if (file.exists(file.path(PATH_RESULTS, file.cont))) {
  load(file.path(PATH_RESULTS, file.cont))
} else { 
  data.cont <- mly.cont(data.ext, 0.05, min.study.number = 2)
  save(data.cont, file =  file.path(PATH_RESULTS, file.cont))
}

file.bin <- "mly.bin.RData"
if (file.exists(file.path(PATH_RESULTS, file.bin))) {
  load(file.path(PATH_RESULTS, file.bin))
} else { 
  data.bin <- mly.bin(data.ext, 0.05, min.study.number = 2)
  save(data.bin, file =  file.path(PATH_RESULTS, file.bin))
}

load(file.path(PATH_RESULTS, file = "mly.RData"))
load(file.path(PATH_RESULTS, file = "data.processed.RData"))

require(biostatUZH)
require(tidyverse)
require(meta)
require(metasens)
require(gridExtra)
require(xtable)

###################################################################################################################
###################################################################################################################

#Barbiturate Examples
barbi1 <- arrange(filter(data, file.nr == 21) %>% 
          select(Study = study.name, Comparison = comparison.name, Outcome = outcome.name, 
                 Events = events1, Total = total1, Events_c = events2, Total_c = total2) %>% 
            slice(c(1,3)), Study) 
barbi2 <- arrange(filter(data, file.nr == 21) %>% 
          select(Study = study.name, Comparison = comparison.name,Outcome = outcome.name), Study)
barbi2[barbi2$Study == "P\303\251rez-B\303\241rcena 2008", 1] = "Perez-Barcena 2008"

#Missing values
cont.out <- data %>% filter(outcome.measure.new == "Mean Difference" | outcome.measure.new == "Std. Mean Difference" | 
                                 outcome.measure.new == "Hedges' G")

missing.mean <- which(is.na(cont.out$mean1) | is.na(cont.out$mean2))
missing.effect <- which(is.na(cont.out$effect))
missing.means <- intersect(missing.mean, missing.effect) #Cont. results that have neither complete means nor an effect
wrong.effect <- which(cont.out$mean1 == 0 & cont.out$mean2 == 0 & cont.out$effect == 0) #Cont.results that have only zeros
missing.effects <- length(c(missing.means, wrong.effect))

missing.sd <- which(is.na(cont.out$sd1) | is.na(cont.out$sd2))
missing.se <- which(is.na(cont.out$se))
missing.sds <- intersect(missing.se, missing.sd) #Cont. results that have neither complete means nor an effect
wrong.sds <- which(cont.out$sd1 == 0 & cont.out$sd2 == 0 & cont.out$se == 0) #Cont.results that have only zeros
missing.sds <- length(c(missing.sds, wrong.sds))

missing.ss <- which(is.na(data$total1) | is.na(data$total2))
wrong.ss <- which(data$total1 <= 0 | data$total2 <= 0)
missing.ssize <- length(c(missing.ss, wrong.ss))

missing.year <- sum(is.na(data$study.year)) + length(c(which(data$study.year < 1920), which(data$study.year > 2019)))

missing.table <- rbind("Missing mean values and mean differences" = missing.effects,
                         "Missing standard deviations and standard errors" = missing.sds,
                         "Missing sample sizes" = missing.ssize,
                         "Missing study year" = missing.year)

#Mean and median number of different outcome types of reviews
mean.diff.out <- data %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>%
  ungroup() %>% group_by(file.nr) %>%
  count() %>% ungroup %>% summarise(mean = mean(n))

median.diff.out <- data %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>%
  ungroup() %>% group_by(file.nr) %>%
  count() %>% ungroup %>% summarise(mean = median(n))

#Study.year quantiles:
data.ext <- data.ext %>% mutate(study.year = ifelse(study.year > 1920, study.year, NA))
publication.year.range <- c(quantile(data.ext$study.year, 0.05, na.rm = T), quantile(data.ext$study.year, 0.25, na.rm = T), 
                            quantile(data.ext$study.year, 0.5, na.rm = T), quantile(data.ext$study.year, 0.75, na.rm = T),
                            round(mean(data.ext$study.year, na.rm = T), 2))

#Outcome.measure frequency table:
outcome.measure.frequencies <- rbind(data %>% group_by(outcome.measure.new) %>% count() %>% 
                                       arrange(desc(n)) %>% ungroup() %>% filter(row_number() < 9),
                                     data %>% group_by(outcome.measure.new) %>% count() %>% arrange(desc(n)) %>% 
                                       ungroup() %>% filter(row_number() > 8) %>% 
                                       summarise(outcome.measure.new = "other", n = sum(n)))

outcome.measure.frequencies <- outcome.measure.frequencies %>% mutate(percentage = round(n/sum(n),3)*100)
names(outcome.measure.frequencies) <- c("Outcome measure", "n", "Percentage")
outcome.measure.frequencies$Percentage <- paste(outcome.measure.frequencies$Percentage, "%", sep = "")

#Total and group size quantiles:
data.ext <- data.ext %>% ungroup() %>%  mutate(total.n = total1 + total2)
samplesize.range <- c(quantile(data.ext$total.n, 0.05, na.rm = T), quantile(data.ext$total.n, 0.25, na.rm = T), 
                            quantile(data.ext$total.n, 0.5, na.rm = T), quantile(data.ext$total.n, 0.75, na.rm = T),
                            round(mean(data.ext$total.n, na.rm = T), 2))
treatment.group.size.range <- c(quantile(data.ext$total1, 0.05, na.rm = T), quantile(data.ext$total1, 0.25, na.rm = T), 
                      quantile(data.ext$total1, 0.5, na.rm = T), quantile(data.ext$total1, 0.75, na.rm = T),
                      round(mean(data.ext$total1, na.rm = T), 2))

#Frequencies of results:
comp.freq <- data %>% group_by(file.nr) %>% summarise(n = n())
comp.range <- c(length(which(comp.freq$n < 6)), quantile(comp.freq$n, 0.25, na.rm = T), 
                      quantile(comp.freq$n, 0.5, na.rm = T), quantile(comp.freq$n, 0.75, na.rm = T),
                      round(mean(comp.freq$n, na.rm = T), 2))

#Frequencies of studies:
study.freq <- data %>% group_by(file.nr) %>% distinct(study.name) %>% count()
study.range <- c(length(which(study.freq$n < 3)), quantile(study.freq$n, 0.25, na.rm = T), 
                      quantile(study.freq$n, 0.5, na.rm = T), quantile(study.freq$n, 0.75, na.rm = T),
                      round(mean(study.freq$n, na.rm = T), 2))

#Pooling studies
cum.repr.trials.subg <- data %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>%
  filter(n < 15) %>%
  full_join( data %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
               filter(n > 14) %>% ungroup %>% summarise(n = 15, nn = sum(nn))) %>% 
  ungroup() %>% arrange(desc(n)) %>% mutate(csum  = cumsum(nn)) %>% arrange(n)
# cum.repr.trials.nosub <- data %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
#   filter(n < 15) %>%
#   full_join( data %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
#                filter(n > 14) %>% ungroup %>% summarise(n = 15, nn = sum(nn))) %>% 
#   ungroup() %>% arrange(desc(n)) %>% mutate(csum  = cumsum(nn)) %>% arrange(n)

# cum.repr.trials <- cbind(cum.repr.trials.nosub, cum.repr.trials.subg)[,c(1,3,6)]

colnames(cum.repr.trials.subg)  <- c("n","Number of groups", "Cumulative sum of groups")
@

\chapter{The Cochrane Dataset}

\subsection{Structure and Content}
The dataset consists of \Sexpr{length(unique(data$file.nr))} systematic reviews from the Cochrane Library with \Sexpr{length(unique(data$study.name))} studies and \Sexpr{dim(data)[1]} results. A result compares clinical or medical interventions or treatments. Each study provides (multiple) results of clinical interventions.

In Table \ref{barbiturate.row}, two results from a systematic review about effects of barbiturates are shown as they are given in the dataset. As can be seen, the result is further specified by the variables in the columns. The comparison variable specifies what treatments or interventions are compared, the outcome variable how it is compared, and the subgroup variable (not given in table) indicates if the result belongs to a certain subgroup. Here, the result is of a binary outcome, so the events in the barbiturate treatment group and the total number of participants are given in columns ``Events'' and ``Total'' and the number of events in the control group ``Events\_c'' and participants ``Total\_c''. Events denotes here number of deaths at the end of follow up.

<<echo=FALSE, results = 'asis'>>=
print(xtable(barbi1, label = "barbiturate.row", digits = 0,
             caption = "Example of two results as given in the dataset. Events denotes the count of events in the treatment group while Events c the count of events in the group compared to. Further descriptive variables have been ommitted"), 
      include.rownames = F, size = "scriptsize")
@

A complete listing of the variables of a result is given in Table \ref{variable}. They can roughly be separated into variables that specify the review in which the result is contained and variables that specify the result itself (separated by a horizontal line in Table \ref{variable}).

\begin{table}[ht]
  \begin{center}
    \begin{tabular}{l l}
      \textbf{Variable} & \textbf{Description}\\
      \hline
      \textbf{file.nr} & The number of the file from which the review data has been \\&gathered. This file corresponds to a file available in the. \\& Cochrane library\\
      \textbf{doi} & Digital object identifier. A unique id of the review such that  \\ &the full text of the review can be found on the web.\\
      \textbf{file.index} & Internal index of the file in the Cochrane library.\\
      \textbf{file.version} & Denotes the version of the review, since the reviews are \\ &occasionally updated.\\
      %\multicolumn{2}{c}{textbf{Study level variables}}\\ 
      \hline
      \textbf{study.name} & Name of the study to which the result belongs\\
      \textbf{study.year} & Year in which the study was published\\
      \hline
      \textbf{comparison.name/.nr} & Specification of the interventions compared in the study  \\ &and a unique number for the comparison\\
      \textbf{outcome.name/.nr} & Specification by which outcome the interventions are compared\\ &and a unique number for the outcome\\
      \textbf{subgroup.name/.nr} & Potentially indication of affiliation to subgroups and a \\ &unique number for the subgroup\\
      \textbf{outcome.measure} & Indication of the quantification method of the effect \\ &(of one intervention compared to the other).\\
      \textbf{effect} & Measure of the effect given in the quantity denoted by \\ &``outcome measure''.\\
      \textbf{se} & Standard error of the measure of the effect,\\
      \textbf{events1/events2} & The counts of patients with an outcome \textit{if}\\ &measurement/outcome is binary or dichotomous \\ &2 (1 for treatment group and 2 for control group).\\
      \textbf{total1/total2} & Number of patients in groups.\\
      \textbf{mean1/mean2} & Mean of patient measurements \textit{if} outcome is continuous.\\
      \textbf{sd1/sd2} & Standard deviation of mean \textit{if} \\ &outcome is continuous.
    \end{tabular}
  \caption{Dataset variable names and descriptions  \label{variable}}

  \end{center}
\end{table}

Results are part of studies that are again part of a (systematic) review. This structure of a review is shown in Figure \ref{review}.

\begin{figure}
\tikzstyle{every node}=[draw=black,thick,anchor=west]
\tikzstyle{selected}=[draw=red,fill=red!30]
\tikzstyle{optional}=[dashed,fill=gray!50]
\begin{tikzpicture}
[grow = right, anchor = west, 
  growth parent anchor=east, % added code
  parent anchor=east]
  \node {Review} [edge from parent fork right]
    child { node {Comparison 2}
      child { node {Outcome 2}}
      child { node {Outcome 1}
        child { node {Subgroup 2}}
        child { node {Subgroup 1}
          child  { node {Result 2}}
          child  { node {Result 1}}
          }}
    }
    child [missing] {}		
    child { node {Comparison 1  }};
\end{tikzpicture}
\caption{Structure of a hypothetical review with two different comparisons\label{review.structure}}
\label{review.structure}
\end{figure}

\vspace{0mm}
The structure of a review will now be outlined based on an example of the dataset. Lets consider the previously mentioned barbiturate and head injury review. The aim was to ``assess the effects of barbiturates in reducing mortality, disability and raised ICP (intra-cranial pressure) in people with acute traumatic brain injury'' as well as to ``quantify any side effects resulting from the use of barbiturates''. Since there are arguments for and against use of barbiturates, the authors of the review did a comprehensive literature search and collected all available study findings.

\vspace{0mm}
The review comprises five studies in total. Three of them compared barbiturate to placebo, one compared barbiturate to Mannitol and one Pentobarbital to Thiopental, which would be the comparison to speak in the previously introduced notion. The studies have different outcomes, for example, death or death and severe disability at follow up. One study split up outcomes for patients with and without haematoma, which would be subgroups.

\vspace{0mm}
The complete listing of outcomes is in table \ref{barbiturates}. The table also gives an illustration of the variety of data that can be included in a review. We have for example continuous (mean body temperature) and binary outcome data (death). Additionally to primary and secondary outcomes, often also dropouts and adverse effects are included in a review.

<<echo=FALSE, results='asis'>>=
print(xtable(barbi2, label = "barbiturates", caption = "Barbiturate and head injury review. In the columns, study names, comparison and outcome measure of the results are given"), include.rownames = F, size = "footnotesize")
@

It is important not to confuse results with studies. A study can contribute multiple results to a systematic review, for example, primary and secondary outcomes and adverse effects.

\vspace{0mm}
Information about missing values in the dataset is given in Table \ref{missing}. For variables as research subject, outcome and subgroup name and event counts there are no missing values. The relative amount of missing values is very low except for study years. In the case of continuous outcomes, the cases have been counted were effect sizes and standard errors of effect sizes are not available. This means that neither mean values of groups nor mean differences are given, or that neither standard deviation nor standard error is given for a result. Study years before 1920 and after 2019 have been counted as missing, as well as sample sizes below zero.

<<echo=FALSE, results='asis'>>=
print(xtable(missing.table, label = "missing", caption = "Number of missing variables and measurements in the dataset", align = "lr"), include.colnames = F,
      size = "footnotesize", digits = 0)
@

The studies that are included in the reviews and have been published are most often from the years after 1980 (5\% quantile = \Sexpr{publication.year.range[1]}). The median of the publication years is \Sexpr{publication.year.range[3]}, the mean \Sexpr{publication.year.range[5]} and the quartiles are \Sexpr{publication.year.range[2]} and \Sexpr{publication.year.range[4]}. Only a handful ($n$ = 18) have been published in 2018, none in 2019. No information is available concerning unpublished results and studies.

\vspace{0mm}
The results of a study are summarised in a result by a effect measure. This effect measure varies, depending on whether data is continuous, binary or time-to-event. The most abundant effect measures used are summarised in Table \ref{outcome.measure.frequencies}. One can conclude of the table that roughly 30 \% of outcomes in the dataset are continuous and the being some sort of discrete or binary outcomes, most often binary (> 65\%).

<<echo=FALSE, results='asis'>>=
print(xtable(outcome.measure.frequencies, label = "outcome.measure.frequencies", 
             caption = "Frequencies of outcome measures among results. n denotes the total number 
             of results with the outcome measure and percentage the percentage of the outcome measure,", 
             align = "llll", digits = 1), include.rownames = F, size = "footnotesize")
@

The sample sizes amongst results vary to some extent. There are 5\% of treatment group sample sizes that are smaller tha \Sexpr{treatment.group.size.range[1]}, the 5\% quantile. The first quartile is \Sexpr{treatment.group.size.range[2]}, the median \Sexpr{treatment.group.size.range[3]}, the mean \Sexpr{treatment.group.size.range[5]} and the third quartile \Sexpr{treatment.group.size.range[4]}. The large difference between median and mean is caused by very large groups with over 2,000,000 participants. Analogously, the quantiles of the total sample size are: 5\% quantile = \Sexpr{samplesize.range[1]}, first quartile = \Sexpr{samplesize.range[2]}, median = \Sexpr{samplesize.range[3]} and third quartile = \Sexpr{samplesize.range[4]}. The mean is \Sexpr{samplesize.range[5]}.

\vspace{0mm}
The mean and median number of results per review are \Sexpr{study.range[5]} and \Sexpr{study.range[3]}. There are \Sexpr{comp.range[1]} reviews with five or fewer results, and the quartiles are \Sexpr{comp.range[2]} and \Sexpr{comp.range[4]}. Similarly, the number of reviews with a maximum of two studies included is \Sexpr{study.range[1]}, the mean study number is \Sexpr{study.range[5]}, the median \Sexpr{study.range[3]} and the interquartile range \Sexpr{study.range[2]} and \Sexpr{study.range[4]}. The discrepancy between mean and median is again due to large reviews with a high number of studies and results, most extreme in \citet{largest.review} which is a systematic review about antibiotic prophylaxisfor preventing infection after cesarean section, with 95 studies and 1497 results in total.

\vspace{0mm}
For results to be suitable for usage in meta-analysis, they have to be identical with respect to comparison and outcome. More speicifically, the studies in the dataset that have the same comparison, outcome and subgroup can be pooled in a meta-analysis since their research subject and experimental setup can be considered sufficiently homogeneous. %Another approach that is not made here is to analyze all subgroups together, which is possible if enough studies are given.
\vspace{0mm}
The amount of studies that can be pooled for meta-analyses is of special interest in this masters thesis. Thus, the dataset is divided in groups whith identical experimental setup. The size of the group denotes how many results are included in a group.

\vspace{0mm}
Table \ref{repr.groups} shows the number of \textit{groups} of groups with equal or more than $n$ results. Practically, this means that a given number of meta analyses can be performed with each having at least $n$ results.

<<echo=FALSE, results='asis'>>=

print(xtable(cum.repr.trials.subg, label = "repr.groups", caption = "Cumulative number of groups with number of reproduction trials >= n", align = "llll", digits = 0), include.rownames = F, size = "footnotesize")
@