### 2019-05-13 communication/newplots_discussion.Rmd

---
title: "Plots for Presentation and Report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, echo=FALSE, warning=FALSE}
rm(list = ls())
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
PATH2 = file.path(PATH_HOME, 'PubBias')
FILE = 'cochrane_2018-06-09.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH2, 'code')
PATH_RESULTS = file.path(PATH2, 'results')
PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

file_results = "pb.RData"

source(file.path(PATH_CODE, 'PubBias_functions.R'))

# data = pb.readData(path = PATH_DATA, file = FILE)
# tmp = pb.clean(data)
# data = tmp[[1]]
# aliases = tmp[[2]]

file.dat <- "data.RData"
if (file.exists(file.path(PATH_RESULTS, file.dat))) {
  load(file.path(PATH_RESULTS, file.dat))
} else {
  data = pb.readData(path = PATH_DATA, file = FILE)
  tmp = pb.clean(data)
  data = tmp[[1]]
  aliases = tmp[[2]]
  save(data, file =  file.path(PATH_RESULTS, file.dat))
}

load(file.path(PATH_RESULTS, file = "mly.RData"))
load(file.path(PATH_RESULTS, file = "data.processed.RData"))

file.bin <- "pb.bin.RData"
if (file.exists(file.path(PATH_RESULTS, file.bin))) {
  load(file.path(PATH_RESULTS, file.bin))
} else {
  meta.bin <- meta.bin.complete(data.ext, min.study.number = 10, sig.level = 0.05, sm = "OR")
  save(meta.bin, file =  file.path(PATH_RESULTS, file.bin))
}

file.cont <- "pb.cont.RData"
if (file.exists(file.path(PATH_RESULTS, file.cont))) {
  load(file.path(PATH_RESULTS, file.cont))
} else {
  meta.cont <- meta.cont.complete(data.ext, min.study.number = 10, sig.level = 0.05)
  save(meta.cont, file =  file.path(PATH_RESULTS, file.cont))
}

file.meta <- "meta.RData"
if (file.exists(file.path(PATH_RESULTS, file.meta))) {
  load(file.path(PATH_RESULTS, file.meta))
} else {
  meta <- pb.meta.merge(meta.bin, meta.cont)
  save(meta, file =  file.path(PATH_RESULTS, file.meta))
}

file.cont <- "mly.cont.RData"
if (file.exists(file.path(PATH_RESULTS, file.cont))) {
  load(file.path(PATH_RESULTS, file.cont))
} else {
  data.cont <- mly.cont(data.ext, 0.05, min.study.number = 2)
  save(data.cont, file =  file.path(PATH_RESULTS, file.cont))
}

file.bin <- "mly.bin.RData"
if (file.exists(file.path(PATH_RESULTS, file.bin))) {
  load(file.path(PATH_RESULTS, file.bin))
} else {
  data.bin <- mly.bin(data.ext, 0.05, min.study.number = 2)
  save(data.bin, file =  file.path(PATH_RESULTS, file.bin))
}

require(biostatUZH)
require(tidyverse)
require(meta)
require(metasens)
require(gridExtra)

#Meta filtering: 
metac.bin <- meta.bin %>% filter(n.sig.type.bin > 1) %>% filter((se.max^2)/(se.min^2) > 4) %>% filter(I2 < 0.5)
metac.cont <- meta.cont %>% filter(n.sig.type.cont > 1) %>% filter((se.max^2)/(se.min^2) > 4) %>% filter(I2 < 0.5)
metac <- meta %>% filter(n.sig.type > 1) %>% filter((se.max^2)/(se.min^2) > 4) %>% filter(I2 < 0.5)

#Significant proportion
test.bin <- metac.bin %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                          schwarzer.test = mean(schwarzer.test),
                                                          rucker.test = mean(rucker.test),
                                                          harbord.test = mean(harbord.test),
                                                          peter.test = mean(peter.test))
test.bin <- test.bin %>% gather(key = "test.type", value = "mean")

p.bin <- metac.bin %>% ungroup() %>% 
  select(egger.test, schwarzer.test, rucker.test, harbord.test, peter.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>%  
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) + ggtitle("Binary Outcomes") + theme(legend.position = "top") +
  guides(fill=guide_legend(title=NULL))+
  annotate("text", x = test.bin$test.type, y = 500, 
           label = paste(round(test.bin$mean, 2)*100, "% rejected"), 
           color = "white")

#Test Results: Continuous
test.cont <- metac.cont %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                            begg.test = mean(begg.test),
                                                            thomson.test = mean(thomson.test))

test.cont <- test.cont %>% gather(key = "test.type", value = "mean")

p.cont <- metac.cont %>% ungroup() %>% 
  select(egger.test, thomson.test, begg.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>% 
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) + ggtitle("Continuous Outcomes") + theme(legend.position = "top") +
	guides(fill=guide_legend(title=NULL))+
  annotate("text", x = test.cont$test.type, y = 150, 
           label = paste(round(test.cont$mean, 2)*100, "% rejected"), 
           color = "white")

#P-value distribution:
pval.bin <- metac.bin %>% ungroup() %>% 
  select(pval.egger.bin, pval.schwarzer.bin, pval.rucker.bin, pval.harbord.bin, pval.peter.bin) %>% 
  gather(key = "test.type", value = "p.value") %>% 
  ggplot(aes(x = p.value)) + geom_histogram() + theme_bw() + facet_wrap(~test.type)

pval.cont <- metac.cont %>% ungroup() %>% 
  select(pval.egger.cont, pval.thomson.cont, pval.begg.cont) %>% 
  gather(key = "test.type", value = "p.value") %>% 
  ggplot(aes(x = p.value)) + geom_histogram() + theme_bw() + facet_wrap(~test.type)

#Sum of significant tests
agree.bin <- meta.bin %>% mutate(n.sig = peter.test + rucker.test + egger.test + harbord.test + schwarzer.test) %>% 
  group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
  ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Binary Outcomes")

agree.cont <- meta.cont %>% mutate(n.sig = egger.test + thomson.test + begg.test) %>% 
  group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
  ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Binary Outcomes")
```

#Publication Bias/Small study effects tests

Publication bias tests for all meta-analyses with n<= 10, $I^2$ < 0.5, largest se/smallest se > 4 and at least one significant single study result. Continuous and binary results are shown separatedly. The first plot shows the proportion of tests with p-values < 0.05. Afterwards, a histogramm with the p-values for the tests is shown.

```{r, fig.height = 3, message = FALSE, echo = FALSE}
grid.arrange(p.bin, p.cont, ncol = 2)
```

```{r, fig.height = 3.2, message = FALSE, echo = FALSE}
plot(pval.bin)
plot(pval.cont)
```

The following plot shows how many meta-analyses were significant when tested only with one test, two tests etc..

```{r, fig.height = 2.5, echo = FALSE}
grid.arrange(agree.bin, agree.cont, ncol = 2)
```