### 2019-05-14 Presentation/present.Rnw
presentation

\documentclass[english]{beamer}
%% The most common packages are already included in:
\usetheme{biostat}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage{amsmath,amsfonts,tikz, amssymb}
\usetikzlibrary{trees}

%% Header data: (adjust to your needs:
\def\uzhunit{Biostatistics}             %% if (not) needed comment/uncomment
%\def\uzhunitext{STA480}

\title{Publication Bias in Meta-Analysis}%[Publication Bias]
%% Optional Argument in [Brackets]: Short Title for Footline

%% The following are all optional, simply comment them
%\subtitle{Publication Bias in the Cochrane Libary}
%\institute{Biostatistics Journal Club}  %% optional
\author{Giuachin Kreiliger}
%\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<include = FALSE>>= 
library(knitr)
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
PATH2 = file.path(PATH_HOME, 'PubBias')
FILE = 'cochrane_2018-06-09.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH2, 'code')
PATH_RESULTS = file.path(PATH2, 'results')
PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

file_results = "pb.RData"

source(file.path(PATH_CODE, 'PubBias_functions.R'))

file.dat <- "data.RData"
if (file.exists(file.path(PATH_RESULTS, file.dat))) {
  load(file.path(PATH_RESULTS, file.dat))
} else {
  data = pb.readData(path = PATH_DATA, file = FILE)
	tmp = pb.clean(data)
	data = tmp[[1]]
	aliases = tmp[[2]]
  save(data, file =  file.path(PATH_RESULTS, file.dat))
}

load(file.path(PATH_RESULTS, file = "data.processed.RData"))

require(tidyverse)
require(meta)
require(xtable)

#Meta-analysis example
data.ext <- data.ext %>% mutate(study.year = ifelse(study.year < 2019, study.year, NA)) %>% 
    mutate(study.year = ifelse(study.year > 1920, study.year, NA))
biased.rev <- data.ext %>% filter(meta.id == 122943)
unbiased.rev <- data.ext %>% filter(meta.id == 181863)
meta.example <- metacont(n.e = total1, mean.e = mean1, sd.e = sd1, n.c = total2, mean.c = mean2, sd.c = sd2, 
												 studlab = study.name, data = biased.rev)
meta.nexample <- metacont(n.e = total1, mean.e = mean1, sd.e = sd1, n.c = total2, mean.c = mean2, sd.c = sd2, 
												 studlab = study.name, data = unbiased.rev)

#Barbiturate Examples
barbi1 <- arrange(filter(data, file.nr == 21) %>%
          select(Study = study.name, Comparison = comparison.name, Outcome = outcome.name,
                 Events = events1, Total = total1, Events_c = events2, Total_c = total2) %>%
            slice(c(1,3)), Study)
barbi1 <- barbi1 %>% mutate(Comparison = "Barbiturate vs ..", Outcome = "Death at ..")

barbi2 <- arrange(filter(data, file.nr == 21) %>%
          select(Study = study.name, Comparison = comparison.name,Outcome = outcome.name,
                 Events = events1, Total = total1, Events_c = events2, Total_c = total2), Study) %>% group_by(Comparison) %>% distinct(Outcome, .keep_all = T)
barbi2[barbi2$Study == "P\303\251rez-B\303\241rcena 2008", 1] = "Perez-Barcena 2008"

barbi2$Comparison <- strtrim(barbi2$Comparison, 19)
barbi2$Outcome <- strtrim(barbi2$Outcome, 19)

#Missing values
cont.out <- data %>% filter(outcome.measure.new == "Mean Difference" | outcome.measure.new == "Std. Mean Difference" |
                                 outcome.measure.new == "Hedges' G")

missing.mean <- which(is.na(cont.out$mean1) | is.na(cont.out$mean2))
missing.effect <- which(is.na(cont.out$effect))
missing.means <- intersect(missing.mean, missing.effect) #Cont. results that have neither complete means nor an effect
wrong.effect <- which(cont.out$mean1 == 0 & cont.out$mean2 == 0 & cont.out$effect == 0) #Cont.results that have only zeros
missing.effects <- length(c(missing.means, wrong.effect))

missing.sd <- which(is.na(cont.out$sd1) | is.na(cont.out$sd2))
missing.se <- which(is.na(cont.out$se))
missing.sds <- intersect(missing.se, missing.sd) #Cont. results that have neither complete means nor an effect
wrong.sds <- which(cont.out$sd1 == 0 & cont.out$sd2 == 0 & cont.out$se == 0) #Cont.results that have only zeros
missing.sds <- length(c(missing.sds, wrong.sds))

missing.ss <- which(is.na(data$total1) | is.na(data$total2))
wrong.ss <- which(data$total1 <= 0 | data$total2 <= 0)
missing.ssize <- length(c(missing.ss, wrong.ss))

missing.year <- sum(is.na(data$study.year)) + length(c(which(data$study.year < 1920), which(data$study.year > 2019)))

missing.table <- rbind("Missing mean values and mean differences" = missing.effects,
                         "Missing standard deviations and standard errors" = missing.sds,
                         "Missing sample sizes" = missing.ssize,
                         "Missing study year" = missing.year)

#Dataset properties:
#Review level
study.rev <- data %>% group_by(file.nr) %>% distinct(study.name) %>% count() %>% ungroup() %>% 
	summarise(lower = quantile(n, 0.05), median = quantile(n, 0.5), mean = mean(n), upper = quantile(n, 0.95))
comparison.rev <- data %>% group_by(file.nr) %>% distinct(comparison.name) %>% count() %>% ungroup() %>% 
	summarise(lower = quantile(n, 0.05), median = quantile(n, 0.5), mean = mean(n), upper = quantile(n, 0.95))
meta.rev <- data.ext %>% group_by(file.nr) %>% distinct(meta.id) %>% count() %>% ungroup() %>% 
	summarise(lower = quantile(n, 0.05), median = quantile(n, 0.5), mean = mean(n), upper = quantile(n, 0.95))

#Global level
study.years <- data.ext %>% group_by(file.nr) %>% distinct(study.name, .keep_all = T) %>% ungroup() %>% 
	summarise(lower = quantile(study.year, 0.05, na.rm = T), median = quantile(study.year, 0.5, na.rm = T),
						mean = mean(study.year, na.rm = T), upper = quantile(study.year, 0.95, na.rm = T))
sample.size <- data.ext %>% group_by(file.nr) %>% distinct(study.name, .keep_all = T) %>% ungroup() %>% 
	mutate(samplesize = total1 + total2) %>% 
	summarise(lower = quantile(samplesize, 0.05, na.rm = T), median = quantile(samplesize, 0.5, na.rm = T),
						mean = mean(samplesize, na.rm = T), upper = quantile(samplesize, 0.95, na.rm = T))

dataset.properties <- rbind(study.rev, comparison.rev, meta.rev, study.years, sample.size)
rownames(dataset.properties) <- c("Study number", "Comparison number", "Group number", "Study years", "Study sample size")
colnames(dataset.properties) <- c("5% quantile", "median", "mean", "95% quantile")

#Pooling studies
cum.repr.trials.subg <- data %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>%
  filter(n < 10) %>%
  full_join( data %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
               filter(n > 9) %>% ungroup %>% summarise(n = 10, nn = sum(nn))) %>%
  ungroup() %>% arrange(desc(n)) %>% mutate(csum  = cumsum(nn)) %>% arrange(n)

colnames(cum.repr.trials.subg)  <- c("n","Number of groups", "Cumulative sum of groups")

#Loading meta-analysis results:
file.bin <- "pb.bin.RData"
load(file.path(PATH_RESULTS, file.bin))

file.cont <- "pb.cont.RData"
load(file.path(PATH_RESULTS, file.cont))

file.meta <- "meta.RData"
load(file.path(PATH_RESULTS, file.meta))

library(gridExtra)

#Meta filtering: 
metac.bin <- meta.bin %>% filter(n.sig.type.bin > 1) %>% filter((se.max^2)/(se.min^2) > 4) %>% filter(I2 < 0.5)
metac.cont <- meta.cont %>% filter(n.sig.type.cont > 1) %>% filter((se.max^2)/(se.min^2) > 4) %>% filter(I2 < 0.5)
metac <- meta %>% filter(n.sig.type > 1) %>% filter((se.max^2)/(se.min^2) > 4) %>% filter(I2 < 0.5)

#Significant proportion
test.bin <- metac.bin %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                          schwarzer.test = mean(schwarzer.test),
                                                          rucker.test = mean(rucker.test),
                                                          harbord.test = mean(harbord.test),
                                                          peter.test = mean(peter.test))
test.bin <- test.bin %>% gather(key = "test.type", value = "mean")

p.bin <- metac.bin %>% ungroup() %>% 
  select(egger.test, schwarzer.test, rucker.test, harbord.test, peter.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>%  
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) + ggtitle("Binary Outcomes") + theme(legend.position = "top") +
  guides(fill=guide_legend(title=NULL))+
  annotate("text", x = test.bin$test.type, y = 500, 
           label = paste(round(test.bin$mean, 2)*100, "% rejected"), 
           color = "white")

test.cont <- metac.cont %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                            begg.test = mean(begg.test),
                                                            thomson.test = mean(thomson.test))

test.cont <- test.cont %>% gather(key = "test.type", value = "mean")

p.cont <- metac.cont %>% ungroup() %>% 
  select(egger.test, thomson.test, begg.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>% 
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) + ggtitle("Continuous Outcomes") + theme(legend.position = "top") +
	guides(fill=guide_legend(title=NULL))+
  annotate("text", x = test.cont$test.type, y = 150, 
           label = paste(round(test.cont$mean, 2)*100, "% rejected"), 
           color = "white")

dat_text <- data.frame(
  label = paste(c(sum(metac.cont$begg.test), sum(metac.cont$egger.test), sum(metac.cont$thomson.test)), "> 0.05"),
  test.type   = c("pval.begg.cont", "pval.egger.cont", "pval.thomson.cont"))

p.dist.cont <- metac.cont %>% ungroup() %>% 
  select(pval.egger.cont, pval.thomson.cont, pval.begg.cont) %>% 
  gather(key = "test.type", value = "p.value") %>% 
  ggplot(aes(x = p.value)) + geom_histogram(bins = 20) + theme_bw() + facet_wrap(~test.type) + 
	geom_text(data = dat_text, mapping = aes(x = -Inf, y = -Inf, label = label), hjust   = -1,
  vjust   = -4, color = "white")

dat_text <- data.frame(
  label = paste(c(sum(metac.bin$egger.test), sum(metac.bin$harbord.test), sum(metac.bin$peter.test), 
  										sum(metac.bin$rucker.test), sum(metac.bin$schwarzer.test)), "> 0.05"),
  test.type   = c("pval.egger.bin", "pval.harbord.bin", "pval.peter.bin", "pval.rucker.bin", "pval.schwarzer.bin"))

p.dist.bin <- metac.bin %>% ungroup() %>% 
  select(pval.egger.bin, pval.harbord.bin, pval.peter.bin, pval.rucker.bin, pval.schwarzer.bin) %>% 
  gather(key = "test.type", value = "p.value") %>% 
  ggplot(aes(x = p.value)) + geom_histogram(bins = 20) + theme_bw() + facet_wrap(~test.type) + 
	geom_text(data = dat_text, mapping = aes(x = -Inf, y = -Inf, label = label), hjust   = -0.65,
  vjust   = -4, color = "white")

#Test agreement
agree.bin <- meta.bin %>% mutate(n.sig = peter.test + rucker.test + egger.test + harbord.test + schwarzer.test) %>% 
  group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
  ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Binary Outcomes")

agree.cont <- meta.cont %>% mutate(n.sig = egger.test + thomson.test + begg.test) %>% 
  group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
  ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Continuous Outcomes")

#Adjustment example plots regression:
reg.1 <- qplot(x = biased.rev$se, y = biased.rev$effect, xlim = c(0, 1.2)) + geom_point() + 
	stat_smooth(method="lm", fullrange=TRUE, se = F) + coord_flip() + theme_bw() +
	ylab("Std. Mean Difference") + xlab("Standard error")

reg.2 <- qplot(x = unbiased.rev$se, y = unbiased.rev$effect, xlim = c(0, 6.5)) + geom_point() + 
	stat_smooth(method="lm", fullrange=TRUE, se = F) + coord_flip() + theme_bw() +
	ylab("Std. Mean Difference") + xlab("Standard error")

@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{document}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Cochrane Library}
Database of high-quality, systematic reviews in clinical science.

Currently $\sim$ 8,000 reviews, prepared by independent groups.

Reviews are peer-reviewed and prepared after guidelines.
\end{frame}

\begin{frame}{Cochrane Library Dataset}
\Sexpr{format(length(unique(data$file.nr)), big.mark=",")} systematic reviews with studies published until 2018.

\Sexpr{format(length(unique(data$study.name)), big.mark=",")} studies.

\Sexpr{format(dim(data)[1], big.mark=",")} study results.
\end{frame}

\begin{frame}{Dataset Structure}

\begin{figure}
\tikzstyle{every node}=[draw=black,thick,anchor=west,scale=.65]
\tikzstyle{selected}=[draw=red,fill=red!30]
\tikzstyle{optional}=[dashed,fill=gray!50]
\begin{tikzpicture}
[grow = right, anchor = west,
  growth parent anchor=east, % added code
  parent anchor=east, level distance=.5cm,
  sibling distance=2em, level 1/.style={sibling distance=2em}, level 2/.style={sibling distance=2em},
  level 3/.style={sibling distance=2em}, level 4/.style={sibling distance=1.2em}]
  \node {Review} [edge from parent fork right]
    child { node {Comparison 2}
      child { node {Outcome 2}}
      child { node {Outcome 1}
        child { node {Subgroup 2}}
        child { node {Subgroup 1}
          child  { node {Result 3}}
          child  { node {Result 2}}
          child  { node {Result 1}}
          }}
    }
    child [missing] {}
    child { node {Comparison 1}};
\end{tikzpicture}
%\caption{Structure of a hypothetical review with two different comparisons\label{review.structure}}
\label{review.structure}
\end{figure}
\end{frame}

\begin{frame}{Dataset Structure}
\begin{itemize}
\item Comparison: What is compared, e.g. treatment vs. control
\item Outcome: How it is compared
\item Subgroup: Subgroup affiliation
\item Meta-Analysis Group: Results from same comparison, outcome and subgroup
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Review Example: binary outcome}
Barbiturate efficacy for head injury treatment
\vspace{-5mm}
<<echo = FALSE, results = 'asis'>>=
print(xtable(barbi2, label = "barbiturates", digits = 0), include.rownames = F, size = "tiny")
@
\end{frame}

\begin{frame}[fragile]{Dataset Properties}
Missing data:
<<echo=FALSE, results = 'asis'>>=
print(xtable(missing.table, align = "lr"), include.rownames = T, include.colnames = F)
@
\end{frame}

\begin{frame}[fragile]{Dataset Properties}
Review and study properties:
<<echo=FALSE, results = 'asis'>>=
print(xtable(dataset.properties, digits = 0, align = "lrrrr"), include.rownames = T, size = "footnotesize", hline = c(0,1,1,4))
@
\end{frame}

\begin{frame}[fragile]{Pooling Studies - Meta-Analysis}
Multiple results in a meta-analysis group can be pooled:
<<echo=FALSE, results = 'asis'>>=
print(xtable(cum.repr.trials.subg, label = "repr.groups", align = "lrrr", digits = 0), include.rownames = F, size = "footnotesize")
@
\end{frame}

\begin{frame}{Meta-analysis}
Benefits:
\begin{itemize}
\item Summary of evidence (e.g. of a treatment)
\item More reliable evidence (?)
\end{itemize}

Assumptions:
\begin{itemize}
\item Identical study settings (can be relaxed)
\item Random sample of studies
\end{itemize}
\end{frame}

\begin{frame}{Small Study Effects}
``The tendency for the smaller studies to show larger treatment effects'' \citep{Sterne}
\end{frame}

\begin{frame}{Small Study Effects}
Causes:
\begin{itemize}
\item Selective publication of studies with large effects
\item Bias in smaller studies
\item Systematical differences in study settings
\item \ldots
\end{itemize}
\end{frame}

\begin{frame}{Small Study Effect Tests}
Tests applicable if:
\begin{itemize}
\item $n$ large
\item variation in the estimated variances of effects \\(here: $\frac{\sigma_{\textrm{max}}}{\sigma_{\textrm{min}}} > 4$)
\end{itemize}
Adjustments required if variance is dependent on effect size (e.g. log odds ratios)
\end{frame}

\begin{frame}[fragile]{Small Study Effect Tests}
Funnel plots (continuous outcome examples):

\vspace{-1.2cm}

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1, 2))
funnel(meta.example)
funnel(meta.nexample)
@

\end{frame}

\begin{frame}[fragile]{Regression based Tests}
Radial plots (continuous outcome examples):

\vspace{-1.2cm}

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1,2))
radial(meta.example, comb.fixed = F)
radial(meta.nexample, comb.fixed = F)
@

\end{frame}

\begin{frame}{Regression based Tests}
Continuous outcomes tests:

$i$ being the $i$th study in a meta-analysis:

Let $y_{i} = \textrm{effect}_{i}/\textrm{se}_{i}$ and $x_{i} = 1/\textrm{se}_{i}$
\begin{itemize}
\item \citet{Egger} : \\ $y_{i} \sim N(\beta_{0} + \beta_{1}x_{i}, v_{i})$
\item \citet{thompson.sharp} : \\ $y_{i} \sim N(\beta_{0} + \beta_{1}x_{i}, v_{i} + \tau^2)$
\end{itemize}
%Test for non-zero intercept $\beta_{0}$
\end{frame}

\begin{frame}[fragile]{Egger's Tests}

\vspace{-6cm}

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1,2))
metabias(meta.example, plotit = TRUE, method.bias = "linreg")
metabias(meta.nexample, plotit = TRUE, method.bias = "linreg")
@

\end{frame}

\begin{frame}[fragile]{Thompson and Sharp's Tests}

\vspace{-6.5cm}

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height= 4>>=
par(las = 1, mfrow = c(1,2))
metabias(meta.example, plotit = TRUE, method.bias = "mm")
metabias(meta.nexample, plotit = TRUE, method.bias = "mm")
@

\end{frame}

\begin{frame}{Regression based Tests}
Adjustments for binary outcomes:
\begin{itemize}
\item \citet{Peters} :$x_{i}$ = inverse of total sample size, $\textrm{variance}_{i}$ as weight.
\item \citet{Harbord} :$x:{i}$ = score of the log-likelihood of a proportion, $\textrm{variance}_{i}$ as weight.
\item \citet{Rucker} :Use arcsine variance stabilizing transformation for variances and effects, do e.g. Egger's test.
\end{itemize}
\end{frame}

\begin{frame}{Rank based tests}
\citet{begg.ties}
Let $y_{i}$ the standardized effect size of a study $i$, $v_{i}$ it's variance and $n$ the number of studies

$u$ the number of pairs $(y_{i}, v_{i})$ ranked in the same order, $l$
the number of pairs in the opposite order

$Z = (u - l)/\sqrt{n(n-1)(2n + 5)/18}$ \\ can be used as a test statistic (based on Kendall's Tau)
\end{frame}

\begin{frame}{Rank based tests}
\citet{Schwarzer}
Let $e_{t}$ the number of events in the treatment group

Given constant log odds ratio, $E_{t}$ follows a hypergeometric distribution.

$\mathbb{E}(E_{t})$ and the variance can be estimated and used as in 
\citet{begg.ties}
\end{frame}

\begin{frame}{Test Results}
Application of tests if:
\begin{itemize}
\item $n \geq 10$
\item at least one statistically significant effect in a study
\item $\frac{\sigma_{\textrm{max}}}{\sigma_{\textrm{min}}} > 4$
\item $I^2 < 0.5$
\end{itemize}

From \Sexpr{dim(meta)[1]} with $n \geq 10$, \Sexpr{dim(metac)[1]} remain.
\end{frame}

% \begin{frame}[fragile]{Test Results: Significance}
% 
% <<echo = FALSE, fig.height = 4>>=
% grid.arrange(p.bin, p.cont, ncol = 2)
% @
% \end{frame}

\begin{frame}[fragile]{Continuous Outcome Test Results}
$p$-values distribution:

<<echo = FALSE, fig.height = 4, message = FALSE>>=
plot(p.dist.cont)
@
\end{frame}

\begin{frame}[fragile]{Binary Outcome Test Results}
$p$-values distribution:

<<echo = FALSE, fig.height = 4, message = FALSE>>=
plot(p.dist.bin)
@
\end{frame}

\begin{frame}[fragile]{Agreement in significance}
Number of significant test results per meta-analysis:

<<echo = FALSE, fig.height = 3.8, message = FALSE>>=
grid.arrange(agree.cont, agree.bin, ncol = 2)
@
\end{frame}

\begin{frame}{Small Study Effect Adjustment}
Three methods:
\begin{itemize}
\item Regression
\item Copas selection model
\item Trim-and-fill
\end{itemize}
\end{frame}

\begin{frame}{Adjustment by regression}
Similar to the tests, but with unnormalized effect $y_{i}$:

\begin{align}
y_{i} = \beta_{0} + \beta_{1}x_{i}
\end{align}

$\beta_{0}$	 corresponds to $y_{i}$ with $x_{i} = 0$
\end{frame}

\begin{frame}[fragile]{Adjustment by regression}
Linear regression method:

<<echo = FALSE, fig.height = 3>>=
grid.arrange(reg.1, reg.2, ncol = 2)
@
\end{frame}

\begin{frame}[fragile]{Shrinkage Regression}
Extended random effects model:

\vspace{-4mm}
\begin{align}
y_{i} = \beta_{0} + \beta_{1}(\sqrt{v_{i} + \tau^2}) + \epsilon_{i}(\sqrt{s_{i} + \tau^2}),  
\epsilon_{i} \stackrel{\textrm{iid}}{\sim} N(0,1) \nonumber
\end{align}

\vspace{-8mm}
\begin{align}
\mathbb{E}(y_{i}) \rightarrow \beta_{0} + \beta_{1}\tau \textrm{ if } \sqrt{v_{i}} \rightarrow 0 \nonumber
\end{align}

$\beta_{0}, \beta_{1}$ and $\tau$ can be estimated e.g. by ML and REML.
\end{frame}

\begin{frame}[fragile]{Shrinkage Regression}
shrinking the within study variance:

\vspace{-4mm}
\begin{align}
y_{i} &= \beta_{0}^\star + \beta_{1}^\star(\sqrt{v_{i}/M + \tau^2}) + \epsilon_{i}(\sqrt{s_{i}/M + \tau^2}) \nonumber
\end{align}

Letting $M \rightarrow \infty$ and substituting for all parameters and the observed residual

\vspace{-8mm}
\begin{align}
y_{\infty,i} &= \beta_{0}^\star + \sqrt{\frac{\tau^2}{v_{i} + \tau^2}}(y_{i} - \beta_{0}^\star)
\end{align}

\end{frame}

\begin{frame}[fragile]{Shrinkage Regression}
Three different treatment effect estimates:

\begin{align}
y_{\infty,i} &= \beta_{0}^\star + \sqrt{\frac{\tau^2}{v_{i} + \tau^2}}(y_{i} - \beta_{0}^\star)
\end{align}

\end{frame}

\begin{frame}{References}
  \small
  \bibliographystyle{apalike}
\bibliography{illustration}
\end{frame}

%\appendix
%% Possible backup slides...

%% chapter division is accomplished with:
%% \part{Appendix}

\end{document}