### 2019-06-11 communication/Complete_Analysis.Rmd
Methods update

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

```{r, echo=FALSE}
rm(list = ls())
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
PATH2 = file.path(PATH_HOME, 'PubBias')
FILE = 'cochrane_2018-06-09.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH2, 'code')
PATH_RESULTS = file.path(PATH2, 'results')
PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

file_results = "pb.RData"

source(file.path(PATH_CODE, 'PubBias_functions.R'))

file.dat <- "data.RData"
if (file.exists(file.path(PATH_RESULTS, file.dat))) {
	load(file.path(PATH_RESULTS, file.dat))
} else {
	data = pb.readData(path = PATH_DATA, file = FILE)
	tmp = pb.clean(data)
	data = tmp[[1]]
	aliases = tmp[[2]]
	save(data, file =  file.path(PATH_RESULTS, file.dat))
}

load(file.path(PATH_RESULTS, "meta_analyses_summary_complete.RData"))
load(file.path(PATH_RESULTS, "meta.bin.RData"))
load(file.path(PATH_RESULTS, "meta.cont.RData"))
load(file.path(PATH_RESULTS, "meta.surv.RData"))
load(file.path(PATH_RESULTS, "data_used_for_analysis.RData"))
```

##Main analysis

```{r}
sig.level <- 0.1
meta.f <- meta.f %>% 
	mutate(egger.test = ifelse(pval.egger < sig.level, 1, 0),
				 thompson.test = ifelse(pval.thompson < sig.level, 1, 0),
				 begg.test = ifelse(pval.begg < sig.level, 1, 0),

				 schwarzer.test = ifelse(pval.schwarzer < sig.level, 1, 0),
				 rucker.test = ifelse(pval.rucker < sig.level, 1, 0),
				 harbord.test = ifelse(pval.harbord < sig.level, 1, 0),
				 peter.test = ifelse(pval.peter < sig.level, 1, 0))

meta.bin <- meta.f %>% filter(outcome.type == "bin")
meta.cont <- meta.f %>% filter(outcome.type == "cont")
meta.surv <- meta.f %>% filter(outcome.type == "surv")

#Overall proportions, to be finsihed yet:
# test <- meta.f %>% group_by(outcome.type) %>% summarize(
#   schwarzer.test = mean(schwarzer.test),
#   rucker.test = mean(rucker.test),
#   harbord.test = mean(harbord.test),
#   peter.test = mean(peter.test),
#   egger.test = mean(egger.test),
#   begg.test = mean(begg.test),
#   thompson.test = mean(thompson.test))
# test <- test %>% gather(key = "test.type", value = "mean", schwarzer.test:thompson.test) 
# # test$mean[which(!is.na(test$mean))] <- round(test$mean[which(!is.na(test$mean))], 2)
# 
# #Compare proportions of total concordance vs. proportion of at least single positive test results among all
# meta.f %>% ungroup() %>% 
#   select(outcome.type, schwarzer.test, rucker.test, harbord.test, peter.test, egger.test, begg.test, thompson.test) %>% 
#   gather(key = "test.type", value = "null.hypothesis", schwarzer.test:thompson.test) %>%  
#   mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
#   ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
#   theme_bw() + xlab(label = NULL) + facet_wrap(~outcome.type, scales = "free") + theme(legend.position = "top") +
#   guides(fill=guide_legend(title=NULL)) 
# 
# # + annotate("text", x = test$test.type, y = 10,
# #            label = paste(round(test$mean, 2,)*100, "% rejected"),
# #            color = "white")

test.bin <- meta.bin %>% ungroup() %>% summarize(
																								 schwarzer.test = mean(schwarzer.test),
																								 rucker.test = mean(rucker.test),
																								 harbord.test = mean(harbord.test),
																								 peter.test = mean(peter.test))
test.bin <- test.bin %>% gather(key = "test.type", value = "mean")

p.bin <- meta.bin %>% ungroup() %>% 
	select(schwarzer.test, rucker.test, harbord.test, peter.test) %>% 
	gather(key = "test.type", value = "null.hypothesis") %>%  
	mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
	ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
	theme_bw() + xlab(label = NULL) + ggtitle("Binary Outcomes") + theme(legend.position = "top") +
	guides(fill=guide_legend(title=NULL))+
	annotate("text", x = test.bin$test.type, y = 500, 
					 label = paste(round(test.bin$mean, 2)*100, "% rejected"), 
					 color = "white")

dat_text <- data.frame(
	label = paste(c(sum(meta.bin$harbord.test), sum(meta.bin$peter.test), 
									sum(meta.bin$rucker.test), sum(meta.bin$schwarzer.test)), "< 0.1,",
								round(c(mean(meta.bin$harbord.test), mean(meta.bin$peter.test), 
												mean(meta.bin$rucker.test), mean(meta.bin$schwarzer.test)),2)*100, "%"),
	test.type   = c("pval.harbord", "pval.peter", "pval.rucker", "pval.schwarzer"))

labels <- c(pval.harbord = "Harbord", pval.peter = "Peter", pval.rucker = "Rucker", pval.schwarzer = "Schwarzer")

p.dist.bin <- meta.bin %>% ungroup() %>% 
	select(pval.harbord, pval.peter, pval.rucker, pval.schwarzer) %>% 
	gather(key = "test.type", value = "p.value") %>% 
	ggplot(aes(x = p.value)) + geom_histogram(bins = 20) + theme_bw() + facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
	geom_text(data = dat_text, mapping = aes(x = 0.5, y = 100, label = label), color = "black") + 
	theme(strip.text.x = element_text(size=15))

test.cont <- meta.cont %>% ungroup() %>% summarize(egger.test = mean(egger.test),
																									 begg.test = mean(begg.test),
																									 thompson.test = mean(thompson.test))

test.cont <- test.cont %>% gather(key = "test.type", value = "mean")

p.cont <- meta.cont %>% ungroup() %>% 
	select(egger.test, thompson.test, begg.test) %>% 
	gather(key = "test.type", value = "null.hypothesis") %>% 
	mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
	ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
	theme_bw() + xlab(label = NULL) + ggtitle("Continuous Outcomes") + theme(legend.position = "top") +
	guides(fill=guide_legend(title=NULL))+
	annotate("text", x = test.cont$test.type, y = 150, 
					 label = paste(round(test.cont$mean, 2)*100, "% rejected"), 
					 color = "white")

dat_text <- data.frame(
	label = paste(c(sum(meta.cont$begg.test), sum(meta.cont$egger.test), sum(meta.cont$thompson.test)), "< 0.1,",
								round(c(mean(meta.cont$begg.test), mean(meta.cont$egger.test), mean(meta.cont$thompson.test)),2)*100, "%"),
	test.type   = c("pval.begg", "pval.egger", "pval.thompson"))

labels <- c(pval.begg = "Begg Mazumdar", pval.egger = "Egger", pval.thompson = "Thompson Sharp")
p.dist.cont <- meta.cont %>% ungroup() %>% 
	select(pval.egger, pval.thompson, pval.begg) %>% 
	gather(key = "test.type", value = "p.value") %>% 
	ggplot(aes(x = p.value)) + geom_histogram(bins = 20) + theme_bw() + 
	facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
	geom_text(data = dat_text, mapping = aes(x = 0.5, y = 25, label = label),  color = "black") + 
	theme(strip.text.x = element_text(size=15))

test.surv <- meta.surv %>% ungroup() %>% summarize(egger.test = mean(egger.test),
																									 begg.test = mean(begg.test),
																									 thompson.test = mean(thompson.test))

test.surv <- test.surv %>% gather(key = "test.type", value = "mean")

p.surv <- meta.surv %>% ungroup() %>% 
	select(egger.test, thompson.test, begg.test) %>% 
	gather(key = "test.type", value = "null.hypothesis") %>% 
	mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
	ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
	theme_bw() + xlab(label = NULL) + ggtitle("Continuous Outcomes") + theme(legend.position = "top") +
	guides(fill=guide_legend(title=NULL))+
	annotate("text", x = test.surv$test.type, y = 150, 
					 label = paste(round(test.surv$mean, 2)*100, "% rejected"), 
					 color = "white")

dat_text <- data.frame(
	label = paste(c(sum(meta.surv$begg.test), sum(meta.surv$egger.test), sum(meta.surv$thompson.test)), "< 0.1,",
								round(c(mean(meta.surv$begg.test), mean(meta.surv$egger.test), mean(meta.surv$thompson.test)),2)*100, "%"),
	test.type   = c("pval.begg", "pval.egger", "pval.thompson"))

labels <- c(pval.begg = "Begg Mazumdar", pval.egger = "Egger", pval.thompson = "Thompson Sharp")
p.dist.surv <- meta.surv %>% ungroup() %>% 
	select(pval.egger, pval.thompson, pval.begg) %>% 
	gather(key = "test.type", value = "p.value") %>% 
	ggplot(aes(x = p.value)) + geom_histogram(bins = 20) + theme_bw() + 
	facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
	geom_text(data = dat_text, mapping = aes(x = 0.5, y = 25, label = label),  color = "black") + 
	theme(strip.text.x = element_text(size=15))

#Test agreement
agree.bin <- meta.bin %>% mutate(n.sig = peter.test + rucker.test + harbord.test + schwarzer.test) %>% 
	group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
	ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Binary Outcomes")

agree.cont <- meta.cont %>% mutate(n.sig = egger.test + thompson.test + begg.test) %>% 
	group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
	ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Continuous Outcomes")

agree.surv <- meta.surv %>% mutate(n.sig = egger.test + thompson.test + begg.test) %>% 
	group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
	ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Continuous Outcomes")

################################################################################################
################################################################################################
#CHANGE OF EFFECT SIZES AFTER ADJUSTMENT
################################################################################################
################################################################################################

#Comparison of z-scores:
sig.zcor <- meta.f %>% mutate(z.fixef = est.z.fixef/se.est.z.fixef,
																 z.ranef = est.z.ranef/se.est.z.ranef,
																 z.reg = est.z.reg/se.est.z.reg,
																 z.copas = est.z.copas/se.est.z.copas) %>%  
	select(z.fixef, z.ranef, z.reg, z.copas) %>% gather(key = "method", value = "fisher.z") %>% 
	group_by(method) %>% 
	summarise(significant = length(which(abs(fisher.z) > 1.96)),
						p.significant = significant/length(fisher.z))
sig.zcor <- data.frame(method = sig.zcor$method,
											 label = paste("n = ", sig.zcor$significant, ", ", round(sig.zcor$p.significant,3)*100, "% significant", sep = ""))

meta.f %>% mutate(z.fixef = est.z.fixef/se.est.z.fixef,
										 z.ranef = est.z.ranef/se.est.z.ranef,
										 z.reg = est.z.reg/se.est.z.reg,
										 z.copas = est.z.copas/se.est.z.copas) %>%  
	select(z.fixef, z.ranef, z.reg, z.copas) %>% gather(key = "method", value = "fisher.z") %>% 
	# filter(abs(fisher.z) < 10) %>% 
	ggplot(aes(x = abs(fisher.z))) + geom_histogram() + theme_bw() + facet_wrap(~method, ncol = 2) + 
	geom_vline(xintercept = 1.96, color = "red") + 
	geom_text(data = sig.zcor, aes(x = 30, y = 500, label = label), position = "dodge")

#Comparison of m.a. with copas correction method applied:
sig.zcor <- meta.f %>% filter(!is.na(est.z.copas)) %>% 
	mutate(z.fixef = est.z.fixef/se.est.z.fixef,
				 z.ranef = est.z.ranef/se.est.z.ranef,
				 z.reg = est.z.reg/se.est.z.reg,
				 z.copas = est.z.copas/se.est.z.copas) %>%  
	select(z.fixef, z.ranef, z.reg, z.copas) %>% gather(key = "method", value = "fisher.z") %>% 
	group_by(method) %>% 
	summarise(significant = length(which(abs(fisher.z) > 1.96)),
						p.significant = significant/length(fisher.z))
sig.zcor <- data.frame(method = sig.zcor$method,
											 label = paste("n = ", sig.zcor$significant, ", ", round(sig.zcor$p.significant,3)*100, "% significant", sep = ""))

meta.f %>% filter(!is.na(est.z.copas)) %>% 
	mutate(z.fixef = est.z.fixef/se.est.z.fixef,
				 z.ranef = est.z.ranef/se.est.z.ranef,
				 z.reg = est.z.reg/se.est.z.reg,
				 z.copas = est.z.copas/se.est.z.copas) %>%  
	select(z.fixef, z.ranef, z.reg, z.copas) %>% gather(key = "method", value = "fisher.z") %>% 
	# filter(abs(fisher.z) < 10) %>% 
	ggplot(aes(x = abs(fisher.z))) + geom_histogram() + theme_bw() + facet_wrap(~method, ncol = 2) + 
	geom_vline(xintercept = 1.96, color = "red") + 
	geom_text(data = sig.zcor, aes(x = 10, y = 125, label = label), position = "dodge")

#Comparison of "original effect sizes":
effect.diff <- meta.f %>% mutate(
	est.fixef = ifelse(outcome.type == "bin", exp(est.fixef), est.fixef),
	est.ranef = ifelse(outcome.type == "bin", exp(est.ranef), est.ranef),
	est.copas = ifelse(outcome.type == "bin", exp(est.copas), est.copas),
	est.reg = ifelse(outcome.type == "bin", exp(est.reg), est.reg)
)

effect.diff <- effect.diff %>% mutate(
	est.fixef = ifelse(outcome.type == "bin", est.fixef - 1, est.fixef),
	est.ranef = ifelse(outcome.type == "bin", est.ranef - 1, est.ranef),
	est.copas = ifelse(outcome.type == "bin", est.copas - 1, est.copas),
	est.reg = ifelse(outcome.type == "bin", est.reg - 1, est.reg),
	fixef.ranef = ifelse(abs(est.ranef) > abs(est.fixef), "Reduction", "Amplification"),
	fixef.copas = ifelse(abs(est.fixef) > abs(est.copas), "Reduction", "Amplification"),
	fixef.reg = ifelse(abs(est.fixef) > abs(est.reg), "Reduction", "Amplification"),
	fixef.copas.s = ifelse(sign(est.fixef) == sign(est.copas), "Unchanged", "Reversed"),
	fixef.reg.s = ifelse(sign(est.fixef) == sign(est.reg), "Unchanged", "Reversed"))

effect.diff <- effect.diff %>% 
  mutate(change.fixef.copas = ifelse(fixef.copas == "Reduction", -abs(est.fixef - est.copas),
                                     abs(est.fixef - est.copas)),
         change.fixef.reg = ifelse(fixef.ranef == "Reduction", -abs(est.fixef - est.reg),
                                   abs(est.fixef - est.reg)))

# hist.ranef <- effect.diff %>% ggplot(aes(x = abs(est.fixef - est.ranef))) +
# 	geom_histogram() + facet_grid(~factor(fixef.ranef)) + xlim(0, 2) + xlab("Effect difference") + 
# 	theme_bw() + theme(legend.position = "bottom", strip.text.x = element_text(size=15)) + ylim(0, 250)

#Copas selection model:
copas.miss <- effect.diff %>% group_by(outcome.type) %>% mutate(dif = abs(est.fixef - est.copas)) %>% 
	filter(dif > 2) %>% mutate(label = round(dif, 1)) %>% arrange(label) %>% select(label, meta.id, outcome.measure.new)
copas.label <- data.frame(label = paste("missing:", c(paste(copas.miss$label[copas.miss$outcome.type == "bin"], collapse = ", "),
																											paste(copas.miss$label[copas.miss$outcome.type == "cont"], collapse = ", "),
																											paste(copas.miss$label[copas.miss$outcome.type == "surv"], collapse = ", "))),
													outcome.type = c("bin", "cont", "surv"), 
													outcome.measure.new = c("Risk Ratio", "Mean Difference", "Hazard Ratio"))
copas.label$label <- as.character(copas.label$label)
copas.label$label[2] <- sub("2.8,", "2.8,\n", copas.label$label[2])

effect.diff %>% ggplot(aes(x = change.fixef.copas, fill = outcome.measure.new)) + 
	facet_wrap(outcome.type~., scales = "free", labeller = labeller(outcome.type = NULL)) +
	geom_histogram() + xlim(-2, 2) + xlab("Effect change") + 
	theme_bw() + theme(legend.position = "bottom", strip.text = element_blank(), legend.title = element_blank())  +
	geom_text(data = copas.label, 
						mapping = aes(x = 0, y = c(220, 47.4, 10), label = label, vjust = "center"), position = "dodge") #+ ylim(0, 250)

#Regression model:
reg.miss <- effect.diff %>% group_by(outcome.type) %>% mutate(dif = abs(est.fixef - est.reg)) %>% 
	filter(dif > 2) %>% mutate(label = round(dif, 1)) %>% arrange(label) %>% select(label, meta.id, outcome.measure.new)
reg.label <- data.frame(label = paste("missing:", c(paste(reg.miss$label[reg.miss$outcome.type == "bin"], collapse = ", "),
																										paste(reg.miss$label[reg.miss$outcome.type == "cont"], collapse = ", "),
																										paste(reg.miss$label[reg.miss$outcome.type == "surv"], collapse = ", "))),
												outcome.type = c("bin", "cont", "surv"), 
												outcome.measure.new = c("Risk Ratio", "Mean Difference", "Hazard Ratio"))
reg.label$label <- as.character(reg.label$label)
reg.label$label[2] <- sub("2.2,","2.2,\n ", reg.label$label[2])
reg.label$label[2] <- sub("2.6,","2.6,\n ", reg.label$label[2])
reg.label$label[2] <- sub(", 3.1,",", 3.1,\n ", reg.label$label[2])
reg.label$label[2] <- sub("4.4, 4.4,", "4.4, 4.4,\n", reg.label$label[2])
reg.label$label[2] <- sub(", 9.5,", ", 9.5,\n", reg.label$label[2])

reg.label$label[1] <- sub(", 3.3,", ", 3.3,\n", reg.label$label[1])
reg.label$label[1] <- sub(", 7.6,", ", 7.6,\n", reg.label$label[1])

effect.diff %>% ggplot(aes(x = est.fixef - est.reg, fill = outcome.measure.new)) + 
	facet_wrap(outcome.type~., scales = "free", labeller = labeller(outcome.type = NULL)) +
	geom_histogram() + xlim(-2, 2) + xlab("Effect change") + 
	theme_bw() + theme(legend.position = "bottom", strip.text = element_blank(), legend.title = element_blank())  +
	geom_text(data = reg.label, 
						mapping = aes(x = 0, y = c(330, 50, 10), label = label, vjust = "center"), position = "dodge") #+ ylim(0, 250)

################################################################################################
################################################################################################
#CHANGE OF SIGNIFICANT TREATMENT EFFECT AFTER ADJUSTMENT
################################################################################################
################################################################################################

#Explanation of coding: 0 = (unsig, unsig), 1 = (unsig, sig), 2 = (sig, sig), 3 = (sig, unsig)

sig.level <- 0.05
meta.f <- meta.f %>% rowwise() %>% 
	mutate(
		pval.copas = 2*(1-pnorm(abs(est.copas/se.est.copas))),

		sig.fixef = ifelse(pval.fixef > sig.level, 0, 1),
		sig.ranef = ifelse(pval.ranef > sig.level, 0, 1),
		sig.reg = ifelse(pval.reg > sig.level, 0, 1),
		sig.copas = ifelse(pval.copas > sig.level, 0, 1),

		sig.change.ranef.reg = sig.change(sig.before = sig.ranef, sig.after =  sig.reg), #Function to register if significance of estimate has changed after correction
		sig.change.ranef.copas = sig.change(sig.before = sig.ranef, sig.after =  sig.copas),
		sig.change.fixef.reg = sig.change(sig.before = sig.fixef, sig.after =  sig.reg),
		sig.change.fixef.copas = sig.change(sig.before = sig.fixef, sig.after =  sig.copas))

################################################################################################
################################################################################################
#PUBLICATION BIAS TEST RESULTS ANALYSIS
################################################################################################
################################################################################################

#Agreement proportions of publication bias tests:

#Binary:
meta.bin <- meta.bin %>%ungroup() %>%  mutate(egger.schwarzer = ifelse(egger.test == schwarzer.test, "agree", "disagree"),
																							egger.peter = ifelse(egger.test == peter.test, "agree", "disagree"),
																							egger.rucker = ifelse(egger.test == rucker.test, "agree", "disagree"),
																							egger.harbord = ifelse(egger.test == harbord.test, "agree", "disagree"),
																							schwarzer.peter = ifelse(schwarzer.test == peter.test, "agree", "disagree"),
																							schwarzer.rucker = ifelse(schwarzer.test == rucker.test, "agree", "disagree"),
																							schwarzer.harbord = ifelse(schwarzer.test == harbord.test, "agree", "disagree"),
																							rucker.peter = ifelse(rucker.test == peter.test, "agree", "disagree"),
																							rucker.harbord = ifelse(rucker.test == harbord.test, "agree", "disagree"),
																							harbord.peter = ifelse(harbord.test == peter.test, "agree", "disagree"))
agreement.bin <- meta.bin %>% ungroup() %>% summarise(egger.schwarzer = sum(egger.schwarzer == "agree")/n(),
																											egger.peter = sum(egger.peter == "agree")/n(),
																											egger.rucker = sum(egger.rucker == "agree")/n(),
																											egger.harbord = sum(egger.harbord == "agree")/n(),
																											schwarzer.peter = sum(schwarzer.peter == "agree")/n(),
																											schwarzer.rucker = sum(schwarzer.rucker == "agree")/n(),
																											schwarzer.harbord = sum(schwarzer.harbord == "agree")/n(),
																											rucker.peter = sum(rucker.peter == "agree")/n(),
																											harbord.peter = sum(harbord.peter == "agree")/n())
correlation.bin <- meta.bin %>%ungroup() %>%  summarise(egger.schwarzer = cor(pval.egger, pval.schwarzer),
																												egger.peter = cor(pval.egger, pval.peter),
																												egger.rucker = cor(pval.egger, pval.rucker),
																												egger.harbord = cor(pval.egger, pval.harbord),
																												schwarzer.peter = cor(pval.schwarzer, pval.peter),
																												schwarzer.rucker = cor(pval.schwarzer, pval.rucker),
																												schwarzer.harbord = cor(pval.schwarzer, pval.harbord),
																												rucker.peter = cor(pval.rucker, pval.peter),
																												harbord.peter = cor(pval.harbord, pval.peter))

binary.tests.agreement <- rbind(agreement.bin, correlation.bin)
rownames(binary.tests.agreement) <- c("Test Agreement","P-value Correlation")

#Continuous:
meta.cont <- meta.cont %>% ungroup() %>% mutate(thompson.egger = ifelse(thompson.test == egger.test, "agree", "disagree"),
																								thompson.begg = ifelse(thompson.test == begg.test, "agree", "disagree"),
																								egger.begg = ifelse(egger.test == begg.test, "agree", "disagree"))
agreement.cont <- meta.cont %>% ungroup() %>%  summarise(thompson.egger = sum(thompson.egger == "agree")/n(),
																												 thompson.begg = sum(thompson.begg == "agree")/n(),
																												 egger.begg = sum(egger.begg == "agree")/n())
correlation.cont <- meta.cont %>% ungroup() %>% summarise(thompson.egger = cor(pval.thompson, pval.egger),
																													thompson.begg = cor(pval.thompson, pval.begg),
																													egger.begg = cor(pval.egger, pval.begg))

cont.tests.agreement <- rbind(agreement.cont, correlation.cont)
rownames(cont.tests.agreement) <- c("Test Agreement","P-value Correlation")
```

```{r}

```