### 2019-06-24 report/MSc_Report.Rnw
Results first draft

\documentclass[11pt,a4paper,twoside]{book}
\input{header.sty}   % packages, layout and standard macros

\usepackage{verbatim}
\usepackage{tikz}
\usepackage{amssymb}
\usetikzlibrary{trees}

\begin{document}
\input{title}

<<'main-setup',include=FALSE,purl=FALSE>>=
library(knitr)
opts_chunk$set(
    concordance=TRUE,
    child.command = 'include',
    cache=TRUE)
@

\graphicspath{{./figure/}}
\DeclareGraphicsExtensions{.pdf,.png}
\setcounter{tocdepth}{1}

\thispagestyle{empty}
\begin{center}
  \vspace*{6cm}{\bfseries\Huge
  Publication Bias in the Cochrane Database of Systematic Reviews
  }
  \vfill
  \rm

  \LARGE
  Giuachin Kreiliger\\[12mm]

  \normalsize
  Version \today
\end{center}
\newpage
\thispagestyle{empty}~
\newpage
\pagenumbering{roman}

\thispagestyle{plain}\markboth{Contents}{Contents}
\tableofcontents
\setkeys{Gin}{width=.8\textwidth}

\chapter*{Preface}
\addtocontents{toc}{\protect \vspace*{13.mm}}
\addcontentsline{toc}{chapter}{\bfseries{Preface}}
\thispagestyle{plain}\markboth{Preface}{Preface}

% Howdy!

\bigskip

\begin{flushright}
  Giuachin Kreiliger\\
  June 2019
\end{flushright}

\addtocontents{toc}{\protect \vspace*{10mm}}

\cleardoublepage
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter01', child='chapter01.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter02', child='chapter02.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter03', child='chapter03.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter04', child='chapter04.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%<<'child-chapter05', child='chapter05.Rnw'>>=
%@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix
<<'child-chapterA1', child='chapterA1.Rnw'>>=
@

\cleardoublepage
\phantomsection
\addtocontents{toc}{\protect \vspace*{10mm}}
\addcontentsline{toc}{chapter}{\bfseries Bibliography}

\bibliographystyle{mywiley} 
\bibliography{biblio}

\cleardoublepage

\end{document}