### 2019-07-12 report/chapter01.Rnw
Report based on the new data

% LaTeX file for Chapter 01
<<'preamble01',include=FALSE>>=
library(knitr)
opts_chunk$set(
    fig.path='figure/ch01_fig', 
    self.contained=FALSE,
    cache=TRUE
)

@

\chapter{Introduction}

Studies get more attention and are more likely to be published, read and cited if they are implying large effects. Studies with no evidence for an effect are less likely to get published. This generates a bias called ``publication bias'', which distorts the view of the evidence for an effect. Recently, publication bias has been identified as one of the major concerns in irreproducible research \citep{Bishop.2019}. The issue has been pin-pointed in clinical science by many researchers (\citet{pb.clinicalscience1}, \citet{sterne2001}). In medical research, clinical trials study the efficacy of therapies and drugs. The gold standard in such intervention studies are randomized, (placebo) controlled, trials (RCTs). Results from such clinical studies influence the treatment of patients in daily clinical practice. The results from multiple intervention studies are summarized in meta-analyses and reflect the overall evidence for a treatment effect. Summarized treatment effects and uncertainities should be more robust and reliable. Publication bias can bias the overall effect estimates from meta-analyses and eventually lead to ineffective treatments that could lead to patient harm. \\
There is extensive literature on publication bias in meta-analyses, e.g. (\citet{pb.clinicalscience.2013}, %\citet{Hopewell02}, 
\citet{grey.literature.4}, \citet{grey.literature.3}, \citet{grey.literature.2}). 
All authors agree in that the exclusion of unpublished results in meta-analyses is responsible for overestimation of treatment effects. Although there are policies that make it mandatory to make accessible all study results \citep{fda}, it is not clear if the situation has improved yet.\\
The Cochrane Organisation has specialized on systematic reviews of healthcare interventions. Researchers that write a systematic review collect data across studies, review them and try to provide up-to-date information about specific treatment efficacies \citep{cochrane.handbook}. By extensive literature scrutinization, they try to circumvent the issue of publication bias. Earlier research however suggests that the efforts are only partially succesful, and that there still is publication bias within the reviews (\citet{Egger}, \citet{Ioannidis2007}, \citet{kicinsky}, \citet{vanAert.2019}). In this publications, the results of a part of the Cochrane Database of systematic reviews is analysed with small study effect tests or bayesian hierarchical selection models \citep{kicinsky}. They all find moderate to large evidence for publication bias in the Database. A more thorough discussion of their methods and results will follow in chapter \ref{ch:Discussion}. However, none of the research so far has estimated the amount and impact of publication bias on meta-analytical findings thoroughly and with the most suitable methods. Also, the results are often presented in the form of dichotomized hypothesis tests, a practice known to contribute to publication bias itself (cite sb).\\
The aim of this thesis is to use prevailed methods to detect publication bias, and make use of the full amount of data that the Cochrane Organisation provides. At the end, an up-to-date estimate of the prevalence of publication bias in the data shall be given, as well as its implications regarding the effect sizes of the treatments. To achieve this, methods to detect and adjust for publication bias in meta-analysis are applied on the data.

Already in 1978, scientists raised concern that effect sizes could be overestimated systematically if journals were ``not willing to drop their policy to use statistical significance as one of their criteria for publication'' \citep{dunlap1978}. Scientific findings get more attention and are more likely to be published, read and cited if they are implying large effects. The issue has been pin-pointed in clinical science by many researchers (\citet{pb.clinicalscience1}, \citet{sterne2001}). The concern is that overestimated efficacy of the treatments could lead to harm to patients. \\
A measure to counter the presence of single, overestimated study results is a meta-analysis \citep{metaanalysis}, a practice originally developped in psychology. Summarized treatment effects and uncertainities should be more robust and reliable. The pitfall is that when the entire body of evidence is again a biased sample of studies with high effects, the estimate is again overestimating the true effect. 
% 
% 
% Also after the introduction of mandatory registration of clinical trials
% 
% 
% 
% Typically, we expect empirical scientific results to be distorted by random noise, such that the results are not equal to the real process that they describe. Additionally, one can reasonably assume that the results are also biased to some extent, as the statistical paradigm of precision versus bias would predict: The smaller the noise part of distortion, the more the result should be distorted by bias. However, the trade-off is not inevitable and can be minimized by scientific rigor and foresightfull experimental design. Large efforts have been made both in theory and practice to improve the quality of experiments and their analysis, such as defining experimental settings that prohibit influence of expectations of the researcher (e.g. blinding) or increase sample size in experiments. The introduction of randomized clinical trials is for example seen nowadays as a benchmark in clinical science, heavily improving the reliability of its findings. \\
% It is more and more considered a new scientific field by itself to think of and argue about circumstances that improve the quality of scientific findings. Statistics is in some sense predestined to contribute to it, because probability and chance concepts are key to understanding of any empirical science. \\
% There are also issues, in which statistics is not thought to be able to contribute much to understanding and to be of little help for solving them. For example, the selective attention that the agents in science are paying to new, strong and clear findings in each science is considered merely a psychological and social %, socioeconomic 
% issue. It is often explained by the affinity of humans for stories rather than mere facts and for novelties. The issue that scientific findings get more attention and are more likely to be published, read and cited is known for many years, but has gained new traction in meta science when the so-called reproducibility crisis emerged in (2003 to 2005). Studies such as ... showed that even rigid prosecution of study protocols did often not lead to reproduction of previous results when experiments were repeated, which struck empirical science in its core. Most reasonably, some concluded that it was a large misunderstanding of statistical tools such as the p-value in the scientific community that lead to this situation (...). Although that some measures have been taken since then, and some progress is made, the non-reproducibility of scientific findings remain an acute threat to the relevance and reliability of science. Some even argue that certain scientific fields sooner or later will repeat the experience that for example psychology or medicine have made, and are yet to encounter their own reproducibility crisis (...). \\
% Clearly, reproduction of experimental results is the gold standard of testing empirical science, because it reassures the universality and reliability by which certain procedures or effects appear if measured in the correct experimental context. However, there are also problems there. One can almost never fully exclude chance events to have played a major role in negating or reaffirming the experimental findings (which is of course always the case where noise is present in the data). Furthermore, reproduction studies are generally costly and the precise experimental conditions might be hard to reproduce because of lack of information or other reasons. \\
% There is another way to assess the strength of scientific results if the experiment has been conducted more than once (however not with exactly identical protocol as in reproduction studies): Meta-analyses. In a meta-analyses, results of multiple (fairly identical) experiments (studies) are summarized to a single result. Very often, meta-analyses are not only regarded to be a synthesis of results, but of evidence, thus reflecting the overall and summarized evidence regarding to a scientific question. It is on purpose that selective attention to positive, strong results in the scientific literature have been mentioned beforehand, because it is clear how they will affect this second way of reassurance of scientific findings: The meta-analysis will again lead to irreproducible findings if it is based on a set of results that is itself biased. Thus, meta-analysis will in this case not reach its purpose of assessing reliability of science, but worsen the problem by reinforcing the confidence in the overestimated over-optimistic effects. However, and this is the main topic in this masters thesis, there are some ways to detect irregularities in the body of results that can provide indirect hints that a selective rather than a random sample from a hypothetical population of experiments is present. The abundance of the methods to link some features of the sample of experiments to publication bias, as the tendency of scientific literature to over-proportionally include large effects is generally called, also speaks for the relevance of the topic. A review in 2017 (...) for example identified 147 (!) conceptually differing methods.\\
% This speaks as well for the inevitable difficulties that such methods encounter. First of all, it is most often impossible to estimate the real selection process, that is the rate by which smaller effects go unnoticed because of selective publication and reporting, because the number of missing results in the studies is not known at the first place and impossible to retrieve. Secondly, there is almost no real world data of complete and unbiased meta-analyses, such that evaluations of methods is dependent on simulations. So we have, after applying the methods, only indirect information of publication bias, and the extent to what it might influences scientific findings. Almost all the time, alternative explanations for the results of the methods might relieve the operators and publishers from the reproval of publication bias. However, given the large body of evidence for publication bias collected from other sources (...), in many cases, the results from these methods indeed give a quantitative measure of publication bias.\\
% In this masters thesis, the answers for two questions are investigated: What is the extent and effect of publication bias in clinical science? For this purpose, a dataset from the Cochrane Organization is analysed with commonly used techniques to detect and adjust for publication bias (more precisely, small study effects, as it will be seen later). The discussion of the results of the analysis will result in the light of the literature to publication bias. 
%Because the extent of the analysis that is operated here is, to my knowledge, unmatched so far, this dataset

%Before beginning with the main part of this masters thesis, I want to recall some of the most pointed evidence for publication bias that has been collected for publication bias so far. \\

% Meta-analysis is at the core of evidence based medicine because it allows to summarise evidence over multiple studies and provide a more broad view on success and effectivity of clinical treatments. The necessity of meta-analyses is also increased by the abundance of data and publications. Especially when the findings differ or even contradict between studies, meta-analysis is the only way to go if one wants to make decisions based on quantitative and scientific criteria.
% 
% \vspace{0mm}
% For this, meta-analyses do not only benefit research, but also clinical practice, and may lead to better health care and prevention. However, the usefulness of meta-analysis does not restrict to clinical science, but to any empirical and quantiative science. 
% 
% \vspace{0mm}
% Usually, a meta-analysis is part of a systematic review where researchers decided to summarise all research in a given field or more specifically, that concerns a given question. Meta-analysis can be applied to all studies that are approximately identical in their experimental setup and the way the outcome of the experiments is measured. In systematic reviews where meta-analyses are used, the conclusions are most often strongly based on the results and the interpretation of the meta-analysis.
% 
% \vspace{0mm}
% However, there are problems that potentially limit the validity of meta-analysis; the number of studies available can be incomplete or the results of the studies can be biased. Some of those problems can be solved or asserted by special statistical methods. 
% 
% \subsection{Small Study Effects or Publication Bias}
% When study sample size decreases, the probability of extreme and missleading results in a study increases. This becomes a problem if results are selectively published, and therefore available, based on their results. When this is the case, one speaks of a small study effect or of ``Publication bias''. 
% 
% \vspace{0mm}
% The issue has been discussed extensively in the last years, most often in the context of what has came to be known as the replication crisis. The reasons for small study effects are manifold, but originate most often in the myopical acting of agents in science and the lack of statistical education. Studies are reported by scientists, published by journals and noticed by readers more often if their findigs are positive and find e.g. a substantive difference or effect. When doing a meta-analysis, one again obtains biased results. 
% 
% \vspace{0mm}
% The reason why that is less of an issue for larger studies is that extreme results are in general less likely and that due to larger effort, a result is published although there has been no clear and positive findings.
% 
% \vspace{0mm}
% While there is generally no way to assert poor study quality, small study effect can in principle be asserted and corrected for statistically. This masters thesis will mainly be about statistical methods to detect and adjust for small study effects. It can furthermore be divided in two parts:
% \begin{itemize}
% \item Methodological part: Collection and discussion of statistical tests and correction methods for small study effects.
% \item Applied part: Application of the methods to studies of the Cochrane Library of systematic Reviews. Subsequent discussion of the implications
% of the results for clinical science.
% \end{itemize}
% 
% In contrast to simulation studies, it is not possible to assess critical properties of the methods such as the power of a test, since the truth is not known. But based on the amount of data, one can of course try to make extrapolation to tendencies in clinical science in general. Moreover, it is still interesting to see how the methods behave in general, especially with respect to each other. It may, as an example, be possible to answer the question which statistical test is most conservative and which pooling method is most optimistic on average. Comparison with results from simulations may allow to speculate about the reasons when simulation and real world results diverge.
% 
% 
% \subsection{Cochrane and the Cochrane Database of Systematic Reviews}
% The Cochrane Organization has specialized on systematic reviews in clinical science. It publishes and maintains a library with a large number of systematic reviews that are available in some countries to the public.
% 
% \vspace{0mm}
% The data analyzed in this thesis stems completely from the Cochrane Library of systematic Reviews (cite). 
% 
% \vspace{0mm}
% The reviews are arguably of good quality, since the authors are following elaborated guidelines, and there are control-mechanisms within the organisation that should prohibit conflicts of interests. This might further improve the validity and precision of findings and conclusions that have been made based on this data. 
%