### 2019-07-27 report/chapter04.Rnw

% LaTeX file for Chapter 03

<<'preamble03',include=FALSE, echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch03_fig',
               echo=TRUE, message=FALSE,
               fig.width=8.1, fig.height=3,
               self.contained = FALSE,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

<<'Data03', echo=FALSE, warning=FALSE>>= 
#Load data:
rm(list = ls())
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
PATH2 = file.path(PATH_HOME, 'PubBias')
FILE = 'cochrane_2019-07-04.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH2, 'code')
PATH_RESULTS = file.path(PATH2, 'results_new')
# PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

source(file.path(PATH_CODE, 'PubBias_functions.R'))
require(corrgram)

load(file.path(PATH_DATA, "PubBias_2019-07-19.RData"))
load(file.path(PATH_RESULTS, "data_used_for_analysis.RData"))
load(file.path(PATH_RESULTS, "meta_analyses_summary_complete.RData"))
load(file.path(PATH_RESULTS, "meta_id_vector.RData"))
load(file.path(PATH_RESULTS, "meta_id_I2.RData"))
load(file.path(PATH_RESULTS, "meta.bin.RData"))
load(file.path(PATH_RESULTS, "meta.cont.RData"))
load(file.path(PATH_RESULTS, "meta.iv.RData"))

@

\chapter{Results} \label{ch:Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<echo=FALSE, warning=FALSE>>=
p.samplesize.z  <- data.ext2 %>% filter(total1 + total2 > 15 & total1 + total2 < 300) %>% ungroup() %>%
  mutate(std.z = abs(z), sample.size = total1 + total2) %>%
  group_by(sample.size) %>%
  summarize(median.effect = median(std.z, na.rm = T)) %>%
  ggplot(aes(x = sample.size, y = median.effect)) + geom_point(size = 0.8) +
  theme_bw() + xlab("sample size") + ylab("median absolute z-score") +
  xlim(c(15, 300)) + ylim(c(0, 0.25)) +
  scale_x_continuous(limits = c(15, 300), breaks = c(15, 100, 200, 300))

labels <- c(RR = "Risk Ratio", OR = "Odds Ratio", MD = "Mean Difference", SMD = "Std. Mean Difference")

p.samplesize.effect.sep <- data %>% filter(total1 + total2 > 15 & total1 + total2 < 300) %>%
  filter(id != "CD001293") %>%  #filter out this crazy review..
  filter(outcome.measure.merged == "RR" |
           outcome.measure.merged == "OR" |
           outcome.measure.merged == "MD" |
           outcome.measure.merged == "SMD") %>%
  group_by(outcome.measure.merged) %>%
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size, outcome.measure.merged) %>%
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>%
  ggplot(aes(x = sample.size, y = median.effect, group = outcome.measure.merged)) + geom_point(size = .5) + facet_wrap(~outcome.measure.merged, scales = "free", labeller = labeller(outcome.measure.merged = labels)) +
  theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size") +
  xlim(c(15, 300)) +
  scale_x_continuous(limits = c(15, 300), breaks = c(15, 100, 200, 300))
@

Using Fisher's $z$ transformation, we can transform all effect sizes of the dataset on the same scale such that they are comparable to each other. Thus, as a first step, an exploratory plot of the median effect size and it's dependence on the study sample size can be shown. \\
The absolute value of the median $z$-score for a given sample size of a trial is shown in Figure \ref{z.samplesize}. It is clearly visible that the absolute value decreases with increasing sample size, i.e. that the effect size is becoming smaller. The trend flattens off after $\sim$ sample size = 400 (not shown). \\
The pattern is similar in Figure \ref{effect.samplesize.separated}, where the original effect size measures ``Odds Ratio'', ``Risk Ratio'', ``Mean Difference'' and ``Std. Mean Difference'' are used (the most common measurs in the dataset). Here, the effect sizes are scaled (mean subtraction and divison through standard error) with respect to all other effects of the same measure.

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
plot(p.samplesize.z)
@
\caption{Median of the absolute $z$-score plotted against the total sample size.}
\label{z.samplesize}
\end{figure}

\begin{figure}
<<echo=FALSE>>=
plot(p.samplesize.effect.sep)
@
\caption{Median of the absolute value of the normalized, original effect size plotted against the total sample size.}
\label{effect.samplesize.separated}
\end{figure}

\section{Publication Bias Test Results} \label{sec:publication.bias.tests}
<<echo=FALSE, warning=FALSE>>=
#Agreement proportions of publication bias tests:

#Binary:
meta.bin <- meta.bin %>%ungroup() %>%  mutate(tes.d.schwarzer = ifelse(tes.d.test == schwarzer.test, "agree", "disagree"),
                                              tes.d.peter = ifelse(tes.d.test == peter.test, "agree", "disagree"),
                                              tes.d.rucker = ifelse(tes.d.test == rucker.test, "agree", "disagree"),
                                              tes.d.harbord = ifelse(tes.d.test == harbord.test, "agree", "disagree"),
                                              schwarzer.peter = ifelse(schwarzer.test == peter.test, "agree", "disagree"),
                                              schwarzer.rucker = ifelse(schwarzer.test == rucker.test, "agree", "disagree"),
                                              schwarzer.harbord = ifelse(schwarzer.test == harbord.test, "agree", "disagree"),
                                              rucker.peter = ifelse(rucker.test == peter.test, "agree", "disagree"),
                                              rucker.harbord = ifelse(rucker.test == harbord.test, "agree", "disagree"),
                                              harbord.peter = ifelse(harbord.test == peter.test, "agree", "disagree"))
agreement.bin <- meta.bin %>% ungroup() %>% summarise(tes.d.schwarzer = sum(tes.d.schwarzer == "agree")/n(),
                                                      tes.d.peter = sum(tes.d.peter == "agree")/n(),
                                                      tes.d.rucker = sum(tes.d.rucker == "agree")/n(),
                                                      tes.d.harbord = sum(tes.d.harbord == "agree")/n(),
                                                      schwarzer.peter = sum(schwarzer.peter == "agree")/n(),
                                                      schwarzer.rucker = sum(schwarzer.rucker == "agree")/n(),
                                                      schwarzer.harbord = sum(schwarzer.harbord == "agree")/n(),
                                                      rucker.peter = sum(rucker.peter == "agree")/n(),
                                                      harbord.peter = sum(harbord.peter == "agree")/n())
meta.bin <- meta.bin %>%ungroup() %>%  mutate(tes.d.schwarzer = ifelse(tes.d.test + schwarzer.test > 1, "agree", "disagree"),
                                              tes.d.peter = ifelse(tes.d.test + peter.test > 1, "agree", "disagree"),
                                              tes.d.rucker = ifelse(tes.d.test + rucker.test > 1, "agree", "disagree"),
                                              tes.d.harbord = ifelse(tes.d.test + harbord.test > 1, "agree", "disagree"),
                                              schwarzer.peter = ifelse(schwarzer.test + peter.test > 1, "agree", "disagree"),
                                              schwarzer.rucker = ifelse(schwarzer.test + rucker.test > 1, "agree", "disagree"),
                                              schwarzer.harbord = ifelse(schwarzer.test + harbord.test > 1, "agree", "disagree"),
                                              rucker.peter = ifelse(rucker.test + peter.test > 1, "agree", "disagree"),
                                              rucker.harbord = ifelse(rucker.test + harbord.test > 1, "agree", "disagree"),
                                              harbord.peter = ifelse(harbord.test + peter.test > 1, "agree", "disagree"))
agreement.sig.bin <- meta.bin %>% ungroup() %>% summarise(tes.d.schwarzer = sum(tes.d.schwarzer == "agree")/sum(tes.d.test),
                                                      tes.d.peter = sum(tes.d.peter == "agree")/sum(tes.d.test),
                                                      tes.d.rucker = sum(tes.d.rucker == "agree")/sum(tes.d.test),
                                                      tes.d.harbord = sum(tes.d.harbord == "agree")/sum(tes.d.test),
                                                      schwarzer.peter = sum(schwarzer.peter == "agree")/sum(schwarzer.test),
                                                      schwarzer.rucker = sum(schwarzer.rucker == "agree")/sum(schwarzer.test),
                                                      schwarzer.harbord = sum(schwarzer.harbord == "agree")/sum(schwarzer.test),
                                                      rucker.peter = sum(rucker.peter == "agree")/sum(rucker.test),
                                                      harbord.peter = sum(harbord.peter == "agree")/sum(peter.test))

binary.tests.agreement <- rbind(agreement.bin, agreement.sig.bin)
rownames(binary.tests.agreement) <- c("Agreement (overall)", "Agreement (significance)")
colnames(binary.tests.agreement) <- c("Excess significance, Schwarzer", "Excess significance, Peter",
                                      "Excess significance, Rucker", "Excess significance, Harbord",
                                      "Peter, Schwarzer", "Schwarzer, Rucker", "Schwarzer, Harbord",
                                      "Rucker, Peter", "Harbord, Peter")
#--------------------------------------------------------------------------------------------------------------------#

#Continuous:
meta.cont <- meta.cont %>% ungroup() %>% mutate(thompson.egger = ifelse(thompson.test == egger.test, "agree", "disagree"),
                                                thompson.begg = ifelse(thompson.test == begg.test, "agree", "disagree"),
                                                egger.begg = ifelse(egger.test == begg.test, "agree", "disagree"),
                                                tes.d.egger = ifelse(tes.d.test == egger.test, "agree", "disagree"),
                                                thompson.tes.d = ifelse(thompson.test ==tes.d.test, "agree", "disagree"),
                                                tes.d.begg = ifelse(tes.d.test == begg.test, "agree", "disagree"))
agreement.cont <- meta.cont %>% ungroup() %>%  summarise(tes.d.egger = sum(tes.d.egger == "agree")/n(),
                                                         thompson.tes.d = sum(thompson.tes.d == "agree")/n(),
                                                         tes.d.begg = sum(tes.d.begg == "agree")/n(),
                                                         thompson.egger = sum(thompson.egger == "agree")/n(),
                                                         thompson.begg = sum(thompson.begg == "agree")/n(),
                                                         egger.begg = sum(egger.begg == "agree")/n())
meta.cont <- meta.cont %>% ungroup() %>% mutate(thompson.egger = ifelse(thompson.test + egger.test > 1, "agree", "disagree"),
                                                thompson.begg = ifelse(thompson.test + begg.test > 1, "agree", "disagree"),
                                                egger.begg = ifelse(egger.test + begg.test > 1, "agree", "disagree"),
                                                tes.d.egger = ifelse(tes.d.test + egger.test > 1, "agree", "disagree"),
                                                thompson.tes.d = ifelse(thompson.test +tes.d.test > 1, "agree", "disagree"),
                                                tes.d.begg = ifelse(tes.d.test + begg.test > 1, "agree", "disagree"))
agreement.sig.cont <- meta.cont %>% ungroup() %>%  summarise(tes.d.egger = sum(tes.d.egger == "agree")/sum(tes.d.test),
                                                             thompson.tes.d = sum(thompson.tes.d == "agree")/sum(tes.d.test),
                                                             tes.d.begg = sum(tes.d.begg == "agree")/sum(tes.d.test),
                                                             thompson.egger = sum(thompson.egger == "agree")/sum(thompson.test),
                                                             thompson.begg = sum(thompson.begg == "agree")/sum(begg.test),
                                                             egger.begg = sum(egger.begg == "agree")/sum(begg.test))

cont.tests.agreement <- rbind(agreement.cont, agreement.sig.cont)
rownames(cont.tests.agreement) <- c("Agreement (overall)", "Agreement (significance)")
colnames(cont.tests.agreement) <- c( "Excess significance, Egger", "Excess significance, Thompson",
                                     "Excess significance, Begg","Thompson, Egger", "Thompson, Begg", "Egger, Begg")

#--------------------------------------------------------------------------------------------------------------------#

#IV:
meta.iv <- meta.iv %>% ungroup() %>% mutate(thompson.egger = ifelse(thompson.test == egger.test, "agree", "disagree"),
                                                thompson.begg = ifelse(thompson.test == begg.test, "agree", "disagree"),
                                                egger.begg = ifelse(egger.test == begg.test, "agree", "disagree"),
                                                tes.d.egger = ifelse(tes.d.test == egger.test, "agree", "disagree"),
                                                thompson.tes.d = ifelse(thompson.test ==tes.d.test, "agree", "disagree"),
                                                tes.d.begg = ifelse(tes.d.test == begg.test, "agree", "disagree"))
agreement.iv <- meta.iv %>% ungroup() %>%  summarise(tes.d.egger = sum(tes.d.egger == "agree")/n(),
                                                         thompson.tes.d = sum(thompson.tes.d == "agree")/n(),
                                                         tes.d.begg = sum(tes.d.begg == "agree")/n(),
                                                         thompson.egger = sum(thompson.egger == "agree")/n(),
                                                         thompson.begg = sum(thompson.begg == "agree")/n(),
                                                         egger.begg = sum(egger.begg == "agree")/n())
meta.iv <- meta.iv %>% ungroup() %>% mutate(thompson.egger = ifelse(thompson.test + egger.test > 1, "agree", "disagree"),
                                                thompson.begg = ifelse(thompson.test + begg.test > 1, "agree", "disagree"),
                                                egger.begg = ifelse(egger.test + begg.test > 1, "agree", "disagree"),
                                                tes.d.egger = ifelse(tes.d.test + egger.test > 1, "agree", "disagree"),
                                                thompson.tes.d = ifelse(thompson.test +tes.d.test > 1, "agree", "disagree"),
                                                tes.d.begg = ifelse(tes.d.test + begg.test > 1, "agree", "disagree"))
agreement.sig.iv <- meta.iv %>% ungroup() %>%  summarise(tes.d.egger = sum(tes.d.egger == "agree")/sum(egger.test),
                                                             thompson.tes.d = sum(thompson.tes.d == "agree")/sum(thompson.test),
                                                             tes.d.begg = sum(tes.d.begg == "agree")/sum(begg.test),
                                                             thompson.egger = sum(thompson.egger == "agree")/sum(thompson.test),
                                                             thompson.begg = sum(thompson.begg == "agree")/sum(begg.test),
                                                             egger.begg = sum(egger.begg == "agree")/sum(begg.test))

iv.tests.agreement <- rbind(agreement.iv, agreement.sig.iv)
rownames(iv.tests.agreement) <- c("Agreement (overall)","Agreement (significance)")
colnames(iv.tests.agreement) <- c("Excess significance, Egger (IV)", "Excess significance, Thompson (IV)",
                                    "Excess significance, Begg (IV)", "Thompson, Egger (IV)",
                                    "Thompson, Begg (IV)", "Egger, Begg (IV)")

#--------------------------------------------------------------------------------------------------------------------#

#Merging:
test.agreement <- rbind(t(binary.tests.agreement), t(cont.tests.agreement), t(iv.tests.agreement))
#--------------------------------------------------------------------------------------------------------------------#

sig.level <- 0.1
meta.f <- meta.f %>% rowwise() %>% 
  mutate(i2f = factor(ifelse(I2 == 0, 1, 0)))

meta.bin <- meta.f %>% filter(outcome.flag == "DICH")
meta.cont <- meta.f %>% filter(outcome.flag == "CONT")
meta.iv <- meta.f %>% filter(outcome.flag == "IV")

dat_text <- data.frame(
  label = paste(c(sum(meta.bin$harbord.test), sum(meta.bin$peter.test), 
                  sum(meta.bin$rucker.test), sum(meta.bin$rucker.test.linreg), 
                  sum(meta.bin$schwarzer.test), sum(meta.bin$tes.d.test, na.rm = T)), "< 0.1,",
                round(c(mean(meta.bin$harbord.test), mean(meta.bin$peter.test), 
                        mean(meta.bin$rucker.test), mean(meta.bin$rucker.test.linreg), 
                        mean(meta.bin$schwarzer.test), mean(meta.bin$tes.d.test, na.rm = T)),2)*100, "%"),
  test.type   = c("pval1.harbord", "pval1.peter", "pval1.rucker", "pval1.rucker.linreg", "pval1.schwarzer", "pval1.d.tes"),
  i2f = factor(rep(0, times = 6)))

labels <- c(pval1.harbord = "Harbord", pval1.peter = "Peter", pval1.rucker = "Rucker (thompson)", pval1.rucker.linreg = "Rucker (egger)", pval1.schwarzer = "Schwarzer", pval1.d.tes = "Excess Significance")

p.dist.bin <- meta.bin %>% ungroup() %>% 
  select(i2f, pval1.harbord, pval1.peter, pval1.rucker, pval1.rucker.linreg, pval1.d.tes, pval1.schwarzer) %>% 
  gather(key = "test.type", value = "p.value", pval1.harbord:pval1.schwarzer) %>% 
  ggplot(aes(x = p.value, fill = i2f)) + geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
  facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
  geom_text(data = dat_text, mapping = aes(x = 0.5, y = 125, label = label), color = "black") + 
  theme(strip.text.x = element_text(size=7)) + ggtitle("Binary Outcomes") + 
  theme(legend.position = "none")

#--------------------------------------------------------------------------------------------------------------------#

dat_text <- data.frame(
  label = paste(c(sum(meta.cont$begg.test), sum(meta.cont$egger.test), sum(meta.cont$thompson.test), sum(meta.cont$tes.d.test, na.rm = T)), "< 0.1,",
                round(c(mean(meta.cont$begg.test), mean(meta.cont$egger.test), mean(meta.cont$thompson.test), mean(meta.cont$tes.d.test, na.rm = T)),2)*100, "%"),
  test.type   = c("pval1.begg", "pval1.egger", "pval1.thompson", "pval1.d.tes"),
  i2f = factor(rep(0, times = 4)))

labels <- c(pval1.begg = "Begg Mazumdar", pval1.egger = "Egger", pval1.thompson = "Thompson Sharp", pval1.d.tes = "Excess significance")
p.dist.cont <- meta.cont %>% ungroup() %>% 
  select(i2f, pval1.egger, pval1.thompson, pval1.begg, pval1.d.tes) %>% 
  gather(key = "test.type", value = "p.value", pval1.egger:pval1.d.tes) %>% 
  ggplot(aes(x = p.value, fill = i2f)) + geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
  facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
  geom_text(data = dat_text, mapping = aes(x = 0.5, y = 60, label = label),  color = "black") + 
  theme(strip.text.x = element_text(size=7)) + ggtitle("Continuous Outcomes") + 
  theme(legend.position = "none")

#--------------------------------------------------------------------------------------------------------------------#

dat_text <- data.frame(
  label = paste(c(sum(meta.iv$begg.test), sum(meta.iv$egger.test), sum(meta.iv$thompson.test), sum(meta.iv$tes.d.test, na.rm = T)), "< 0.1,",
                round(c(mean(meta.iv$begg.test), mean(meta.iv$egger.test), mean(meta.iv$thompson.test), mean(meta.iv$tes.d.test, na.rm = T)),2)*100, "%"),
  test.type   = c("pval1.begg", "pval1.egger", "pval1.thompson", "pval1.d.tes"),
  i2f = factor(rep(0, times = 4)))

p.dist.iv <- meta.iv %>% ungroup() %>% 
  select(i2f, pval1.egger, pval1.thompson, pval1.begg, pval1.d.tes) %>% 
  gather(key = "test.type", value = "p.value", pval1.egger:pval1.d.tes) %>% 
  ggplot(aes(x = p.value, fill = i2f)) + geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
  facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
  guides(fill=guide_legend(title=expression(paste(I^2, " Heterogeneity")))) + 
  geom_text(data = dat_text, mapping = aes(x = 0.5, y = 25, label = label),  color = "black")  + 
  scale_fill_discrete(labels = c("> 0", "= 0")) +
  theme(strip.text.x = element_text(size=7)) + 
  ggtitle("IV Outcomes") 
#--------------------------------------------------------------------------------------------------------------------#

################################################################################################
################################################################################################
#PUBLICATION BIAS TEST AGREEMENT
################################################################################################
################################################################################################

cor.data <- meta.f %>% select(stat.rucker, stat.harbord, stat.peter, stat.schwarzer, 
                             stat.egger, stat.thompson, stat.begg, I2)
colnames(cor.data) <- c("Rucker", "Harbord", "Peter", "Schwarzer", "Egger", "Thompson",
                        "Begg", "I-squared")

panel.pts.new <- function (x, y, corr = NULL, col.regions, cor.method, ...) 
{

  plot.xy(xy.coords(x, y), type = "p", ...)
  box(col = "lightgray")
  rect(xleft = -1.644854, ybottom = -1.644854, xright = 1.644854, ytop = 1.644854, density = NULL, angle = 45,
       col = NA, border = "white", lty = 1, lwd = 1.4, 
       ...)
  if (!is.null(corr)) 
    return()

}

# Tukey Mean difference plots;
#--------------------------------------------------------------------------------------------------------------------#
logit <- function(x) log(x/(1-x))

egger.thompson.plot <- meta.f %>% 
  mutate(mean = ((logit(pval1.egger)+ logit(pval1.thompson)))/2, 
         difference = logit(pval1.egger) - logit(pval1.thompson)) %>% 
  filter(difference < 10) %>% 
  ggplot(aes(x = mean, y = difference)) + geom_point(size = .5) + 
  geom_abline(aes(slope = 0, intercept = mean(difference, na.rm = T), col = "blue")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) + 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) - 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  # geom_smooth(method = "loess", se = F) + 
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(labels = c("systematic error", "95 % confidence intervals"), 
                     values = c("red", "blue")) +
  ggtitle(expression(paste("logit(", italic(p),"-value Egger) - logit(", italic(p), "-value Thompson)" ))) + geom_smooth(method = "lm", col = "green")
#--------------------------------------------------------------------------------------------------------------------#

egger.excess.plot <- meta.f %>% 
  mutate(mean = ((logit(pval1.egger)+ logit(pval1.d.tes)))/2, 
         difference = logit(pval1.egger) - logit(pval1.d.tes)) %>% 
  filter(difference > -Inf & difference < 10) %>% 
  ggplot(aes(x = mean, y = difference)) + geom_point(size = .5) + 
  geom_abline(aes(slope = 0, intercept = mean(difference, na.rm = T), col = "blue")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) + 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) - 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  # geom_smooth(method = "loess", se = F) + 
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(labels = c("systematic error", "95 % confidence intervals"), 
                     values = c("red", "blue")) +
  ggtitle(expression(paste("logit(", italic(p),"-value Egger) - logit(", italic(p), "-value excess significance)" ))) + geom_smooth(method = "lm", col = "green")
#--------------------------------------------------------------------------------------------------------------------#

harbord.rucker.plot <- meta.f %>% 
  mutate(mean = ((logit(pval1.harbord)+ logit(pval1.rucker)))/2, 
         difference = logit(pval1.harbord) - logit(pval1.rucker)) %>% 
  ggplot(aes(x = mean, y = difference)) + geom_point(size = .4) + 
  geom_abline(aes(slope = 0, intercept = mean(difference, na.rm = T), col = "blue")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) + 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) - 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  # geom_smooth(method = "loess", se = F) + 
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(labels = c("systematic error", "95 % confidence intervals"), 
                     values = c("red", "blue")) +
  ggtitle(expression(paste("logit(", italic(p),"-value Harbord) - logit(", italic(p), "-value Rucker)" ))) + geom_smooth(method = "lm", col = "green")
#--------------------------------------------------------------------------------------------------------------------#

harbord.excess.plot <- meta.f %>% 
  mutate(mean = ((logit(pval1.harbord)+ logit(pval1.d.tes)))/2, 
         difference = logit(pval1.harbord) - logit(pval1.d.tes)) %>% 
  ggplot(aes(x = mean, y = difference)) + geom_point(size = .4) + 
  geom_abline(aes(slope = 0, intercept = mean(difference, na.rm = T), col = "blue")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) + 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) - 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  # geom_smooth(method = "loess", se = F) + 
  theme_bw() + #theme(legend.position = "bottom") +
  scale_color_manual(labels = c("systematic error", "95 % conf. int."), 
                     values = c("red", "blue")) +
  ggtitle(expression(paste("logit(", italic(p),"-value Harbord) - logit(", italic(p), "-value excess significance)" ))) + geom_smooth(method = "lm", col = "green")
#--------------------------------------------------------------------------------------------------------------------#

#Table with proportions of significant test results:
meta.f <- meta.f %>% 
mutate(test.sum.bin = tes.d.test + schwarzer.test + harbord.test + peter.test + rucker.test,
       test.sum.else = tes.d.test + egger.test + thompson.test + begg.test, 
       test.sum.bin2 = tes.d.test + harbord.test,
       test.sum.else2 = tes.d.test + egger.test, 
       test.min.bin = ifelse(schwarzer.test + harbord.test + peter.test + rucker.test > 0, 1, 0),
       test.min.else = ifelse(egger.test + thompson.test + begg.test > 0, 1, 0),
       test.min.bin2 = ifelse(test.sum.bin2 > 0, 1, 0),
       test.min.else2 = ifelse(test.sum.else2 > 0, 1, 0))

mean.min.bin <- round(mean(meta.f$test.min.bin, na.rm = T)*100, 1)
mean.min.else <- round(mean(meta.f$test.min.else, na.rm = T)*100, 1)

mean.min.bin2 <- round(mean(meta.f$test.min.bin2, na.rm = T)*100, 1)
mean.min.else2 <- round(mean(meta.f$test.min.else2, na.rm = T)*100, 1)

both.bin <- round(length(which(meta.f$test.sum.bin2 == 2))/
                    length(which(!is.na(meta.f$test.sum.bin2)))*100, 1)

both.else <- round(length(which(meta.f$test.sum.else2 == 2))/
                     length(which(!is.na(meta.f$test.sum.else2)))*100, 1)

summary.bin <- round(meta.f %>% 
  group_by(test.sum.bin) %>% summarise(n = n()) %>% select(n)/dim(meta.bin)[1] * 100, 1)

summary.else <- round(meta.f %>% group_by(test.sum.else) %>% 
                        summarise(n = n()) %>% 
                        select(n)/(dim(meta.cont)[1] + dim(meta.iv)[1]) * 100, 1)

number.sig.tests <- data.frame(number = seq(from = 0, to = 5),
                               binary = paste(summary.bin$n[-7], "%"),
                               cont = as.character(paste(c(summary.else$n[-c(6,7)], NA), "%")))
number.sig.tests <- as.matrix(number.sig.tests)
number.sig.tests[6,3][1] <- "-"

colnames(number.sig.tests) <- c("Count", "Binary Outcomes", "Continuous and IV")

meta.f <- meta.f %>% mutate(tes.d.test.b = ifelse(pval1.d.tes < 0.05, 1, 0),
                            harbord.test.b = ifelse(pval1.harbord < 0.05, 1, 0),
                            egger.test.b = ifelse(pval1.egger < 0.05, 1, 0),
                            test.sum.bin2 = tes.d.test.b + harbord.test.b,
                            test.sum.else2 = tes.d.test.b + egger.test.b, 
                            test.min.bin2 = ifelse(test.sum.bin2 > 0, 1, 0),
                            test.min.else2 = ifelse(test.sum.else2 > 0, 1, 0))

mean.min.bin2.b <- round(mean(meta.f$test.min.bin2, na.rm = T)*100, 1)
mean.min.else2.b <- round(mean(meta.f$test.min.else2, na.rm = T)*100, 1)

both.bin.b <- round(length(which(meta.f$test.sum.bin2 == 2))/
                    length(which(!is.na(meta.f$test.sum.bin2)))*100, 1)

both.else.b <- round(length(which(meta.f$test.sum.else2 == 2))/
                     length(which(!is.na(meta.f$test.sum.else2)))*100, 1)

meta.f <- meta.f %>% rowwise() %>% mutate(egger.test = ifelse(pval1.egger < 0.1/3, 1, 0),
                            thompson.test = ifelse(pval1.thompson < 0.1/3, 1, 0),
                            begg.test = ifelse(pval1.begg < 0.1/3, 1, 0),

                            schwarzer.test = ifelse(pval1.schwarzer < 0.1/4, 1, 0),
                            rucker.test = ifelse(pval1.rucker < 0.1/4, 1, 0),
                            harbord.test = ifelse(pval1.harbord < 0.1/4, 1, 0),
                            peter.test = ifelse(pval1.peter < 0.1/4, 1, 0),
                            test.sum.bin = tes.d.test + schwarzer.test + harbord.test + peter.test + rucker.test,
                            test.sum.else = tes.d.test + egger.test + thompson.test + begg.test, 
                            test.min.bin = ifelse(schwarzer.test + harbord.test + peter.test + rucker.test > 0, 1, 0),
                            test.min.else = ifelse(egger.test + thompson.test + begg.test > 0, 1, 0))

mean.min.bin.b <- round(mean(meta.f$test.min.bin, na.rm = T)*100, 1)
mean.min.else.b <- round(mean(meta.f$test.min.else, na.rm = T)*100, 1)

#--------------------------------------------------------------------------------------------------------------------#
#Calculate harmonic mean of p-values of pb tests: 
meta.f <- meta.f %>% mutate(hmean = case_when(
  outcome.flag == "DICH" ~ 5/sum(pval1.rucker^-1 + pval1.harbord^-1 + pval1.schwarzer^-1 + pval1.peter^-1 + pval1.d.tes^-1),
  outcome.flag != "DICH" ~ 4/sum(pval1.egger^-1 + pval1.thompson^-1 + pval1.begg^-1 + pval1.d.tes^-1)))

harmonic.means <- c(length(which(meta.f$hmean < 0.1))/dim(meta.f)[1], 
length(which(meta.f$hmean[which(meta.f$outcome.flag == "DICH")] < 0.1))/dim(meta.f[which(meta.f$outcome.flag == "DICH"),])[1],
length(which(meta.f$hmean[which(meta.f$outcome.flag == "CONT")] < 0.1))/dim(meta.f[which(meta.f$outcome.flag == "CONT"),])[1],
length(which(meta.f$hmean[which(meta.f$outcome.flag == "IV")] < 0.1))/dim(meta.f[which(meta.f$outcome.flag == "IV"),])[1])
harmonic.means <- round(harmonic.means*100, 1)

#--------------------------------------------------------------------------------------------------------------------#
#--------------------------------------------------------------------------------------------------------------------#
#--------------------------------------------------------------------------------------------------------------------#
#--------------------------------------------------------------------------------------------------------------------#
#--------------------------------------------------------------------------------------------------------------------#
#--------------------------------------------------------------------------------------------------------------------#
#--------------------------------------------------------------------------------------------------------------------#

@

The meta-analyses fulfilling the criteria from chapter \ref{ch:dataset}, section \ref{sec:Processing}, are analysed with one-sided small publication bias tests and excess significance tests. The direction in which bias is expected is the one on which more significant results are ($p$-value < 0.05, two-sided). The tests are applied on the original effect size measures, since the journal editors and the researchers also base their decisions on them. Different tests are applied depending on the outcome being binary or continuous.\\
Multiple tests are applied on the same data to compare their results. A histogram of $p$-values for each test and all meta-analyses will summarize the overall evidence against the null-hypothesis of no publication bias, as displayed in Figure \ref{fig:test}. Skewedness indicates evidence for publication bias.\\
The abbreviations in Figure \ref{fig:test}  are shortly explained with references to chapter \ref{ch:methods}: \\
``Excess significance'' denotes the excess of significant $p$-values testing method from \citet{excess.significance}, see \ref{sec:excess.significance}. For continuous and \texttt{IV} outcomes, the names refer to:

\begin{itemize}
\item Egger's test, weighted linear regression test described in section \ref{sec:Egger}
\item Thompson and Sharp's test, weighted linear regression test adjusted for between-study heterogeneity, section \ref{sec:Thompson}
\item Begg and Mazumdar's test, rank test described in section \ref{sec:Begg}
\end{itemize}

For binary outcomes, the names refer to:
\begin{itemize}
\item Harbord's test, likelihood score based test (section \ref{sec:Harbord})
\item Peter's test, weighted linear regression with inverse sample size as explanatory variable described in section \ref{sec:Peter}
\item R\"ucker's test, test based on the arcsine transformation of proportions, in combination with Thompson and Sharp's regression test (section \ref{sec:Rucker})
\item Schwarzer's test, rank based test using the expected event counts computed with the hypergeometric distribution (section \ref{sec:Schwarzer})
\end{itemize}

\begin{figure}
<<echo=FALSE, fig.height = 9>>=
grid.arrange(p.dist.bin,
             p.dist.cont,
             p.dist.iv, ncol = 1)
@
\caption{Histogram of one-sided $p$-values for small study effect in direction of larger effect sizes. The testing method is indicated in the header, bin width is equal to 0.1. The proportion of meta-analyses with signifiant publication bias based on the threshold of 0.1 is displayed inside the figures.}
\label{fig:test}
\end{figure}

%Agreement table -> Chunck 1
The histograms in Figure \ref{fig:test} show that the tests are not alway in agreement if publication bias is present in the data. The $p$-values of excess significance and Schwarzer's test and the $p$-values for \texttt{IV} outcomes are rather uniformly distributed, or even skewed to the right. Notably, excess significance test, Schwarzer's test and rank tests have been shown to lack power. The tests that are more suitable (regression based tests in general) all have proportions of significant $p$-values ($p$ > 0.1) clearly above 10 \% which would be the expected false positive rate. The conclusion is that there is publication bias, however the extent can be debated. The meta-analyses with \texttt{outcome.flag} = \texttt{IV} seem to differ from the other meta-analyses. \\
In Figure \ref{fig:test}, the meta-analyses with an estimated $I^2$ of zero are depicted, because some methods are known to only be suitable when no heterogeneity is present (excess significance test and also Egger's test). Others are specially constructed to adjust for between study heterogeneity (Thompson and Sharp's test and R\"ucker's test). In continuous outcomes, when Thompson and Sharp's test is applied, the evidence for publication bias decreases, as can be seen in the smaller proportion of blue on the left corner in Figure \ref{fig:test}. However, this is also due to application to meta-analyses with $I^2$ equal zero, in which case the method is known to lack power. Similarly, the evidence decreases somewhat when R\"ucker's test is extended by Thompson and Sharp's method to account for heterogeneity. The moderate decrease indicates however that the previous restriction to meta-analyses with $I^2 < 0.5$ is sufficient to remove meta-analyses with large heterogeneity and that the test results are not heavily influenced by unaccounted between study heterogeneity. \\
The $p$-values of tests can be summarized by computing their harmonic mean. In the case of binary tests, the $p$-values of R\"ucker's, Peters, Harbord's, Schwarzer's and excess significance tests are used, in the case of continuous and \texttt{outcome.flag} = \texttt{IV} outcomes, Egger's, Thompson and Sharp's, Begg and Mazumdar's and excess significance tests are used. This lead's to an overall of \Sexpr{harmonic.means[1]}\% significant results ($p_\textrm{harmonic}$ < 0.1, \Sexpr{harmonic.means[2]}\% \texttt{outcome.flag} = \texttt{DICH}, \Sexpr{harmonic.means[3]}\% \texttt{outcome.flag} = \texttt{CONT}, \Sexpr{harmonic.means[4]}\% \texttt{outcome.flag} = \texttt{IV}).

\subsection{Publication Bias Test Consistency}
It cannot be seen in Figure \ref{fig:test} if the tests used are finding small study effects for the same meta-analyses. A simple method to check the consistency of test results is to compare scatterplots and empirical Spearman correlations between the test statistics. This is done in Figure \ref{fig:test.agreement}. Here, there is no separation between \texttt{IV} and continuous outcomes. The upper left rectangle is displaying binary outcome results and the lower right continuous and \texttt{IV} outcomes results. Also, the $I^2$ statistic is included. Since no approximately normally distributed test statistic is used for excess significance tests, it is not shown here

\begin{figure}
<<echo=FALSE, fig.height = 7, fig.width = 8.5, warning=FALSE>>=
corrgram(cor.data, 
         upper.panel=panel.pts.new, lower.panel=panel.cor, 
         cex.labels = 1, cex = .25, cex.cor = 2, pch = 1,
         text.panel=panel.txt, cor.method = "spearman")
@
\caption{Pairs-plot for test statistics of small study effect and excess significance. The lower panel gives the Spearman correlations for the different test statistics, and the upper panel displays a scatterplot. The colors indicate magnitude and direction of the correlation coefficients. The rectangle with white borders displays the area within which both tests have absolute value < 1.64 (dots inside are statistically not significant by 0.1 $p$-value threshold).}
\label{fig:test.agreement}
\end{figure}

The observed patterns on the scatterplots differ, and some small study effect test statistics do align better than others. Regression based tests as Egger and Thompsons test which are methodically almost identical are closely aligned, which is reflected in large correlation coefficients. Continuous and\texttt{IV} outcome type tests align more closely than binary outcome tests. While correlation coefficients between binary outcome tests vary between each other, Harbord's test statistic has similar correlation coefficients with the other small study effect test statistics.\\
Because scatterplots and correlation coefficients can be misleading, also a Tukey mean-difference or Bland-Altman of transformed $p$-values plot is shown for four scenarios in Figure \ref{fig:mean.diff.test}:
\begin{itemize}
\item For Egger's and Thompson's tests, which is supposedly the most similar test and should show the least deviations and systematic errors
\item For Egger's and excess significance tests
\item For Harbord's and R\"ucker's tests
\item For Harbord's and excess significance tests
\end{itemize}

This can be justified since all tests are supposed to measure the evidence for publication bias. For the plots, the $p$-values of the tests are transformed on the entire continuous scale by a logit transformation $f(x)  = \log(\frac{p}{1-p}$). The mean $p$-value (($f$($p$-value no. 1) + $f$($p$-value no. 2))/2) is then displayed against the difference between the $f$($p$-value). If no systematic errors and biases exist between the measurement methods, then

\begin{itemize}
\item the mean of the differences should be around zero (no systematic error) 
\item the points should scatter independently on the $y$-axis and no general increase or decrease with the mean of the transformed $p$-values should be visible (and the linear regression fit is flat)
\end{itemize}

There are likely systematic errors and bias between the tests, although the extent seems to vary. Most error seems to be between small study effect tests and excess significance tests, because the slope of the linear regression fit is likely positive. This means that the excess significance test finds less evidence in cases when both $p$-values are small and more evidence when both $p$-values are large, on average.\\
The confidence intervals from Figure \ref{fig:mean.diff.test} can be used as limits of agreement, which gives, after back-transformation, around 0.9 for Egger's and Thompson's test, and around 0.99 for Harbord's and R\"ucker's tests. This suggests that additionally to the bias correspondence between the tests is not very good in general.

\begin{figure}
<<echo=FALSE, fig.height = 8.5, warning=FALSE>>=
grid.arrange(egger.thompson.plot, 
             egger.excess.plot,
             harbord.rucker.plot, 
             harbord.excess.plot, ncol = 1)
@
\caption{Mean - difference plots for logit transformed $p$-values. The mean of logit transformed $p$-values is displayed on the $x$-axis and the difference on the $y$-axis. Blue and red lines display the systematic error and the confidence intervals of the systematic error (limits of agreement).
In green, a linear regression fit is shown with 95\% CI bands.}
\label{fig:mean.diff.test}
\end{figure}

The previous results suggest that the results will also differ substantially after applying the common dichotomization of $p$-values. Some proportion of the meta-analyses will only be significant using a single test, while being non-significant otherwise. This can be seen in Table \ref{number.sig.tests}. It displays the percentage of meta-analyses with a certain number of significant test results. For both outcome types, there is only marginal agreement, very few cases are significant in any test. Around 67\% of the dataset is not significant, no matter which test is used. When only using small study effect tests, \Sexpr{mean.min.bin}\% of binary outcome tests and \Sexpr{mean.min.else}\% of \texttt{IV} and continuous outcome tests had at least one significant result. After applying the Bonferroni correction for multiple testing, this shrinks to \Sexpr{mean.min.bin.b}\% for binary and \Sexpr{mean.min.else.b}\% for continuous and \texttt{IV} outcomes. To compare significant findings for small study effect tests and excess significance tests, Harbord's or Egger's test results are compared with excess significane tests. \Sexpr{mean.min.bin2}\% of binary outcome analyses had at least one of the two test $p$-values being significant, and equivalently, \Sexpr{mean.min.else2}\% for continuous and \texttt{IV} outcomes. The numbers change to \Sexpr{mean.min.bin2.b}\% for binary outcomes and \Sexpr{mean.min.else2.b}\% for continuous and \texttt{IV} outcomes after applying the Bonferroni correction. \Sexpr{both.bin}\% have significant Harbord's test result and significant excess significance test result  (\Sexpr{both.bin.b}\% with Bonferroni). Of the continuous outcomes, we have \Sexpr{both.else}\% with Egger's test and \Sexpr{both.else.b}\% with Bonferroni correction. \\

<<echo = FALSE, results='asis'>>=
print(xtable(number.sig.tests, caption = "Counts Number of significant test results per meta-analysis, separated
       for outcome types. Last entry for continuous and IV outcomes is empty since one test less was 
       applied", label = "number.sig.tests", align = "lccc"), include.rownames = F, size = "footnotesize")
@

The precise proportions of agreement are provided in table \ref{test.agreement}. We see that only in few cases when excess significance test indicates statistical significance, small study effect tests do so. Correspondence between excess significance tests and small study effects thus has to be considered rather random. Linear regression based tests agree more often with other linear regression based tests, and agreement between small study effect tests is in general well above 60 \% (at best, 95 \% test agreement in significance for \texttt{IV} \texttt{outcome.flag}).

<<echo = FALSE, results = 'asis'>>=
print(xtable(test.agreement, align = "lcc", caption = "Overall proportion of agreement if significant or unsignificant, and for significance only. Horizontal lines separate binary, continuous and IV outcomes (order as in table). The reference significance test is the one with more significant results.",
             label = "test.agreement"),, size = "scriptsize")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Small Study Effects Adjustment}
<<echo=FALSE, warning=FALSE, message=FALSE>>=

#HISTOGRAMS:

#Comparison of treatment effect p-values:
sig.zcor <- meta.f %>% mutate(z.fixef = est.z.fixef/se.est.z.fixef,
                              z.ranef = est.z.ranef/se.est.z.ranef,
                              z.reg = est.z.reg/se.est.z.reg,
                              z.copas = est.z.copas/se.est.z.copas) %>%  
  select(z.fixef, z.ranef, z.reg, z.copas) %>% gather(key = "method", value = "fisher.z") %>% 
  mutate(p.fisher.z = 2*(1-pnorm(abs(fisher.z)))) %>% 
  group_by(method) %>% 
  summarise(significant = length(which(abs(fisher.z) > 1.96)),
            p.significant = significant/length(fisher.z))
sig.zcor <- data.frame(method = sig.zcor$method,
                       label = paste(round(sig.zcor$p.significant,3)*100, "% significant", ", (n = ", sig.zcor$significant, ")", sep = ""))

method_names <- c(
  z.fixef = "Fixed Effects",
  z.ranef = "Random Effects",
  z.reg = "Regression",
  z.copas = "Copas"
)

adjustment.p.z <- meta.f %>% 
  mutate(z.fixef = est.z.fixef/se.est.z.fixef,
         z.ranef = est.z.ranef/se.est.z.ranef,
         z.reg = est.z.reg/se.est.z.reg,
         z.copas = est.z.copas/se.est.z.copas) %>%  
  select(z.fixef, z.ranef, z.reg, z.copas) %>% 
  gather(key = "method", value = "fisher.z") %>% 
  mutate(method = factor(method, levels = c("z.fixef", "z.ranef", "z.reg", "z.copas"))) %>% 
  mutate(p.fisher.z = 2*(1-pnorm(abs(fisher.z)))) %>% 
  ggplot(aes(x = p.fisher.z)) + 
  geom_histogram(boundary = 0, bins = 20) + 
  theme_bw() + 
  facet_wrap(~method, ncol = 2, labeller = as_labeller(method_names)) +
  geom_text(data = sig.zcor, aes(x = 0.5, y = 750, label = label), position = "dodge") + 
  xlab(expression(paste(italic(p),"-value of test statistic of the ", italic(z),"-score"))) +
  ggtitle(expression(paste(italic(z), " Score ", italic(p),"-value of Meta-Analysis and Adjustment Method")))
#--------------------------------------------------------------------------------------------------------------------#

#Comparison of SMD's:
sig.d <- meta.f %>% mutate(d.fixef = est.d.fixef/se.est.d.fixef,
                           d.ranef = est.d.ranef/se.est.d.ranef,
                           d.reg = est.d.reg/se.est.d.reg,
                           d.copas = est.d.copas/se.est.d.copas) %>%  
  select(d.fixef, d.ranef, d.reg, d.copas) %>% gather(key = "method", value = "smd") %>% 
  mutate(p.smd = 2*(1-pnorm(abs(smd)))) %>% 
  group_by(method) %>% 
  summarise(significant = length(which(abs(smd) > 1.96)),
            p.significant = significant/length(smd))
sig.d <- data.frame(method = sig.d$method,
                    label = paste(round(sig.d$p.significant,3)*100, "% significant", ", (n = ", sig.d$significant, ")", sep = ""))

method_names <- c(
  d.fixef = "Fixed Effects",
  d.ranef = "Random Effects",
  d.reg = "Regression",
  d.copas = "Copas"
)

adjustment.p.d <- meta.f %>% 
  mutate(d.fixef = est.d.fixef/se.est.d.fixef,
         d.ranef = est.d.ranef/se.est.d.ranef,
         d.reg = est.d.reg/se.est.d.reg,
         d.copas = est.d.copas/se.est.d.copas) %>%  
  select(d.fixef, d.ranef, d.reg, d.copas) %>% 
  gather(key = "method", value = "smd") %>% 
  mutate(method = factor(method, levels = c("d.fixef", "d.ranef", "d.reg", "d.copas"))) %>% 
  mutate(p.smd = 2*(1-pnorm(abs(smd)))) %>% 
  ggplot(aes(x = p.smd)) + 
  geom_histogram(boundary = 0, bins = 20) + 
  theme_bw() + 
  facet_wrap(~method, ncol = 2, labeller = as_labeller(method_names)) +
  geom_text(data = sig.d, aes(x = 0.5, y = 750, label = label), position = "dodge") + 
  xlab(expression(paste(italic(p),"-value of test statistic of std. mean differences"))) +
  ggtitle(expression(paste("std. mean difference ", italic(p),"-value of Meta-Analysis and Adjustment Method")))
#--------------------------------------------------------------------------------------------------------------------#

#Comparison of log hazard ratios:
sig.d <- meta.f %>% filter(outcome.flag == "IV") %>% 
  mutate(d.fixef = est.fixef/se.est.fixef,
         d.ranef = est.ranef/se.est.ranef,
         d.reg = est.reg/se.est.reg,
         d.copas = est.copas/se.est.copas) %>%  
  select(d.fixef, d.ranef, d.reg, d.copas) %>% gather(key = "method", value = "smd") %>% 
  group_by(method) %>% 
  summarise(significant = length(which(abs(smd) > 1.96)),
            p.significant = (significant + sum(is.na(smd)))/(length(smd)))
sig.d <- data.frame(method = sig.d$method,
                    label = paste(round(sig.d$p.significant,3)*100, "% significant", ", (n = ", sig.d$significant, ")", sep = ""))

method_names <- c(
  d.fixef = "Fixed Effects",
  d.ranef = "Random Effects",
  d.reg = "Regression",
  d.copas = "Copas"
)

adjustment.p.log.hazard.ratio <- meta.f %>% filter(outcome.flag == "IV") %>% 
  mutate(d.fixef = est.fixef/se.est.fixef,
         d.ranef = est.ranef/se.est.ranef,
         d.reg = est.reg/se.est.reg,
         d.copas = est.copas/se.est.copas) %>%  
  select(d.fixef, d.ranef, d.reg, d.copas) %>% gather(key = "method", value = "smd") %>% 
  mutate(method = factor(method, levels = c("d.fixef", "d.ranef", "d.reg", "d.copas"))) %>% 
  mutate(p.smd = 2*(1-pnorm(abs(smd)))) %>% 
  ggplot(aes(x = p.smd)) + 
  geom_histogram(boundary = 0, bins = 20) + 
  theme_bw() + 
  facet_wrap(~method, ncol = 2, labeller = as_labeller(method_names)) +
  geom_text(data = sig.d, aes(x = 0.5, y = 100, label = label), position = "dodge") + 
  xlab(expression(paste(italic(p),"-value of test statistic of Hedges"))) +
  ggtitle(expression(paste("Log IV outcome measures ", italic(p),"-value of Meta-Analysis and Adjustment Method")))
#--------------------------------------------------------------------------------------------------------------------#

#Plots separated by effect measure:

sig.level <- 0.1
meta.f <- meta.f %>% rowwise() %>% 
  mutate(egger.test = ifelse(pval1.egger < sig.level, 1, 0),
         thompson.test = ifelse(pval1.thompson < sig.level, 1, 0),
         begg.test = ifelse(pval1.begg < sig.level, 1, 0),

         schwarzer.test = ifelse(pval1.schwarzer < sig.level, 1, 0),
         rucker.test = ifelse(pval1.rucker < sig.level, 1, 0),
         harbord.test = ifelse(pval1.harbord < sig.level, 1, 0),
         peter.test = ifelse(pval1.peter < sig.level, 1, 0),
         I2 = max(c(0, 1 - (k-1)/Q)),
         i2f = factor(ifelse(I2 == 0, 1, 0)),
         se.stat.c = ifelse(outcome.flag == "DICH", stat.rucker, stat.thompson),
         se.stat = cut(se.stat.c,
                       breaks =  c(-7,1, 2, 4.5)))

meta.bin <- meta.f %>% filter(outcome.flag == "DICH")
meta.cont <- meta.f %>% filter(outcome.flag == "CONT")
meta.iv <- meta.f %>% filter(outcome.flag == "IV")

#Meta-Analysis and adjusted treatment effect estimate difference:

#Z-score:
diff.z.fixef.reg <- meta.f %>% 
  mutate(fixef = est.z.fixef, ranef = est.z.ranef,
         copas = est.z.copas, regression = est.z.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(fixef) ~  abs(copas),
                           sign(copas) != sign(fixef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(fixef) ~  abs(regression),
                                sign(regression) != sign(fixef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(se.stat, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.z.score", regression) %>% 
  mutate(difference = fixef - adjusted.z.score) %>% 
  filter(difference > -.75 & difference < .75) %>% 
  ggplot(aes(x = difference, fill = se.stat)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + theme(legend.position = "none") + 
  scale_fill_manual(values  = c("seagreen1", "seagreen3", "green4")) +
  xlab(expression(paste("Fixed effects - adjusted ", z, "-score"))) + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

diff.z.fixef.copas <- meta.f %>% 
  mutate(fixef = est.z.fixef, ranef = est.z.ranef,
         copas = est.z.copas, regression = est.z.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(fixef) ~  abs(copas),
                           sign(copas) != sign(fixef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(fixef) ~  abs(regression),
                                sign(regression) != sign(fixef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(se.stat, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.z.score", copas) %>% 
  mutate(difference = fixef - adjusted.z.score) %>% 
  filter(difference > -.75 & difference < .75) %>% 
  ggplot(aes(x = difference, fill = se.stat)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + theme(legend.position = "none") + 
  scale_fill_manual(values  = c("seagreen1", "seagreen3", "green4")) +
  xlab(expression(paste("Fixed effects - adjusted ", z, "-score"))) + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

diff.z.ranef.reg <- meta.f %>% 
  mutate(fixef = est.z.fixef, ranef = est.z.ranef,
         copas = est.z.copas, regression = est.z.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(se.stat, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.z.score", regression) %>% 
  mutate(difference = ranef - adjusted.z.score) %>% 
  filter(difference > -.75 & difference < .75) %>% 
  ggplot(aes(x = difference, fill = se.stat)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + theme(legend.position = "none") + 
  scale_fill_manual(values  = c("seagreen1", "seagreen3", "green4")) + ylab("") + 
  xlab(expression(paste("Random effects - adjusted ", z, "-score"))) + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

diff.z.ranef.copas <- meta.f %>% 
  mutate(fixef = est.z.fixef, ranef = est.z.ranef,
         copas = est.z.copas, regression = est.z.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(se.stat, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.z.score", copas) %>% 
  mutate(difference = ranef - adjusted.z.score) %>% 
  filter(difference > -.75 & difference < .75) %>% 
  ggplot(aes(x = difference, fill = se.stat)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + theme(legend.position = "none") + 
  scale_fill_manual(values  = c("seagreen1", "seagreen3", "green4")) + 
  xlab(expression(paste("Random effects - adjusted ", z, "-score"))) + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

#SMD:
diff.d.fixef.reg <- meta.f %>% 
  mutate(fixef = est.d.fixef, ranef = est.d.ranef,
         copas = est.d.copas, regression = est.d.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(fixef) ~  abs(copas),
                           sign(copas) != sign(fixef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(fixef) ~  abs(regression),
                                sign(regression) != sign(fixef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(se.stat, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.smd", regression) %>% 
  mutate(difference = fixef - adjusted.smd) %>% 
  filter(difference > -1 & difference < 1) %>% 
  ggplot(aes(x = difference, fill = se.stat)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + theme(legend.position = "none") + 
  scale_fill_manual(values  = c("seagreen1", "seagreen3", "green4")) + 
  xlab(expression(paste("Fixed effects - adjusted Hedges"))) + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

diff.d.fixef.copas <- meta.f %>% 
  mutate(fixef = est.d.fixef, ranef = est.d.ranef,
         copas = est.d.copas, regression = est.d.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(fixef) ~  abs(copas),
                           sign(copas) != sign(fixef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(fixef) ~  abs(regression),
                                sign(regression) != sign(fixef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(se.stat, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.smd", copas) %>% 
  mutate(difference = fixef - adjusted.smd) %>% 
  filter(difference > -1 & difference < 1) %>% 
  ggplot(aes(x = difference, fill = se.stat)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + theme(legend.position = "none") + 
  scale_fill_manual(values  = c("seagreen1", "seagreen3", "green4")) + 
  xlab(expression(paste("Fixed effects - adjusted std. mean difference"))) + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

diff.d.ranef.reg <- meta.f  %>% 
  mutate(fixef = est.d.fixef, ranef = est.d.ranef,
         copas = est.d.copas, regression = est.d.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  gather(key = "method", value = "adjusted.smd", regression) %>% 
  mutate(difference = ranef - adjusted.smd) %>% 
  filter(difference > -1 & difference < 1) %>% 
  ggplot(aes(x = difference, fill = se.stat)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + guides(fill=guide_legend(title="test stat.")) + 
  scale_fill_manual(values  = c("seagreen1", "seagreen3", "green4")) +  ylab("") +
xlab(expression(paste("Random effects - adjusted Hedges"))) + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

diff.d.ranef.copas <- meta.f  %>% 
  mutate(fixef = est.d.fixef, ranef = est.d.ranef,
         copas = est.d.copas, regression = est.d.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  gather(key = "method", value = "adjusted.smd", copas) %>% 
  mutate(difference = ranef - adjusted.smd) %>% 
  filter(difference > -1 & difference < 1) %>% 
  ggplot(aes(x = difference, fill = se.stat)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + guides(fill=guide_legend(title="test stat.")) +
  scale_fill_manual(values  = c("seagreen1", "seagreen3", "green4")) + ylab("") +
xlab(expression(paste("Random effects - adjusted std. mean difference"))) + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

#Log hazard ratios:
diff.log.hazard.ratio.fixef <- meta.iv %>% 
  mutate(fixef = est.fixef, ranef = est.ranef,
         copas = est.copas, regression = est.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(fixef) ~  abs(copas),
                           sign(copas) != sign(fixef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(fixef) ~  abs(regression),
                                sign(regression) != sign(fixef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  gather(key = "method", value = "adjusted.log.hazard.ratio", copas:regression)  %>% 
  ggplot(aes(x = (fixef - adjusted.log.hazard.ratio))) + 
  geom_histogram(binwidth = 0.1, center = 0) +
  facet_wrap(~method) + theme_bw() + xlab("Fixed effects - adjusted log effect") + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

diff.log.hazard.ratio.ranef <- meta.iv %>% 
  mutate(fixef = est.fixef, ranef = est.ranef,
         copas = est.copas, regression = est.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  gather(key = "method", value = "adjusted.log.hazard.ratio", copas:regression)  %>% 
  ggplot(aes(x = (ranef - adjusted.log.hazard.ratio))) + 
  geom_histogram(binwidth = 0.1, center = 0) +
  facet_wrap(~method) + theme_bw() + xlab("Random effects - adjusted log effect") + geom_vline(xintercept = 0, linetype = "dashed" )
#--------------------------------------------------------------------------------------------------------------------#

sig.level <- 0.1
meta.f <- meta.f %>% 
	mutate(egger.test = ifelse(pval1.egger < sig.level, 1, 0),
				 thompson.test = ifelse(pval1.thompson < sig.level, 1, 0),
				 begg.test = ifelse(pval1.begg < sig.level, 1, 0),

				 tes.d.test = ifelse(pval1.d.tes < sig.level, 1, 0),

				 schwarzer.test = ifelse(pval1.schwarzer < sig.level, 1, 0),
				 rucker.test = ifelse(pval1.rucker < sig.level, 1, 0),
				 harbord.test = ifelse(pval1.harbord < sig.level, 1, 0),
				 peter.test = ifelse(pval1.peter < sig.level, 1, 0))

#Quantile, mean, largerthan and missing table of differences:
adjustment.diff.keep <- meta.f %>% 
  mutate(est.copas.f = case_when(sign(est.copas) == sign(est.fixef) ~  abs(est.copas),
                                 sign(est.copas) != sign(est.fixef) ~ -abs(est.copas)),
         est.reg.f = case_when(sign(est.reg) == sign(est.fixef) ~  abs(est.reg),
                               sign(est.reg) != sign(est.fixef) ~ -abs(est.reg)),
         est.copas.r = case_when(sign(est.copas) == sign(est.ranef) ~  abs(est.copas),
                                 sign(est.copas) != sign(est.ranef) ~ -abs(est.copas)),
         est.reg.r = case_when(sign(est.reg) == sign(est.ranef) ~  abs(est.reg),
                               sign(est.reg) != sign(est.ranef) ~ -abs(est.reg)),

         est.d.copas.f = case_when(sign(est.d.copas) == sign(est.d.fixef) ~  abs(est.d.copas),
                                   sign(est.d.copas) != sign(est.d.fixef) ~ -abs(est.d.copas)),
         est.d.reg.f = case_when(sign(est.d.reg) == sign(est.d.fixef) ~  abs(est.d.reg),
                                 sign(est.d.reg) != sign(est.d.fixef) ~ -abs(est.d.reg)),
         est.d.copas.r = case_when(sign(est.d.copas) == sign(est.d.ranef) ~  abs(est.d.copas),
                                   sign(est.d.copas) != sign(est.d.ranef) ~ -abs(est.d.copas)),
         est.d.reg.r = case_when(sign(est.d.reg) == sign(est.d.ranef) ~  abs(est.d.reg),
                                 sign(est.d.reg) != sign(est.d.ranef) ~ -abs(est.d.reg)),

         est.z.copas.f = case_when(sign(est.z.copas) == sign(est.z.fixef) ~  abs(est.z.copas),
                                   sign(est.z.copas) != sign(est.z.fixef) ~ -abs(est.z.copas)),
         est.z.reg.f = case_when(sign(est.z.reg) == sign(est.z.fixef) ~  abs(est.z.reg),
                                 sign(est.z.reg) != sign(est.z.fixef) ~ -abs(est.z.reg)),
         est.z.copas.r = case_when(sign(est.z.copas) == sign(est.z.ranef) ~  abs(est.z.copas),
                                   sign(est.z.copas) != sign(est.z.ranef) ~ -abs(est.z.copas)),
         est.z.reg.r = case_when(sign(est.z.reg) == sign(est.z.ranef) ~  abs(est.z.reg),
                                 sign(est.z.reg) != sign(est.z.ranef) ~ -abs(est.z.reg)),

         est.fixef = abs(est.fixef),
         est.ranef = abs(est.ranef),
         est.z.fixef = abs(est.z.fixef),
         est.z.ranef = abs(est.z.ranef),
         est.d.fixef = abs(est.d.fixef),
         est.d.ranef = abs(est.d.ranef))

adjustment.diff.keep <- adjustment.diff.keep %>% 
  mutate(bias.side = ifelse(est.fixef - est.reg.f > 0, 1, 0), #Decide if effect size declines or increases

         egger.test = ifelse(bias.side == 1, egger.test, 0),
         thompson.test = ifelse(bias.side == 1, thompson.test, 0),
         begg.test = ifelse(bias.side == 1, begg.test, 0),

         schwarzer.test = ifelse(bias.side == 1, schwarzer.test, 0),
         peter.test = ifelse(bias.side == 1, peter.test, 0),
         rucker.test = ifelse(bias.side == 1, rucker.test, 0),
         harbord.test = ifelse(bias.side == 1, harbord.test, 0))

adjustment.diff <- adjustment.diff.keep %>%

  transmute(copas.z.f = est.z.fixef - est.z.copas.f,
            copas.z.r = est.z.ranef - est.z.copas.r,
            reg.z.f = est.z.fixef - est.z.reg.f,
            reg.z.r = est.z.ranef - est.z.reg.r,

            copas.d.f = est.d.fixef - est.d.copas.f,
            copas.d.r = est.d.ranef - est.d.copas.r,
            reg.d.f = est.d.fixef - est.d.reg.f,
            reg.d.r = est.d.ranef - est.d.reg.r)

m.adjustment.diff <- as.matrix(adjustment.diff)

adjustmend.diff.iv <- meta.iv %>% 
  transmute(est.copas.f = case_when(sign(est.copas) == sign(est.fixef) ~  abs(est.copas),
                                    sign(est.copas) != sign(est.fixef) ~ -abs(est.copas)),
            est.copas.r = case_when(sign(est.copas) == sign(est.ranef) ~  abs(est.copas),
                                    sign(est.copas) != sign(est.ranef) ~ -abs(est.copas)),

            est.reg.f = case_when(sign(est.reg) == sign(est.fixef) ~  abs(est.reg),
                                  sign(est.reg) != sign(est.fixef) ~ -abs(est.reg)),
            est.reg.r = case_when(sign(est.reg) == sign(est.ranef) ~  abs(est.reg),
                                  sign(est.reg) != sign(est.ranef) ~ -abs(est.reg)),

            est.fixef = abs(est.fixef),
            est.ranef = abs(est.ranef)) %>%

  transmute(copas.iv.f = est.fixef - est.copas.f,
            copas.iv.r = est.ranef - est.copas.r,

            reg.iv.f =  est.fixef - est.reg.f,
            reg.iv.r = est.ranef - est.reg.r)

m.adjustment.diff.iv <- as.matrix(adjustmend.diff.iv)

adjustment.diff.quantile.table <- apply(m.adjustment.diff, 2, 
                                        FUN = function(column){quantile(x = column, 
                                                                        probs = c(0.05, 0.25, 0.5, 0.75, 0.95), na.rm = T)})
adjustment.diff.mean <- apply(m.adjustment.diff, 2, FUN = function(column) mean(column, na.rm = T))
adjustment.diff.largerzero <- (apply(m.adjustment.diff, 2, FUN = function(column) length(which(column > 0)))/apply(m.adjustment.diff, 2, FUN = function(column) length(column)))*100
adjustment.diff.equalzero <- (apply(m.adjustment.diff, 2, FUN = function(column) length(which(column == 0)))/apply(m.adjustment.diff, 2, FUN = function(column) length(column)))*100
adjustment.diff.largerequalzero <- (apply(m.adjustment.diff, 2, FUN = function(column) length(which(column >= 0)))/apply(m.adjustment.diff, 2, FUN = function(column) length(column)))*100
adjustment.diff.comp.errors <- c(4,4,0,0,0,0,0,0)
adjustment.diff.noest <- c(sum(meta.f$est.z.copas.na[meta.f$outcome.flag != "IV"]), 
                                sum(meta.f$est.z.copas.na[meta.f$outcome.flag != "IV"]),
                                sum(meta.f$est.z.reg.na[meta.f$outcome.flag != "IV"]), 
                                sum(meta.f$est.z.reg.na[meta.f$outcome.flag != "IV"]),
                                sum(meta.f$est.d.copas.na[meta.f$outcome.flag != "IV"]), 
                                sum(meta.f$est.d.copas.na[meta.f$outcome.flag != "IV"]),
                                sum(meta.f$est.d.reg.na[meta.f$outcome.flag != "IV"]), 
                                sum(meta.f$est.d.reg.na[meta.f$outcome.flag != "IV"]))
adjustment.diff.missing <- (adjustment.diff.comp.errors + adjustment.diff.noest)/apply(m.adjustment.diff, 2, FUN = function(column) length(column))*100

adjustment.diff.table <- rbind(adjustment.diff.quantile.table, mean = adjustment.diff.mean,
                               equalzero = adjustment.diff.equalzero,
                               largerequal = adjustment.diff.largerequalzero, larger = adjustment.diff.largerzero,
                               missing = adjustment.diff.missing)

adjustment.diff.quantile.table.iv <- apply(m.adjustment.diff.iv, 2, 
                                             FUN = function(column){quantile(x = column, 
                                                                             probs = c(0.05, 0.25, 0.5, 0.75, 0.95), na.rm = T)})
adjustment.diff.mean.iv <- apply(m.adjustment.diff.iv, 2, FUN = function(column) mean(column, na.rm = T))
adjustment.diff.largerzero.iv <- (apply(m.adjustment.diff.iv, 2, FUN = function(column) length(which(column >= 0)))/apply(m.adjustment.diff.iv, 2, FUN = function(column) length(column)))*100
adjustment.diff.equalzero.iv <- (apply(m.adjustment.diff.iv, 2, FUN = function(column) length(which(column == 0)))/apply(m.adjustment.diff.iv, 2, FUN = function(column) length(column)))*100
adjustment.diff.largerequalzero.iv <- (apply(m.adjustment.diff.iv, 2, FUN = function(column) length(which(column > 0)))/apply(m.adjustment.diff.iv, 2, FUN = function(column) length(column)))*100
adjustment.diff.comp.errors.iv <- c(0,0,0,0)
adjustment.diff.noest.iv <- c(sum(meta.iv$est.copas.na), sum(meta.iv$est.copas.na),
                                sum(meta.iv$est.reg.na), sum(meta.iv$est.reg.na))
adjustment.diff.missing.iv <- ((adjustment.diff.comp.errors.iv + adjustment.diff.noest.iv)/apply(m.adjustment.diff.iv, 2, FUN = function(column) length(column)))*100

adjustment.diff.table.iv <- rbind(adjustment.diff.quantile.table.iv, mean = adjustment.diff.mean.iv,
                                    equalzero = adjustment.diff.equalzero.iv,
                                    largerequal = adjustment.diff.largerequalzero.iv, larger = adjustment.diff.largerzero.iv,
                                    missing = adjustment.diff.missing.iv)

adjustment.diff.table <- cbind(adjustment.diff.table, (adjustment.diff.table.iv))

adjustment.diff.table <- round(t(adjustment.diff.table), 4)

rownames(adjustment.diff.table) <- c(paste("z:", c("Fixed - Copas", "Random - Copas", "Fixed - Regression", "Random - Regression")),
                                     paste("g:", c("Fixed - Copas", "Random - Copas", "Fixed - Regression", "Random - Regression")),
                                     paste("IV:", c("Fixed - Copas", "Random - Copas", "Fixed - Regression", "Random - Regression")))

colnames(adjustment.diff.table) <- c(colnames(adjustment.diff.table)[1:6], "= 0 (%)", ">= 0 (%)", "> 0 (%)", "No adj. est. (%)")

#--------------------------------------------------------------------------------------------------------------------#
#Meta-Analysis and adjusted treatment effect estimate difference, missing values:

#Z-score:
missing.z.fixef <- meta.f %>% 
  mutate(fixef = est.z.fixef, ranef = est.z.ranef,
         copas = est.z.copas, regression = est.z.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(fixef) ~  abs(copas),
                           sign(copas) != sign(fixef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(fixef) ~  abs(regression),
                                sign(regression) != sign(fixef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(meta.id, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.z.score", copas:regression) %>% 
  mutate(difference = fixef - adjusted.z.score) %>% 
  filter(difference < -.75 | difference > .75) %>% 
  select(meta.id)

#--------------------------------------------------------------------------------------------------------------------#

missing.z.ranef <- meta.f %>% 
  mutate(fixef = est.z.fixef, ranef = est.z.ranef,
         copas = est.z.copas, regression = est.z.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(meta.id, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.z.score", copas:regression) %>% 
  mutate(difference = ranef - adjusted.z.score) %>% 
  filter(difference < -.75 | difference > .75) %>% 
  select(meta.id)

#--------------------------------------------------------------------------------------------------------------------#

#SMD:
missing.d.fixef <- meta.f %>% 
  mutate(fixef = est.d.fixef, ranef = est.d.ranef,
         copas = est.d.copas, regression = est.d.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(fixef) ~  abs(copas),
                           sign(copas) != sign(fixef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(fixef) ~  abs(regression),
                                sign(regression) != sign(fixef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(meta.id, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.smd", copas:regression) %>% 
  mutate(difference = fixef - adjusted.smd) %>% 
  filter(difference < -1 | difference > 1) %>% select(meta.id)

#--------------------------------------------------------------------------------------------------------------------#

missing.d.ranef <- meta.f %>% 
  mutate(fixef = est.d.fixef, ranef = est.d.ranef,
         copas = est.d.copas, regression = est.d.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(meta.id, fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.smd", copas:regression) %>% 
  mutate(difference = ranef - adjusted.smd) %>% 
  filter(difference < -1 | difference > 1) %>% select(meta.id)

missing.differences <- rbind(missing.z.fixef, missing.z.ranef, missing.d.fixef, missing.d.ranef)
missing.differences <- missing.differences %>% distinct(meta.id)

missing.differences <- merge(meta.f[,
        c("meta.id", "id", "comparison.nr", "subgroup.nr",
          "est.z.ranef", "est.z.fixef", "est.z.copas", "est.z.reg", 
          "est.d.fixef", "est.d.ranef", "est.d.copas", "est.d.reg")],
        missing.differences, by = "meta.id")

missing.differences[, c("est.z.ranef", "est.z.fixef", "est.z.copas", "est.z.reg", 
          "est.d.fixef", "est.d.ranef", "est.d.copas", "est.d.reg")] <- round(missing.differences[, c("est.z.ranef", "est.z.fixef", "est.z.copas", "est.z.reg", 
          "est.d.fixef", "est.d.ranef", "est.d.copas", "est.d.reg")], 2)

missing.differences <- missing.differences %>% mutate(meta.id = as.character(meta.id), id = as.character(id),
                                                      comparison.nr = as.character(comparison.nr), subgroup.nr = as.character(subgroup.nr))

colnames(missing.differences) <- c("meta.id", "id", "comparison.nr", "subgroup.nr",
                                   "z fixed", "z random", "z Copas", "z regression",
                                   "g fixed", "g random", "g Copas", "g regression")

#--------------------------------------------------------------------------------------------------------------------#
#Tukey mean difference plots for adjustment methods:

adjustment.mean.diff.orig <- adjustment.diff.keep %>% filter(est.copas.na != 1) %>% 
  mutate(mean = (est.copas + est.reg)/2, 
         difference = est.copas - est.reg) %>% filter(difference < 10 & mean < 20) %>% 
  ggplot(aes(x = mean, y = difference)) + geom_point(size = .5) + 
  geom_abline(aes(slope = 0, intercept = mean(difference, na.rm = T), col = "blue")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) + 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) - 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  # geom_smooth(method = "loess", se = F) + 
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(labels = c("systematic error", "95 % confidence intervals"), 
                     values = c("red", "blue")) +
  ggtitle(expression(paste("Copas adjusted estimate - Regression adjusted estimate (original)" )))

adjustment.mean.diff.z <- adjustment.diff.keep %>% filter(est.z.copas.na != 1) %>% 
  mutate(mean = (est.z.copas + est.z.reg)/2, 
         difference = est.z.copas - est.z.reg) %>% filter(mean < 2.5 & difference < 5) %>% 
  ggplot(aes(x = mean, y = difference)) + geom_point(size = 0.5) + 
  geom_abline(aes(slope = 0, intercept = mean(difference, na.rm = T), col = "blue")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) + 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) - 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  # geom_smooth(method = "loess", se = F) + 
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(labels = c("systematic error", "95 % confidence intervals"), 
                     values = c("red", "blue")) +
  ggtitle(expression(paste("Copas adjusted ", italic(z),"-score estimate - Regression adjusted ", italic(z),"-score estimate (original)" )))

adjustment.mean.diff.d <- adjustment.diff.keep %>% filter(est.d.copas.na != 1) %>% 
  mutate(mean = (est.d.copas + est.d.reg)/2, 
         difference = est.d.copas - est.d.reg) %>% filter(difference < 7 & mean < 10) %>% 
  ggplot(aes(x = mean, y = difference)) + geom_point(size = .5) + 
  geom_abline(aes(slope = 0, intercept = mean(difference, na.rm = T), col = "blue")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) + 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  geom_abline(aes(slope = 0,
                  intercept = mean(difference, na.rm = T) - 1.96*sqrt(var(difference, na.rm = T)), 
                  col = "red")) +
  # geom_smooth(method = "loess", se = F) + 
  theme_bw() + #theme(legend.position = "bottom") +
  scale_color_manual(labels = c("systematic error", "95 % confidence intervals"), 
                     values = c("red", "blue")) + guides(fill=guide_legend(title=NULL)) +
  ggtitle(expression(paste("Copas adjusted std. mean difference estimate - Regression adjusted std. mean difference estimate (original)" )))

#--------------------------------------------------------------------------------------------------------------------#

#Missing study number by copas:
missing.copas.histogram <- meta.f %>% ggplot(aes(x = missing.copas/k)) + geom_histogram(binwidth = 0.05, boundary = 0) + theme_bw() + xlab("Missing study fraction")
missing.copas.zeros <- length(which(meta.f$missing.copas[!is.na(meta.f$missing.copas)]/meta.f$k[!is.na(meta.f$missing.copas)] == 0))
missing.copas.range.frac <- c(missing.copas.zeros, quantile(x = meta.f$missing.copas[!is.na(meta.f$missing.copas)]/meta.f$k[!is.na(meta.f$missing.copas)], c(0.05, 0.25, 0.5, 0.75, 0.95)), mean(meta.f$missing.copas[!is.na(meta.f$missing.copas)]/meta.f$k[!is.na(meta.f$missing.copas)],na.rm = T) )
missing.copas.range.count <- c(missing.copas.zeros, quantile(x = meta.f$missing.copas[!is.na(meta.f$missing.copas)], c(0.05, 0.25, 0.5, 0.75, 0.95)), mean(na.rm = T,meta.f$missing.copas[!is.na(meta.f$missing.copas)])) 
missing.copas.table <- rbind(missing.copas.range.frac, missing.copas.range.count)
rownames(missing.copas.table) <- c("Missing fraction", "Missing study number")
colnames(missing.copas.table) <- c("= 0", colnames(missing.copas.table)[2:6], "mean")

@

\subsection{Change in Effect Size after Adjustment} \label{sec:change.size}

There are methods that can take into account the presence of publication bias in meta-analyses when estimating the overall treatment effect. The methods work in a semi-automatic manner; they will not only adjust for publication bias if smaller studies show larger effects, but also in the opposite case. The latter results in the adjusted overall treatment effect being \textit{larger} than the unadjusted, overall treatment effect. \\
To compare the effects of adjustment between meta-analyses of different outcomes, the outcome measures are transformed to standardized mean differences and Fisher's $z$-scores (see section \ref{sec:transformation.effectsizes} for details). When comparing to unadjusted effects, fixed or random effects meta-analysis estimates are used as references\\
Figure \ref{fig:adjustment.reg} displays the difference $\delta$ between the estimated meta-analysis treatment effect and the regression adjusted treatment effect \ref{sec:regression.adjustment}, $\hat{\theta}_M - \hat{\theta}_\textrm{Adj.}$. The absolute value $|\hat{\theta}_M|$ is taken and $\hat{\theta}_\textrm{Adj.}$ is negative if the signs change in the original $\theta$. Thus, a positive difference indicates a reduction of the original effect size, and the magnitude of the difference indicates the extent of the adjustment. \\
Additionally, the test statistic of heterogeneity adjusted publication bias tests (R\"ucker's and Thompson's test) are displayed with green color. Test statitics smaller $< 1$ (light green colored) are equivalent to no evidence for publication test, test statistics between one and two to weak evidence, and above two they indicate evidence for publication bias (dark green). An adjusted effect with evidence for publication bias can be regarded as a more realistic estimate of treatment efficacy. Some very large and very small differences have been omitted in the $z$-score and std. mean difference histograms; they are shown in Table \ref{missing.differences}. \\

\begin{figure}
<<echo=FALSE, fig.height = 5.2, warning=FALSE>>=
grid.arrange(diff.z.fixef.reg, diff.z.ranef.reg, diff.d.fixef.reg, diff.d.ranef.reg,
             ncol = 2)
@
\caption{Histogram of the treatment effect differences between meta-analysis and regression adjusted meta-analysis. Negative differences indicate greater adjusted effect sizes than meta-analysis effect sizes. The bins are centered at zero and binwidth is equal to 0.1. Deeper green color indicates more evidence for small study effects.}
\label{fig:adjustment.reg}
\end{figure}

Most often, adjustment leads to a reduction of overall treatment effect estimates, because the bins on the positive side of the histograms are larger. Adjustment is stronger when random effects meta-analysis is used as reference, because it gives larger weights to small studies.\\
Contrary to naive expectation, we see cases with large adjustment, but no evidence for small study effects. This is because the linear regression parameter estimates are large, but estimated with high uncertainity, such that there will be few evidence for small study effects (publication bias), but nonetheless, the adjustment will be large. The authors that developped the methods also recommend to use the methods in cases where there is clear evidence for publication bias. The color legend in Figure \ref{fig:adjustment.reg} thus gives a sense of confidence that there is in the credibility of the adjustment. \\
Additionally, some meta-analyses with positively adjusted, larger overall treatment effects after adjustment (left side of the histogram) have at the same time evidence for publication bias. This is no wrong result, although the $p$-value of the test is one-sided, \ie only allowing for publication bias for large effects (small studies with large effects). But the effects and their variances have been transformed, and it is possible that the shape of the funnel plot changes upon transformation; Figure \ref{fig:funnel.plot.change} shows this for illustratory purposes. From the left to the right, the funnel plot is shown for mean differences (the original measure), standardized mean differences and Fisher's $z$ scores. This is quite the most extreme case, with three different effects of adjustment, depending on outcome measure; there is no change with mean differences, reduction of effect sizes with std. mean differences and increase with Fisher's $z$ transformed correlation. Note that while the rank of the effect sizes is usually preserved after transformation, the relative size and especially the variance may vary. One effect of the Fisher's $z$-transformation is that the effect sizes are bounded on $[-1,1]$, and thus, very large effect sizes will influence the fit of the linear regression less than for example in std. mean differences which are not bounded. In contrast, the variance of the correlation is directly tied to the sample size, which makes it a suitable proxy for study size (variance of the mean difference is in contrast strongly influenced by the standard deviations). Thus, cases where the direction of adjustment changes upon transformation are not very rare.

\begin{figure}
<<echo = FALSE, warning=FALSE>>=
par(mfrow = c(1,3))
example <- data.ext2 %>% filter(id == "CD003236", subgroup.nr == 2, outcome.nr == 1, comparison.nr == 1)
funnel(limitmeta(metagen(TE = effect, se = se, studlab = study.name, sm = "MD", data = example)), col.adjust = "blue", col.line = "blue")
funnel(limitmeta(metagen(TE = smd.pool, se = se.smd.pool, studlab = study.name, sm = "SMD", data = example)), col.adjust = "blue", col.line = "blue")
funnel(limitmeta(metacor(cor = z, n = total1 + total2, studlab = study.name, sm = "ZCOR", data = example)), col.adjust = "blue", col.line = "blue")
par(mfrow = c(1,1))
@
\caption{Funnel plots for a meta-analysis based on three different effect size measures: Mean differences, std. mean differences and Fisher's z transformed correlations and corresponding standard errors. Vertical dashed lines indicate meta-analysis estimates, the rhombus with the curved blue line the adjusted treatment effect.}
\label{fig:funnel.plot.change}
\end{figure}

Figure \ref{fig:adjustment.copas} shows the same for effect sizes adjusted by Copas; Copas selection model substitutes its estimates with random effect estimates when it finds no evidence for small study effects. Therefore, the effect of adjustment by Copas can be seen when comparing adjusted with random effects meta analysis estimates. Again, we clearly see that more effect sizes are adjusted downwards. Additionally, there is more coincidence between the publication bias test statistics and adjustment, which is as expected. \\

\begin{figure}
<<echo=FALSE, fig.height = 4, warning=FALSE>>=
grid.arrange(diff.z.ranef.copas, diff.d.ranef.copas, ncol = 2)
@
\caption{Histogram of the treatment effect differences between meta-analysis and Copas adjusted meta-analysis. Negative differences indicate greater adjusted effect sizes than meta-analysis effect sizes. The bins are centered at zero and binwidth is equal to 0.1. Deeper green color indicates more evidence for small study effects.}
\label{fig:adjustment.copas}
\end{figure}

Table \ref{adjustment.difference} shows quantiles and means for the various differences and the overall proportion of downward adjusted effect sizes. When Hedge's $g$ is used as an effect measure, there are (substantially) more reduced effect sizes. The means in Table \ref{adjustment.difference} suggest that the average reduction is small. To recall some other findings out of Table \ref{adjustment.difference}: 5\% or \Sexpr{floor(length(!is.na(meta.f$est.z.reg))*0.05)} cases have their $z$-score reduced by more than 0.15 by regression adjustment (and 5\% or \Sexpr{floor(length(!is.na(meta.f$est.z.reg))*0.05)} increased by -0.13, fixed effects reference). Also, std. mean difference is reduced by 0.39 compared to fixed effects estimates in 5\% or \Sexpr{floor(length(!is.na(meta.f$est.d.reg))*0.05)} meta-analyses (or increased by 0.24).

<<echo=FALSE, results='asis'>>=
print(xtable(adjustment.diff.table, label = "adjustment.difference", caption = "Quantiles and Means of the differences between meta-analysis pooled treatment effects and small study adjusted treatment effects. The column with the names ``> 0'' give the percentages of estimates larger than zero or larger or equal zero. The column ``No adj. est.'' gives the percentage of missing estimates due to non-significant publication bias test (for Copas) and computational errors. The row names indicate which outcome measure, meta-analysis method and adjustment method is used. Abbreviations are used for z-score (= z) and Hedges g (= g).", align = "lcccccccccr"), include.rownames = T, size = "scriptsize",# sanitize.text.function=function(x){x},
      # table.placement = "p"
      )
@

\subsection{Change in Evidence for Treatment Effects} \label{sec:change.evidence}
Adjustment for small study effects in meta-analysis will not only provide new effect sizes, but also standard errors thereof. Thus, also the evidence for efficacy of a treatment can be obtained, which is usually summarized in a suitable test statistic or $p$-value. It is of interest if and how the evidence for treatment effects changes if adjusted for publication bias. There is an important difference between Copas selection model and regression adjustment. The regression adjustment method incorporates an additional parameter with corresponding uncertainity in the treatment effect estimate, while the application of the Copas algorithm does not so. Instead, the copas selection model derives the uncertainity of the estimate from the inverse of the fisher information matrix of the likelihood. Additionally, the Copas selection model algorithm provides only an estimate when it can detect a small study effect with reasonable precision at the first place (see \ref{sec:copas}).\\
The Wald test statistics $p$-value for fixed and random effects meta-analyses and Copas and regression adjusted treatment effects have been calculated. They are shown in Figure \ref{fig:adjustment.stat} for meta-analysis based on $z$-score, std. mean difference and other effect measures (\texttt{outcome.flag} = \texttt{IV}). \\
It can be seen that evidence for a treatment effect decreases after adjusting for publication bias. It does so more for Hedge's $g$ compared to $z$-scores, and for regression adjustment compared to Copas selection model. 
%Also, the evidence decreases for random effects meta-analyses compared to fixed effects meta-analyses. 
In the case of other effect measures (\texttt{outcome.flag} = \texttt{IV}) adjustment, the adjustment has only negligible impact on the $p$-values. \\
Also, Copas selection method has a very small impact on the evidence compared to random effects meta-analysis, at least using $z$-scores (again, we should rely on comparisons to random effects meta-analysis to evaluate Copas method). \\ 
% Interestingly, in contrast to the small study effect tests, there seems to be no large difference between adjusted continuous and binary meta-analysis test statistics --appendix--
%Figures \ref{fig:adjustment.stat.z} for the adjusted $z$-scores, \ref{fig:adjustment.stat.smd} for the adjusted std. mean difference and Cohen's d and \ref{fig:adjustment.stat.log.hazard.ratio} for the adjusted log hazard ratios.

\begin{figure}
<<echo=FALSE, fig.height = 9, warning=FALSE>>=
grid.arrange(adjustment.p.z,
             adjustment.p.d,
             adjustment.p.log.hazard.ratio, ncol = 1)
@
\caption{Histogram of the Wald test-statistic $p$-value of meta-analysis and adjusted pooled treatment effect, based on different treatment effect measures. The method is indicated in the header, bin width is set to 0.05. The significant proportion based on the threshold of 0.05 is displayed inside the figures.}
\label{fig:adjustment.stat}
\end{figure}

The Copas selection model also gives an estimate of the number of missing studies. It finds that \Sexpr{round(sum(meta.f$missing.Copas, na.rm = T))} are missing, which corresponds to \Sexpr{round((sum(meta.f$missing.copas, na.rm = T)/sum(meta.f$k))*100,1)}\% from all \Sexpr{round(sum(meta.f$k))} analysed studies. Figure \ref{fig:copas.missing} shows a histogram of the overall fraction of missing studies. Note that we have excluded \Sexpr{sum(is.na(meta.f$missing.Copas))} out of \Sexpr{dim(meta.f)[1]}, for which no Copas selection model estimate, but a random effects estimate was retained because the algorithm initially found no evidence for small study effects.

\begin{figure}
<<echo = FALSE, warning = FALSE>>=
missing.copas.histogram
@
\caption{Histogram of the fraction of missing studies from the total number of studies in a meta-analyses (only data shown where Copas estimate was obtained, thus \\$n =$ \Sexpr{sum(!is.na(meta.f$missing.Copas))})}
\label{fig:copas.missing}
\end{figure}

We can see that in some occasions, the method finds more than half of all studies are missing. In most occasions, the estimate of missing studies is zero, as can be seen in Table \ref{copas.missing}. The discrepancy between mean and median may indicate that the estimate of \Sexpr{round((sum(meta.f$missing.Copas, na.rm = T)/sum(meta.f$k))*100,1)}\% missing studies depends somewhat on these extreme cases. As can be written of from Table \ref{copas.missing}, 5\%, \ie \Sexpr{round(sum(!is.na(meta.f$missing.Copas))*0.05)} meta-analyses have 17.6 or more studies missing: in fact, these 5\% most extreme make up for \Sexpr{sum(meta.f$missing.Copas[which(meta.f$missing.Copas >= 17.60)], na.rm = T)}, more than 30\% of missing studies.

<<echo= FALSE, results='asis'>>=
print(xtable(missing.copas.table, caption = "Fraction of missing studies and estimates of missing studies with their zero counts (``= 0''), quantiles and means.",
             label = "copas.missing", align = "lrrrrrrr", digits = c(0,0,0,0,1,1,1,1)),
      size = "footnotesize")
@

<<echo=FALSE, results='asis'>>=
print(xtable(missing.differences, 
             caption = "Missing meta-analysis pooled treatment effect and adjusted treatment effects. Abbreviations are used for z-score (= z) and Hedges g (= g).", label = "missing.differences", align = "llcccrrrrrrrr"), 
      include.rownames = F, size = "tiny")
@

\subsection{Comparison of adjustment methods}
Regression adjusted estimates are compared to the estimates of Copas selection model if these are not equal to random effects meta-analysis. A Tukey mean difference plot can serve to reveal systematic differences and biases between the two measurement methods. We will compare estimates based on all three outcome measures, i.e. original measure, $z$-score and std. mean difference in Figure \ref{fig:adjustment.mean.diff}. Note that the \Sexpr{dim(meta.iv)[1]} \texttt{IV} outcome data is not included when using $z$-score and std. mean difference.

\begin{figure}
<<echo = FALSE, fig.height = 7.5>>=
grid.arrange(adjustment.mean.diff.orig,
             adjustment.mean.diff.z,
             adjustment.mean.diff.d, ncol = 1)
@
\caption{Mean - difference plots for publication bias adjustment methods. The mean of the adjusted treatment effects is displayed on the $x$-axis and the difference on the $y$-axis. Blue and red lines display the systematic error and the confidence intervals of the systematic error (limits of agreement). Two values have been omitted in the middle plot for std. mean difference and one for $z$-score (see Table \ref{missing.differences}).}
\label{fig:adjustment.mean.diff}
\end{figure}

No formal tests are provided, but the at least there seems to be no clear bias or systematic error. The limits of agreement in Figure \ref{fig:adjustment.mean.diff} are large. We conclude thus that the impact of regression adjustment on the effect sizes is in general not substantially larger than the impact of Copas selection model in the subset of data where the estimate of the Copas selection model is not equal to a random effects estimate. There is however a small difference, with regression estimates to have a little bit a larger absolute value. There might be some bias between adjusted $z$-scores, where regression estimates seem to be somewhat smaller when the mean is a little above zero, and somewhat larger when the mean is a little below zero. \\