### 2019-08-14 report/MSc_Report.Rnw
Draft for Simon and Prof. Held

\documentclass[11pt,a4paper,twoside]{book}
\input{header.sty}   % packages, layout and standard macros
\input{newCommands.tex}

\usepackage{verbatim}
\usepackage{tikz}
\usepackage{amssymb}
\usetikzlibrary{trees}
\usetikzlibrary{shapes,arrows}

\begin{document}
\input{title}

<<'main-setup',include=FALSE,purl=FALSE>>=
library(knitr)
opts_chunk$set(
    concordance=TRUE,
    child.command = 'include',
    cache=TRUE)
@

\graphicspath{{./figure/}}
\DeclareGraphicsExtensions{.pdf,.png}
\setcounter{tocdepth}{1}

\thispagestyle{empty}
\newpage

\pagenumbering{roman}

\chapter*{Abstract}
% \addtocontents{toc}{\protect \vspace*{13.mm}}
% \addcontentsline{toc}{chapter}{\bfseries{Preface}}
\thispagestyle{plain}\markboth{Abstract}{Abstract}
It is well known that the part of scientific findings that are published in journals are a 
biased sample of the overall amount of research that is done. Often, editors aim to publish
new and influential results that will be distributed further by other researchers and
heighten their journal impact factor. The downside of it is that some of these results
might only be positive by chance, or because of poor study design. When decision
makers use this information to take decisions, they are likely to be mislead. \\
In this study, the prevalence and impact of publication bias is assessed in 
a large scale analysis. The data stems from the Cochrane Library of Systematic Reviews, the biggest collection
of clinical trial and public health intervention data. Meta-Analyses of Cochrane
are scrutinized for small-study effects and excess significance, both signs
of publication bias. We also adjust the treatment effects from the meta-analysis for publication bias 
and compare the newly obtained effects with unadjusted treatment effects. In a exploratory analysis, we find
that around 20\% of meta-analyses are subject to publication bias, and that the
evidence for treatment efficacy decreases after adjustment. Publication bias
thus remains a threat to the validity of findings in Cochrane and in clinical research.
% Because of the
% amount of data that underlies this analysis, we suggest that previous measures 
% to curb publication bias need to be enforced and possibly, that new measures have 
% to be taken to prevent patient harm and strengthen the reliability of clinical research.

\chapter*{Acknowledgments}
\thispagestyle{plain}\markboth{Acknowledgments}{Acknowledgments}
I would like to thank all members of the Department of Biostatistics for promoting me along my studies and making this thesis possible.
Special thanks goes to Dr. Simon Schwab, for thorough support, and to Prof. Dr. Leonhard Held for advice. \\
Among my friends, I would to thank all students of biostatistics in the years 2018 and 2019 for making this time so pleasant. Samuel
Pawel was a very good friend with lots of advice and understanding throughout this time, and it was for Eleftheria Michalopoulou 
that writing this thesis was also fun.\\
Most important, my family always supported me and gave me confidence in myself. Denise Sternlicht and
Annatina Kreiliger are two of the main reasons why this thesis has been done, and I would like to dedicate this
thesis to them.

\bigskip

\begin{flushright}
  Giuachin Kreiliger\\
  June 2019
\end{flushright}

%\addtocontents{toc}{\protect \vspace*{10mm}}
% \addtocontents{toc}{\protect \vspace*{13.mm}}
% \addcontentsline{toc}{chapter}{\bfseries{Preface}}

\thispagestyle{plain}\markboth{Contents}{Contents}
\tableofcontents
\setkeys{Gin}{width=.8\textwidth}

% Howdy!

%\cleardoublepage
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter01', child='chapter01.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter02', child='chapter02.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter03', child='chapter03.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter04', child='chapter04.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<'child-chapter05', child='chapter05.Rnw'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix
<<'child-chapterA1', child='chapterA1.Rnw'>>=
@

\cleardoublepage
\phantomsection
\addtocontents{toc}{\protect \vspace*{10mm}}
\addcontentsline{toc}{chapter}{\bfseries Bibliography}

\bibliographystyle{mywiley} 
\bibliography{biblio}

\cleardoublepage

\end{document}