### 2019-08-28 Presentation Final/present.Rnw
Deletion of unnecessary files

\documentclass[english]{beamer}
%% The most common packages are already included in:
\usetheme{biostat}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage{amsmath,amsfonts,tikz, amssymb}
\usetikzlibrary{trees}

%% Header data: (adjust to your needs:
\def\uzhunit{Biostatistics}             %% if (not) needed comment/uncomment
%\def\uzhunitext{STA480}

\title{Publication Bias in Cochrane Meta-Analyses}%[Publication Bias]
%% Optional Argument in [Brackets]: Short Title for Footline

%% The following are all optional, simply comment them
%\subtitle{Publication Bias in the Cochrane Libary}
%\institute{Biostatistics Journal Club}  %% optional
\author{Giuachin Kreiliger}
\date{\today}

\input{newCommands.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
<<include = FALSE, cache = TRUE>>= 
#Load data:
rm(list = ls())
library(knitr)
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
PATH2 = file.path(PATH_HOME, 'PubBias')
FILE = 'cochrane_2019-07-04.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH2, 'code')
PATH_RESULTS = file.path(PATH2, 'results_new')
# PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

source(file.path(PATH_CODE, 'PubBias_functions.R'))
require(corrgram)
require(nlme)

load(file.path(PATH_DATA, "PubBias_2019-07-19.RData"))
load(file.path(PATH_RESULTS, "data_used_for_analysis.RData"))
load(file.path(PATH_RESULTS, "meta_analyses_summary_complete.RData"))
load(file.path(PATH_RESULTS, "meta_id_vector.RData"))
load(file.path(PATH_RESULTS, "meta_id_I2.RData"))
load(file.path(PATH_RESULTS, "meta.bin.RData"))
load(file.path(PATH_RESULTS, "meta.cont.RData"))
load(file.path(PATH_RESULTS, "meta.iv.RData"))

bias.side.z <- function(yi, sei){
  alpha = 0.05
  to.ommit <- which(is.na(yi) | is.na(sei))

  if(length(to.ommit) > 0){
    yi <-  yi[-to.ommit]
    sei <- sei[-to.ommit]
  }

  to.ommit <- which(sei == 0)

  if(length(to.ommit) > 0){
    yi <-  yi[-to.ommit]
    sei <- sei[-to.ommit]
  }

  if(sum(pnorm(yi/sei) < .05) == sum(pnorm(yi/sei, lower.tail=FALSE) < .05)){
      side <- sign(est.fe <- rma(yi = yi, sei = sei, method = "FE")$b[1] )
  } else{
    side = ifelse(sum(pnorm(yi/sei) < .05) > sum(pnorm(yi/sei, lower.tail=FALSE) < .05),
                  -1, 1)
    }
  return(side)
}

#Inital dataset properties:
initial.id.count <- length(unique(data$id))
initial.study.count <- length(unique(data$study.name))
initial.result.count <- dim(data)[1]

#No withdrawn reviews:
tp7 <- data.ext2 %>% filter(id %in% data.review$id[which(data.review$withdrawn == FALSE)]) 
nowith.id.count <- length(unique(tp7$id))
nowith.study.count <- length(unique(tp7$study.name))
nowith.result.count <- dim(tp7)[1]

#Only efficacy outcomes:
tp8 <- data.ext2 %>% filter(id %in% data.review$id[which(data.review$withdrawn == FALSE)]) %>% 
  filter(outcome.desired == "efficacy")
efficacy.id.count <- length(unique(tp8$id))
efficacy.study.count <- length(unique(tp8$study.name))
efficacy.result.count <- dim(tp8)[1]

#outcome.flag reduction:
tp.2 <- data.ext2 %>% filter(id %in% data.review$id[which(data.review$withdrawn == FALSE)]) %>% 
  filter(outcome.flag == "IV") %>% 
  filter(!all(effect == 0)) %>% 
  filter(outcome.measure.merged == "Rate Ratio" | outcome.measure.merged == "SMD" |
           outcome.measure.merged == "MD" | outcome.measure.merged == "Hazard Ratio" |
           outcome.measure.merged == "OR" | outcome.measure.merged == "RR")
tp.1 <- data.ext2 %>% filter(id %in% data.review$id[which(data.review$withdrawn == FALSE)]) %>% 
  filter(outcome.flag == "DICH" | outcome.flag == "CONT")
tp <- bind_rows(tp.1, tp.2)
flag.id.count <- length(unique(tp$id))
flag.study.count <- length(unique(tp$study.name))
flag.result.count <- dim(tp)[1]
#----------------------------------------------------------------------------------------------#

#Counts of meta-analyses with n larger ten:
m.a.largerten.bin <- data.ext2 %>% filter(id %in% data.review$id[which(data.review$withdrawn == FALSE)]) %>% 
  filter(outcome.desired == "efficacy") %>% 
  filter(outcome.flag == "DICH") %>% 
  filter(total1 != 0 & total2 != 0) %>% 
  group_by(meta.id) %>% filter(n() > 9) %>%  filter(any(events1 != 0) | any(events2 != 0)) %>% 
  distinct(meta.id) %>% ungroup() %>% mutate(n = n())

m.a.largerten.cont <- data.ext2 %>% filter(id %in% data.review$id[which(data.review$withdrawn == FALSE)]) %>% 
  filter(outcome.desired == "efficacy") %>% 
  filter(outcome.flag == "CONT") %>% 
  filter(sd1 != 0 & sd2 != 0) %>% 
  filter(total1 > 0 & total2 > 0) %>% group_by(meta.id) %>% 
  filter(n() > 9) %>% distinct(meta.id) %>% ungroup() %>% mutate(n = n())

m.a.largerten.iv <- data.ext2 %>% filter(id %in% data.review$id[which(data.review$withdrawn == FALSE)]) %>% 
  filter(outcome.desired == "efficacy") %>% 
  filter(outcome.flag == "IV") %>% 
  filter(!all(effect == 0)) %>% 
  filter(outcome.measure.merged == "Rate Ratio" | outcome.measure.merged == "SMD" |
           outcome.measure.merged == "MD" | outcome.measure.merged == "Hazard Ratio" |
           outcome.measure.merged == "OR" | outcome.measure.merged == "RR") %>% 
  filter(se != 0) %>% group_by(meta.id) %>% filter(n() > 9) %>% 
  distinct(meta.id) %>% ungroup() %>% mutate(n = n())

m.a.largerten.id <- rbind(m.a.largerten.bin, m.a.largerten.cont, m.a.largerten.iv)
tp2 <- data.ext2 %>% filter(meta.id %in% m.a.largerten.id$meta.id)

ten.id.count <- length(unique(tp2$id))
ten.study.count <- length(unique(tp2$study.name))
ten.result.count <- dim(tp2)[1]
m.a.largerten.count <- unique(m.a.largerten.bin$n) + unique(m.a.largerten.cont$n) + unique(m.a.largerten.iv$n)

#----------------------------------------------------------------------------------------------#

#variance ratio > 4:
m.a.variance4.count <- data.ext2 %>% filter(meta.id %in% m.a.largerten.id$meta.id) %>% 
  group_by(meta.id) %>% mutate(se.new = case_when(outcome.flag == "DICH" ~ sqrt(var.lrr),
                                                  TRUE ~ se)) %>% 
  summarise(ratio = (max(se.new)^2)/(min(se.new)^2)) %>% 
  filter(ratio >= 4) %>% count
tp3 <- data.ext2 %>% filter(meta.id %in% m.a.largerten.id$meta.id) %>% 
  group_by(meta.id) %>% mutate(se.new = case_when(outcome.flag == "DICH" ~ sqrt(var.lrr),
                                                  TRUE ~ se)) %>% 
  mutate(ratio = (max(se.new)^2)/(min(se.new)^2)) %>% 
  filter(ratio >= 4)
var.id.count <- length(unique(tp3$id))
var.study.count <- length(unique(tp3$study.name))
var.result.count <- dim(tp3)[1]

#At least on sig. estimate:
m.a.one.sig.count <- data.ext2 %>% filter(meta.id %in% m.a.largerten.id$meta.id) %>% 
  group_by(meta.id) %>% mutate(se.new = case_when(outcome.flag == "DICH" ~ sqrt(var.lrr),
                                                  TRUE ~ se)) %>% 
  summarise(ratio = (max(se.new)^2)/(min(se.new)^2),
                                  n.sig = sum(sig.single)) %>% 
  filter(ratio >= 4 & n.sig > 0) %>% count
tp4 <- data.ext2 %>% filter(meta.id %in% m.a.largerten.id$meta.id) %>% 
  group_by(meta.id) %>% mutate(se.new = case_when(outcome.flag == "DICH" ~ sqrt(var.lrr),
                                                  TRUE ~ se)) %>% 
  mutate(ratio = (max(se.new)^2)/(min(se.new)^2),
                                  n.sig = sum(sig.single)) %>% 
  filter(ratio >= 4 & n.sig > 0)
sig.id.count <- length(unique(tp4$id))
sig.study.count <- length(unique(tp4$study.name))
sig.result.count <- dim(tp4)[1]

#Without duplicates:
m.a.no.dupl.number <- data.ext2 %>% filter(meta.id %in% m.a.largerten.id$meta.id) %>% 
  group_by(meta.id) %>% mutate(se.new = case_when(outcome.flag == "DICH" ~ sqrt(var.lrr),
                                                  TRUE ~ se)) %>% 
  summarise(ratio = (max(se.new)^2)/(min(se.new)^2),
                                  n.sig = sum(sig.single),
                                  dupl = unique(dupl.remove)) %>% 
  filter(ratio >= 4 & n.sig > 0 & dupl == 0) %>% count
tp5 <- data.ext2 %>% filter(meta.id %in% m.a.largerten.id$meta.id) %>% 
  group_by(meta.id) %>% mutate(se.new = case_when(outcome.flag == "DICH" ~ sqrt(var.lrr),
                                                  TRUE ~ se)) %>% 
  mutate(ratio = (max(se.new)^2)/(min(se.new)^2),
                                  n.sig = sum(sig.single),
                                  dupl = unique(dupl.remove)) %>% 
  filter(ratio >= 4 & n.sig > 0 & dupl == 0)
dupl.id.count <- length(unique(tp5$id))
dupl.study.count <- length(unique(tp5$study.name))
dupl.result.count <- dim(tp5)[1]

#smaller I2 than 0.5:
m.a.smaller.5.count <- meta.info.I2 %>% filter(I2 < 50) %>% ungroup %>% count()

meta.id.vector <- meta.info.I2 %>% filter(I2 < 50) %>% select(meta.id)
meta.data <- data.ext2 %>% filter(meta.id %in% meta.id.vector$meta.id)

I2.id.count <- length(unique(meta.data$id))
I2.study.count <- length(unique(meta.data$study.name))
I2.result.count <- dim(meta.data)[1]
#----------------------------------------------------------------------------------------------#

#Adjustment dataset:
meta.adj.2 <- meta.f %>% filter(outcome.flag == "IV") %>%
  filter(outcome.measure.merged == "SMD")
meta.adj.1 <- meta.f %>% filter(outcome.flag == "DICH" | outcome.flag == "CONT") 
meta.adj <- bind_rows(meta.adj.2, meta.adj.1)
tp6 <- data.ext2 %>% filter(meta.id %in% meta.adj$meta.id)

adj.id.count <- length(unique(tp6$id))
adj.study.count <- length(unique(tp6$study.name))
adj.result.count <- dim(tp6)[1]

#----------------------------------------------------------------------------------------------#
# missing.copas.histogram <- meta.f %>% ggplot(aes(x = missing.copas/k)) + geom_histogram(binwidth = 0.05, boundary = 0) + theme_bw() + xlab("Missing study fraction")
missing.copas.zeros <- length(which(meta.f$missing.copas[!is.na(meta.f$missing.copas)]/meta.f$k[!is.na(meta.f$missing.copas)] == 0))
missing.copas.range.frac <- c(missing.copas.zeros, quantile(x = meta.f$missing.copas[!is.na(meta.f$missing.copas)]/meta.f$k[!is.na(meta.f$missing.copas)], c(0.05, 0.25, 0.5, 0.75, 0.95)), mean(meta.f$missing.copas[!is.na(meta.f$missing.copas)]/meta.f$k[!is.na(meta.f$missing.copas)],na.rm = T) )
missing.copas.range.count <- c(missing.copas.zeros, quantile(x = meta.f$missing.copas[!is.na(meta.f$missing.copas)], c(0.05, 0.25, 0.5, 0.75, 0.95)), mean(na.rm = T,meta.f$missing.copas[!is.na(meta.f$missing.copas)])) 
missing.copas.table <- rbind(missing.copas.range.frac, missing.copas.range.count)
rownames(missing.copas.table) <- c("Overall fraction", "Missing number")
colnames(missing.copas.table) <- c("= 0", colnames(missing.copas.table)[2:6], "mean")
missing.copas.table <- missing.copas.table[c(2,1),]
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{document}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Publication Bias}
\begin{itemize}
\item Preference of journal editors to publish significant study results
\end{itemize}
\hspace{2mm} $\rightarrow$ \hspace{2mm} non-significant results remain in file-drawer
\end{frame}

\begin{frame}{Systematic Reviews}
\begin{itemize}
\item Summarize all evidence with regard to treatment with meta-analysis
\item Biased if non-significant results are not available and included
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Funnel Plot}
Look for funnel plot asymmetry:

\vspace{-5mm}
\begin{figure}
<<echo = FALSE, fig.height = 4, warning=FALSE, cache = TRUE>>=
ex.se <- meta.f$meta.id[which(meta.f$id == "CD003246" & meta.f$subgroup.nr == 0
                              & meta.f$outcome.nr == 3 & meta.f$comparison.nr == 9)]
ex.nose <- meta.f$meta.id[which(meta.f$id == "CD000213" & meta.f$subgroup.nr == 1
                              & meta.f$outcome.nr == 15 & meta.f$comparison.nr == 1)]
par(mfrow = c(1,2))
funnel(x = meta.id(ex.se), ylim = c(1.60, 0), comb.fixed = F, comb.random = F)
funnel(x = meta.id(ex.nose), ylim = c(1.60, 0), comb.fixed = F, comb.random = F, ylab = NULL)
@
\end{figure}

\end{frame}

\begin{frame}[fragile]{Funnel Plot}
Effects with large standard errors have larger effect sizes (because they are only published
if significant)

\vspace{-10mm}
\begin{figure}
<<echo = FALSE, fig.height = 4, warning=FALSE, cache = TRUE>>=
par(mfrow = c(1,2))
funnel(x = meta.id(ex.se), ylim = c(1.60, 0), comb.fixed = F, comb.random = F, contour.levels = 0.975)
funnel(x = meta.id(ex.nose), ylim = c(1.60, 0), comb.fixed = F, comb.random = F, ylab = NULL)
@
\end{figure}
\end{frame}

\begin{frame}{Detect Publication Bias}
\begin{itemize}
\item Small study effect tests (funnel plot asymmetry)
\item Excess significance tests
\end{itemize}
\end{frame}

\begin{frame}{Excess Significance Test}
Calculate power of each study, given that true effect size is
fixed effects meta-analysis estimate.

Calculate:
\begin{align}
p &= \sum_{i = O}^n\Big({n \choose i} p^i (1-p)^{n - i}\Big) \nonumber
\end{align}

\vspace{-1mm}
$O = $ observed no. of significant results, $E$ expected based on power of studies, $p$ = $E/n$.
\end{frame}

\begin{frame}{Analysis}
\begin{itemize}
\item Use meta-analyses from Cochrane.
\item ``The single most reliable source of evidence in clinical science
\end{itemize}

Analyse meta-analyses with publication bias tests.
\end{frame}

%------------------------------------------------------------------------------------------------%
\begin{frame}{The Cochrane Dataset}
\small
\begin{figure}
\begin{center}
\tikzstyle{block} = [rectangle, draw, fill=blue!20, text centered, rounded corners, minimum height=2.3em]
\tikzstyle{btw} = [rectangle, node distance=0cm, minimum height =1.8em, fill= white]
\tikzstyle{ublock} = [rectangle, draw, fill=blue!20, text centered, rounded corners, minimum height=4em, text width = 10em]
\tikzstyle{line} = [draw, ->, thick]
\tikzstyle{final} = [rectangle, draw, fill=red!20, text centered, rounded corners, minimum height=2.8em]

\begin{tikzpicture}[node distance = 2cm, auto]

        \node [final] (1) {Inital dataset:
    \Sexpr{format(initial.id.count, big.mark = ",")} reviews,
    \Sexpr{format(initial.study.count, big.mark = ",")} studies,
    \Sexpr{format(initial.result.count, big.mark = ",")} results};

    \node [final, below of=1, node distance=2cm] (7) {Analysis dataset: 
    \Sexpr{format(I2.id.count, big.mark = ",")} reviews,
    \Sexpr{format(I2.study.count, big.mark = ",")} studies,
    \Sexpr{format(I2.result.count, big.mark = ",")} results};

    \path [line] (1) -- node [btw, xshift = -2.8cm] {exclusion of unsuitable meta-analyses} (7);
\end{tikzpicture}
\label{Inclusion.criteria}
\end{center}
\end{figure}
\end{frame}

\begin{frame}{The Analysis dataset}
\footnotesize
\begin{figure}
\begin{center}
\tikzstyle{block} = [rectangle, draw, fill=blue!20, text centered, rounded corners, minimum height=2.3em]
\tikzstyle{btw} = [rectangle, node distance=0cm, minimum height =1.8em, fill= white]
\tikzstyle{ublock} = [rectangle, draw, fill=blue!20, text centered, rounded corners, minimum height=4em, text width = 10em]
\tikzstyle{line} = [draw, ->, thick]
\tikzstyle{final} = [rectangle, draw, fill=red!20, text centered, rounded corners, minimum height=2.8em]

\begin{tikzpicture}[node distance = 2cm, auto]

    \node [final, node distance=1cm] (7) {Analysis dataset:
    \Sexpr{format(m.a.smaller.5.count$nn, big.mark = ",")} meta-analyses};

    \node [ublock, below of=7, node distance=3.5cm] (8) {Binary data: \\
    \Sexpr{format(dim(meta.bin)[1], big.mark = ",")} meta-analyses};

    \node [ublock, right of=8, node distance=3.5cm] (9) {Continuous data: \\
    \Sexpr{format(dim(meta.cont)[1], big.mark = ",")} meta-analyses};

    \node [ublock, left of=8, node distance=3.5cm] (10) {Other data (IV): \\
    \Sexpr{format(dim(meta.iv[1])[1], big.mark = ",")} meta-analyses};

    \path [line] (7) -- (8);
    \path [line] (7) --  (9);
    \path [line] (7) --  (10);

\end{tikzpicture}

\end{center}
\end{figure}
\end{frame}

\begin{frame}{Small Study Effect Tests}
\vspace{-4mm}
Weighted linear regression with std. error $x_i$ and effect size $y_i$:
\begin{align}
y_i &= \beta_0 + \beta_1 x_i + \epsilon_i, & \epsilon_i \sim N(0, x_i \sigma^2) \nonumber
\end{align}
Test for $H0: \beta_1 = 0$, no funnel plot asymmetry
\end{frame}

\begin{frame}{Adjustments for Binary Data}
As recommended by \citet{Sterne}
\begin{itemize}
\item Log odds ratio and risk ratio $\theta$ and standard error $\se_\theta$ are not independent
\item Use score of binomial likelihood at log odds ratio $\theta_{\textrm{H0}} = 0$ instead of 
log odds ratio, and the inverse Fisher information instead of $\se_\theta$
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Small Study Effect Tests}
\begin{figure}
<<echo = FALSE, warnings = FALSE, cache = TRUE, fig.height = 4>>=
dat_text <- data.frame(
  label = paste(sum(meta.bin$harbord.test), "< 0.1,",
                round(mean(meta.bin$harbord.test),2)*100, "%"))
p.dist.bin <- meta.bin%>% 
  ggplot(aes(x = pval1.harbord)) + geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
  geom_text(data = dat_text, mapping = aes(x = 0.5, y = 125, label = label), 
  					color = "red", size = 4.5) + 
  ggtitle("Binary data (Harbord)") + 
	xlab(expression(paste(italic(p), "-value"))) + theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 12, margin = margin(b = 20)),
        axis.title.y = element_text(size=16, margin = margin(r = 10)), 
        axis.title.x = element_text(size=16, margin = margin(t = 10)), 
        axis.text= element_text(size=12),
        strip.text.x = element_text(size = 30))

dat_text <- data.frame(
  label = paste(sum(meta.cont$egger.test), "< 0.1,", round(mean(meta.cont$egger.test),2)*100, "%"))

p.dist.cont <- meta.cont %>% 
  ggplot(aes(x = pval1.egger)) + geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
  geom_text(data = dat_text, mapping = aes(x = 0.5, y = 63, label = label),  
  					color = "red", size = 4.5) + 
  ggtitle("Cont. Data (Egger)") + ylab("") + 
	xlab(expression(paste(italic(p), "-value"))) + theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 12, margin = margin(b = 20)),
        #axis.title.y = element_text(size=16, margin = margin(r = 10)), 
        axis.title.x = element_text(size=16, margin = margin(t = 10)), 
        axis.text= element_text(size=12),
        strip.text.x = element_text(size = 30))

dat_text <- data.frame(
  label = paste(sum(meta.iv$egger.test), "< 0.1,", round(mean(meta.iv$egger.test),2)*100, "%"))

p.dist.iv <- meta.iv %>%  ggplot(aes(x = pval1.egger)) + 
	geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
  geom_text(data = dat_text, mapping = aes(x = 0.5, y = 24.5, label = label),  
  					color = "red", size = 4.5) + 
  ggtitle("Other Data (Egger)") + ylab("") +
	xlab(expression(paste(italic(p), "-value")))  + theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 12, margin = margin(b = 20)),
        axis.title.x = element_text(size=16, margin = margin(t = 10)), 
        axis.text= element_text(size=12),
        strip.text.x = element_text(size = 30))

grid.arrange(p.dist.bin, p.dist.cont, p.dist.iv, ncol = 3)
@
\end{figure}
\end{frame}

\begin{frame}[fragile]{Excess Significance Test}
\begin{figure}
<<echo = FALSE, cache = TRUE, fig.height = 4>>=
dat_text <- data.frame(
  label = paste(sum(meta.f$tes.d.test), "< 0.1,", round(mean(meta.iv$tes.d.test),2)*100, "%"))

meta.f %>% ggplot(aes(x = pval1.d.tes)) +  
	geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
  ggtitle("All Data") +
  geom_text(data = dat_text, mapping = aes(x = 0.5, y = 100, label = label),  
  					color = "red", size = 5.5) + 
	xlab(expression(paste(italic(p), "-value")))  + theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 16, margin = margin(b = 20)),
        axis.title.y = element_text(size=16, margin = margin(r = 10)), 
        axis.title.x = element_text(size=16, margin = margin(t = 10)), 
        axis.text= element_text(size=12),
        strip.text.x = element_text(size = 30))

@
\end{figure}
\end{frame}

\begin{frame}{Adjustment}
Calculate unbiased estimates by
\begin{itemize}
\item Sensitivity analysis
\item Regression approach
\end{itemize}
\end{frame}

\begin{frame}{Selection Model}
\citet{Copas2}
\begin{align}
\theta_i &\sim N(\mu_i, \sigma_i^2) &
\mu_i \sim N(\theta, \tau^2) \nonumber
\end{align}
So we have $\theta_i = \mu_i + \sigma_i\epsilon_i \nonumber$. Introduce
\begin{itemize}
\item Baseline study retention rate $a$
\item Std. error dependent retention rate $b$
\item introduces correlation between $\epsilon_i$ and selection probability
\end{itemize}
%Model the selection process for increasing $a$ and $b$
\end{frame}

\begin{frame}{Sensitivity analysis}
To test a pair $a$,$b$,	include $\beta$:
\begin{align}
\theta_i &= \theta + \beta \se_i + \sigma_i \epsilon_i \nonumber
\end{align}
\vspace{-10mm}
\begin{itemize}
\item $\theta_i$ is the effect size of study $i$ and right-hand side is the fitted 
value of the model. 
\item Repeatedly test for $\beta_{\textrm{H0}} = 0$ %and estimate $\theta_{\textrm{Adj.}}$
\end{itemize}
\end{frame}

\begin{frame}{Regression}
Correct for small study effect \citep{Limitmeta}:
\begin{itemize}
\item Use global mean of regression $\beta_0$ from $y_i = \beta_0 + \beta_1 x_i + \epsilon_i, \epsilon_i \sim N(0, x_i \sigma^2) \nonumber$
\item Add minimal bias term $\beta_1 \cdot \tau$
\end{itemize}
$\theta_{\textrm{Adj.}} = \beta_0 + \beta_1\tau$
\end{frame}

\begin{frame}{Adjustment Methods}
Differences:
\begin{itemize}
\item Regression: uncertainty of bias parameter retained
\item Sensitivity analysis: Bias itself is not estimated, but most parsimonious scenario is chosen
\item Sensitivity analysis: No adjustment if no evidence for bias
\end{itemize}
Regression adjustment leads to higher uncertainty of $\theta_{\textrm{Adj.}}$
\end{frame}

\begin{frame}{Effect Measure Transformation}
\vspace{-3mm}
\begin{align}
d &= \theta \frac{\sqrt{3}}{\pi} & \se_d^2 =  \se^2\frac{\sqrt{3}}{\pi} \nonumber
\end{align}
\begin{align}
r &= \frac{d}{\sqrt{d^2 + a}} & a = (n_c + n_t)^2 / n_c n_t \nonumber
\end{align}
\begin{align}
z &= 0.5 \ln\bigg(\frac{1 + r}{1 - r}\bigg) & \nonumber
%z &= \arctan(r) & \nonumber
\se_z^2 &= \frac{1}{n-3}
\end{align}
\end{frame}

\begin{frame}[fragile]{Results}
<<echo = FALSE, cache = TRUE, fig.height = 4, message = FALSE, warning = FALSE>>=
diff.z.ranef.reg <- meta.f %>% 
  mutate(fixef = est.z.fixef, ranef = est.z.ranef,
         copas = est.z.copas, regression = est.z.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.z.score", regression) %>% 
  mutate(difference = ranef - adjusted.z.score) %>% 
  filter(difference > -.75 & difference < .75) %>% 
  ggplot(aes(x = difference)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + ggtitle("Regression") +
  xlab(expression(paste("Random effects ", italic(hat(z)), " - adjusted ", italic(hat(z))))) + 
	geom_vline(xintercept = 0, linetype = "dashed" ) + theme(
		axis.title.y = element_text(size=16, margin = margin(r = 10)), 
        plot.title = element_text(hjust = 0.5, face = "bold", size = 15, margin = margin(b = 22)),
        axis.title.x = element_text(size=14, margin = margin(t = 10)), 
        axis.text= element_text(size=12),
        strip.text.x = element_text(size = 30)) + xlim(c(-0.8, 0.7))
#--------------------------------------------------------------------------------------------------------------------#

diff.z.ranef.copas <- meta.f %>% 
  mutate(fixef = est.z.fixef, ranef = est.z.ranef,
         copas = est.z.copas, regression = est.z.reg) %>% 
  mutate(copas = case_when(sign(copas) == sign(ranef) ~  abs(copas),
                           sign(copas) != sign(ranef) ~ -abs(copas)),
         regression = case_when(sign(regression) == sign(ranef) ~  abs(regression),
                                sign(regression) != sign(ranef) ~ -abs(regression)),
         fixef = abs(fixef),
         ranef = abs(ranef)) %>% 
  select(fixef, ranef, copas, regression) %>% 
  gather(key = "method", value = "adjusted.z.score", copas) %>% 
  mutate(difference = ranef - adjusted.z.score) %>% 
  filter(difference > -.75 & difference < .75) %>% 
  ggplot(aes(x = difference)) + geom_histogram(binwidth = 0.1, center = 0) +
  theme_bw() + ylab("") + ggtitle("Selection Model") +
  xlab(expression(paste("Random effects ", italic(hat(z)), " - adjusted ", italic(hat(z))))) + 
  	geom_vline(xintercept = 0, linetype = "dashed" ) + theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 15, margin = margin(b = 22)),
        axis.title.x = element_text(size=14, margin = margin(t = 10)), 
        axis.text= element_text(size=12),
        strip.text.x = element_text(size = 30)) + xlim(c(-0.8, 0.7))

grid.arrange(diff.z.ranef.reg, diff.z.ranef.copas, ncol = 2)
@
\end{frame}

\begin{frame}{Results}
% \vspace{-3mm}
% Compare wald test statistics (original effect size measures):
% 
% \vspace{-2mm}
<<echo = FALSE, warning = FALSE, cache = TRUE, fig.height = 4>>=
mid <- median(meta.f$pval.se)
method.names <- c(z.reg = "Regression", z.copas = "Selection Model")

decrease_text <- data.frame(label = c("decrease", "decrease"),
  method   = c("z.reg", "z.copas"))
increase_text <- data.frame(label = c("increase", "increase"),
                            method   = c("z.reg", "z.copas"))
meta.f %>% filter(z.reg > -40) %>% 
  select(pval.se, z.fixef, z.ranef, z.reg, z.copas) %>% 
  gather(key = "method", value = "test.stat", z.reg:z.copas) %>%
  ggplot(aes(x = test.stat, y = z.ranef, colour = pval.se)) + 
  facet_wrap(~ method, labeller = as_labeller(method.names)) + 
  scale_color_gradient2(midpoint=mid, low="firebrick4", mid = "gold",
                        high="chartreuse2", space ="Lab", name = expression(paste(italic(p),"-value"))) +
  geom_point(size = .6) + theme_bw() + guides(fill=guide_legend(title="p-value")) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, size = .3) +
  geom_abline(slope = 0, intercept = 0, linetype = 2, size = .3) +
  geom_text(data = decrease_text, mapping = aes(x = -11, y = 6, label = label), 
            color = "tan3") + 
  geom_text(data = decrease_text, mapping = aes(x = 11, y = -6, label = label), 
            color = "tan3") + 
  geom_text(data = increase_text, mapping = aes(x = 15, y = 4, label = label), 
            color = "chartreuse3") + 
  geom_text(data = increase_text, mapping = aes(x = -15, y = -4, label = label), 
            color = "chartreuse3") + 
  xlab(expression(paste("Adjusted ",  italic(z),  " test statistic"), sep = "")) + 
  ylab(expression(paste("Random effects z ", italic(z), " test statistic"), sep = "")) +
  theme(legend.title = element_text(size=14),
        axis.title.y = element_text(size=14, margin = margin(r = 20)), 
        axis.title.x = element_text(size=14, margin = margin(t = 20)), 
        legend.text = element_text(size =10), 
        axis.text= element_text(size=12),
        strip.text.x = element_text(size = 15),
        #plot.title = element_text(hjust = 0.5, face = "bold", size = 40, margin = margin(b = 20)),
  			strip.background = element_rect(fill = "white"))
@

\end{frame}

\begin{frame}{Missing Study Number}
Missing study number according to selection model parameters: \Sexpr{format(round(sum(meta.f$missing.copas, na.rm = T)),big.mark = ",")}, \Sexpr{round(sum(meta.f$missing.copas, na.rm = T)/sum(meta.f$k),3)*100}\%

<<echo= FALSE, results='asis', cache = TRUE>>=
print(xtable(missing.copas.table, caption = "Fraction of missing studies and estimates of missing studies with their zero counts (``= 0''), quantiles and means.",
             label = "copas.missing", align = "lrrrrrrr", digits = c(0,0,0,0,1,1,1,1)),
      size = "footnotesize")
@
\end{frame}

\begin{frame}{Additional Results}
Linear mixed model with random effects for meta-analyses and reviews:

\begin{align}
\mathbf{y}_i \given U_j,U_k,\epsilon_{i} =  \beta_0 + x_i\beta_1 + U_j + U_k + \epsilon_{i} \nonumber
\end{align}

$\rightarrow$ Increased power to estimate $\beta$
\end{frame}

\begin{frame}{Mixed Linear Model}
AIC improvement to null fit: 1820 to 1306. \\
Bias parameter $\beta_1$ = 0.52 (95\%CI: 0.44,0.59)

\vspace{-12mm}
\begin{figure}
<<echo = FALSE, warning=FALSE, cache = TRUE, fig.height = 4>>=
reg.dat <- data.ext2 %>% filter(meta.id %in% meta.f$meta.id) #Filter meta-analysis data
reg.dat <- reg.dat %>% group_by(meta.id) %>%
  filter(var.z != 0) %>% filter(!is.na(z)) %>% filter(!is.na(var.z)) %>%
  mutate(se.z.mirrored = sqrt(var.z)) %>% filter(!is.na(se.z.mirrored)) %>%
  mutate(bias.side.z = bias.side.z(yi = z, sei = se.z.mirrored),
  z.mirrored = z * bias.side.z) #Mirror effect sizes on one side
reg.dat <- reg.dat %>%  filter(!is.na(z.mirrored)) %>%
  filter(!is.na(se.z.mirrored))
reg.dat <- reg.dat %>% filter(se.z.mirrored < 3)
publication.bias.fit <- lme(fixed = z.mirrored ~ se.z.mirrored, 
														random = list(~ 1 | id, ~ 1 | meta.id), 
														weights = ~ se.z.mirrored^2, method = "ML", data = reg.dat)
publication.bias.rs.fit <- lme(fixed = z.mirrored ~ se.z.mirrored,
                          random = list(~ 1 | id, ~ meta.id | meta.id),
                       weights = ~ se.z.mirrored^2, method = "ML", data = reg.dat)
plot(y = reg.dat$z.mirrored, x = reg.dat$se.z.mirrored, cex = .1, cex.lab = 1.3,
		 xlab = "standard error", ylab = "z mirrored")
publication.bias.fit.coef <- summary(publication.bias.fit)$coefficients$fixed
abline(a = publication.bias.fit.coef, col = "red", lty = 2, lwd = 1.5)
publication.bias.rs.fit.coef <- summary(publication.bias.rs.fit)$coefficients$fixed
abline(a = publication.bias.rs.fit.coef, col = "blue", lty = 3, lwd = 1.5)
@
\end{figure}
\end{frame}

\begin{frame}{Limitations}
Small study effect $\neq$ publication bias. Also:
\begin{itemize}
\item True heterogeneity
\item Selective outcome reporting
\item Delayed publication
\item ...
\end{itemize}
\vspace{-3mm}
Small study effect tests do not take into account significance directly.
\end{frame}

\begin{frame}{Limitations}
\begin{itemize}
\item Unknown amount of secondary outcomes
\item Small amount of adverse outcomes
\item Publication bias unknown for excluded data
\item Exploratory Analysis!
\end{itemize}
\end{frame}

\begin{frame}{Implications}
\begin{itemize}
\item Results are largely in line with previous research
\item Underline the need for examination of publication and other biases in meta-analyses
\item Otherwise, validity of meta-analyses can be contested
\end{itemize}
\end{frame}

\begin{frame}{References}
  \small
  \bibliographystyle{apalike}
\bibliography{illustration}
\end{frame}

\begin{frame}
Backup slides:
\end{frame}

\begin{frame}[fragile]{Transformation}

\vspace{-6mm}
<<echo = FALSE, warning=FALSE, fig.height=4.8, cache = TRUE>>=
par(mfrow = c(2,3), cex.lab = 1.5, cex.axis = 1.5)
example <- data.ext2 %>% filter(meta.id == 81244)
funnel(limitmeta(metagen(TE = effect, se = se, studlab = study.name, sm = "MD", data = example)), col.adjust = "blue", col.line = "blue")
funnel(limitmeta(metagen(TE = smd.pool, se = se.smd.pool, studlab = study.name, sm = "SMD", data = example)), col.adjust = "blue", col.line = "blue")
funnel(limitmeta(metagen(TE = z, seTE = sqrt(var.z), studlab = study.name, sm = "ZCOR", data = example)), col.adjust = "blue", col.line = "blue")

example <- data.ext2 %>% filter(meta.id == 32532)
funnel(limitmeta(metagen(TE = log(effect), se = se, studlab = study.name, sm = "RR", data = example)), col.adjust = "blue", col.line = "blue")
funnel(limitmeta(metagen(TE = smd.pool, se = se.smd.pool, studlab = study.name, sm = "SMD", data = example)), col.adjust = "blue", col.line = "blue")
funnel(limitmeta(metagen(TE = z, seTE = sqrt(var.z), studlab = study.name, sm = "ZCOR", data = example)), col.adjust = "blue", col.line = "blue")
@
\end{frame}

% 
% \begin{frame}{Fixed Effects Meta-Analysis}
% Study $i$ of $n$ studies, effects $\theta_i$ and variances $v_i$, s.e. $s_i$
% 
% Fixed effects estimate: weighted mean with $w_i = 1/\hat{v_i}$: 
% 
% \begin{align}
% \theta_f &= \frac{\sum_{i = 1}^n w_{i}\theta_i}{\sum_{i = 1}^n w_i} \nonumber \\ 
% \textrm{se}(\theta_f) &= \frac{1}{\sqrt{\sum_{i = 1}^n w_i}} \nonumber
% \end{align}
% \end{frame}
% 
% \begin{frame}{Random Effects Meta-Analysis}
% Let $\hat{\theta_i}|\theta_i \sim N(\theta_i, v_i) 
% \theta_i \sim N(\theta, \tau^2)$
% 
% Random effects estimate: weigthed mean of $\hat{\theta_i}$ with weights $w_i = 1/(\hat{v_i} + \tau^2)$
% \end{frame}
% 
% \begin{frame}{Evidence Based Medicine}
% Likely to base decisions on over-optimistic findings
% 
% Direct consequences: patient harm and unnecessary research
% \end{frame}
% 
% 
% \begin{frame}{Cochrane Organisation}
% Aim: summarise findings in primary clinical research and health care
% 
% Provide peer-reviewed, systematic reviews
% 
% Public access (for some countries)
% \end{frame}
% 
% \begin{frame}{Research Question}
% Quantify the abundance and impact of publication bias in the Cochrane Library
% \end{frame}
% 
% 
% \begin{frame}{Cochrane Library Dataset}
% \Sexpr{format(length(unique(data$file.nr)), big.mark=",")} systematic reviews with studies published until 2018.
% 
% \Sexpr{format(length(unique(data$study.name)), big.mark=",")} studies.
% 
% \Sexpr{format(dim(data)[1], big.mark=",")} study results.
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Review Example: Binary Outcome}
% Barbiturate efficacy for head injury treatment
% \vspace{-5mm}
% <<echo = FALSE, results = 'asis'>>=
% print(xtable(barbi2, label = "barbiturates", digits = 0), include.rownames = F, size = "tiny")
% @
% \end{frame}
% 
% 
% 
% \begin{frame}{Dataset Structure}
% 
% \begin{figure}
% \tikzstyle{every node}=[draw=black,thick,anchor=west,scale=.65]
% \tikzstyle{selected}=[draw=red,fill=red!30]
% \tikzstyle{optional}=[dashed,fill=gray!50]
% \begin{tikzpicture}
% [grow = right, anchor = west,
%   growth parent anchor=east, % added code
%   parent anchor=east, level distance=.5cm,
%   sibling distance=2em, level 1/.style={sibling distance=2em}, level 2/.style={sibling distance=2em},
%   level 3/.style={sibling distance=2em}, level 4/.style={sibling distance=1.2em}]
%   \node {Review} [edge from parent fork right]
%     child { node {Comparison 2}
%       child { node {Outcome 2}}
%       child { node {Outcome 1}
%         child { node {Subgroup 2}}
%         child { node {Subgroup 1}
%           child  { node {Result 3}}
%           child  { node {Result 2}}
%           child  { node {Result 1}}
%           }}
%     }
%     child [missing] {}
%     child { node {Comparison 1}};
% \end{tikzpicture}
% %\caption{Structure of a hypothetical review with two different comparisons\label{review.structure}}
% \label{review.structure}
% \end{figure}
% \end{frame}
% 
% 
% 
% 
% 
% \begin{frame}[fragile]{Dataset Properties}
% Review or study level:
% <<echo=FALSE, results = 'asis'>>=
% print(xtable(dataset.properties, digits = 0, align = "lrrrr"), include.rownames = T, size = "footnotesize", hline = c(0,1,1,3))
% @
% \end{frame}
% 
% 
% 
% \begin{frame}{Small Study Effects}
% ``The tendency for the smaller studies to show larger treatment effects'' \citep{Sterne}
% \end{frame}
% 
% 
% \begin{frame}{Small Study Effects}
% Causes:
% \begin{itemize}
% \item Selective publication of studies with significant results - publication bias
% \item Systematic differences in study settings
% \end{itemize}
% \end{frame}
% 
% 
% \begin{frame}{Small Study Effect Tests}
% Different approaches:
% \begin{itemize}
% \item Simple linear regression
% \item Rank correlation
% \end{itemize}
% 
% Special methods for binary outcomes	
% \end{frame}
% 
% 
% % \begin{frame}[fragile]{Small Study Effect Tests}
% % Funnel plots (continuous outcome examples):
% % 
% % \vspace{-1.2cm}
% % 
% % <<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
% % par(las = 1, mfrow = c(1, 2))
% % funnel(meta.example)
% % funnel(meta.nexample)
% % @
% % 
% % \end{frame}
% % 
% % 
% % \begin{frame}[fragile]{Regression based Tests}
% % Radial plots (continuous outcome examples):
% % 
% % \vspace{-1.1cm}
% % 
% % <<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
% % par(las = 1, mfrow = c(1,2))
% % biased.rev$inv <- 1/biased.rev$se
% % plot(y = (biased.rev$effect/biased.rev$se), x = (biased.rev$se^-1), xlim = c(0, 6), xlab = "inverse standard error", 
% % 		 ylab = "mean diff. / std. error")
% % 
% % unbiased.rev$inv <- 1/unbiased.rev$se
% % plot(y = (unbiased.rev$effect/unbiased.rev$se), x = (unbiased.rev$inv), xlim = c(0, 1.2), 
% % 		 ylim = c(-18, 0), xlab = "inverse standard error", ylab = "mean diff. / std. error")
% % @
% % 
% % \end{frame}
% 
% 
% \begin{frame}{Regression based Tests}
% study $i$ of $n$ studies, effects $\theta_i$ and variances $v_i$, s.e. $s_i$
% 
% $\theta_M$ is the pooled effect and $\tau^2$ the between-study variance.
% 
% Let $y_{i} = \theta_{i}/_{i}$ and $x_i = 1/s_i$
% \begin{itemize}
% \item \citet{Egger} : Simple linear regression \\ $y_i = \beta_0 + \beta_1 x_i, \epsilon_i \sim N(0, \sigma)$
% \item \citet{thompson.sharp} : extension of Egger with study weights $v_{i} + \tau^2$ 
% \end{itemize}
% 
% \end{frame}
% 
% \begin{frame}[fragile]{Egger's Test examples}
% Test for non-zero intercept $\beta_{0}$
% 
% \vspace{-1.1cm}
% <<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
% par(las = 1, mfrow = c(1,2))
% biased.rev$inv <- 1/biased.rev$se
% m.ex <- lm(formula = (biased.rev$effect/biased.rev$se) ~ (biased.rev$inv))
% plot(y = (biased.rev$effect/biased.rev$se), x = (biased.rev$se^-1), xlim = c(0, 6), xlab = "inverse standard error", 
% 		 ylab = "mean diff. / std. error")
% abline(coef = coef(m.ex), lty = 2, col = 2)
% 
% 
% unbiased.rev$inv <- 1/unbiased.rev$se
% m.nex <- lm(formula = (unbiased.rev$effect/unbiased.rev$se) ~ (unbiased.rev$inv))
% plot(y = (unbiased.rev$effect/unbiased.rev$se), x = (unbiased.rev$inv), xlim = c(0, 1.2), 
% 		 ylim = c(-18, 0), xlab = "inverse standard error", ylab = "mean diff. / std. error")
% abline(coef = coef(m.nex), lty = 2, col = 2)
% @
% \end{frame}
% 
% 
% \begin{frame}{Regression Tests for Binary Outcomes}
% 
% \begin{itemize}
% \item \citet{Peters} :$x_i = 1/n_i$ instead $1/s_i$, inverse variances as weight.
% \item \citet{Harbord} :$x_i$ = score of the log-likelihood of a proportion and inverse variances as weights.
% \item \citet{Rucker} :Use arcsine variance stabilizing transformation for variances and effects, do e.g. Egger's test.
% \end{itemize}
% \end{frame}
% 
% 
% \begin{frame}{Rank based tests}
% \citet{begg.ties}: \\
% Let $y_{i}$ be $\frac{\theta_i - \theta_M}{v_i}$ and $x_i$ its variance ($\neq v_i$)
% 
% $u$ the number of pairs $(y_{i}, x_{i})$ ranked in the same order, $l$
% the number of pairs in the opposite order
% 
% $Z = \frac{(u - l)}{\sqrt{n(n-1)(2n + 5)/18}}$  is a test statistic
% \end{frame}
% 
% \begin{frame}{Rank based tests}
% \citet{Schwarzer}: \\
% $e_t$ number of events in the treatment group
% 
% $E_t$ follows hypergeometric distribution: calculate $\mathbb{E}(E_{t})$ and variances
% 
% proceed as in \citet{begg.ties}
% \end{frame}
% 
% 
% \begin{frame}{Test Results}
% Inclusion criteria (from \citet{Ioannidis2007}):
% \begin{itemize}
% \item $n \geq 10$
% \item at least one statistically significant effect in a study
% \item $\frac{\hat{v_{\textrm{max}}}^2}{\hat{v_{\textrm{min}}}^2} > 4$
% \item $I^2 < 0.5$
% \end{itemize}
% 
% From \Sexpr{dim(meta)[1]} with $n \geq 10$, \Sexpr{dim(meta.f)[1]} remain.
% \end{frame}
% 
% 
% 
% \begin{frame}[fragile]{Continuous Outcome Test Results}
% $p$-values distribution, $n$ = \Sexpr{dim(meta.cont)[1]}:
% 
% \vspace{-2mm}
% <<echo = FALSE, fig.height = 4, message = FALSE>>=
% plot(p.dist.cont)
% @
% \end{frame}
% 
% \begin{frame}[fragile]{Binary Outcome Test Results}
% $p$-values distribution, $n$ = \Sexpr{dim(meta.bin)[1]}:
% 
% \vspace{-2mm}
% <<echo = FALSE, fig.height = 3.6, message = FALSE>>=
% plot(p.dist.bin)
% @
% \end{frame}
% 
% \begin{frame}[fragile]{Agreement in significance}
% Number of significant test results per meta-analysis:
% 
% \vspace{-2mm}
% <<echo = FALSE, fig.height = 3.2, message = FALSE>>=
% grid.arrange(agree.cont, agree.bin, ncol = 2)
% @
% \end{frame}
% 
% \begin{frame}{Small Study Effect Adjustment}
% Three methods:
% \begin{itemize}
% \item Regression
% \item Copas selection model
% \item Trim-and-fill
% \end{itemize}
% \end{frame}
% 
% \begin{frame}{Adjustment by regression}
% $y_i = \theta_i/s_i, x_i = 1/s_i$
% 
% $y_i = \beta_0 + \beta_1 x_i, \epsilon_i \sim N(0, \sigma)$
% 
% $\beta_1$ is the weighted mean treatment effect if $\beta_0 = 0$
% \end{frame}
% 
% 
% \begin{frame}{Adjustment by regression}
% Radial plots (continuous outcome examples):
% 
% \vspace{-1.1cm}
% 
% <<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
% par(las = 1, mfrow = c(1,2))
% biased.rev$inv <- 1/biased.rev$se
% m.ex.0 <- lm(formula = (biased.rev$effect/biased.rev$se) ~ 0 + (biased.rev$inv))
% m.ex <- lm(formula = (biased.rev$effect/biased.rev$se) ~ (biased.rev$inv))
% plot(y = (biased.rev$effect/biased.rev$se), x = (biased.rev$se^-1), xlim = c(0, 6), xlab = "inverse standard error", 
% 		 ylab = "mean diff. / std. error")
% abline(coef = c(0,coef(m.ex.0)), lty = 2)
% abline(coef = coef(m.ex), lty = 2, col = 2)
% 
% 
% unbiased.rev$inv <- 1/unbiased.rev$se
% m.nex.0 <- lm(formula = (unbiased.rev$effect/unbiased.rev$se) ~ 0 + (unbiased.rev$inv))
% m.nex <- lm(formula = (unbiased.rev$effect/unbiased.rev$se) ~ (unbiased.rev$inv))
% plot(y = (unbiased.rev$effect/unbiased.rev$se), x = (unbiased.rev$inv), xlim = c(0, 1), 
% 		 ylim = c(-18, 0), xlab = "inverse standard error", ylab = "mean diff. / std. error")
% abline(coef = c(0,coef(m.nex.0)), lty = 2)
% abline(coef = coef(m.nex), lty = 2, col = 2)
% @
% \end{frame}
% 
% 
% 
% \begin{frame}[fragile]{Limit Meta-Analysis}
% Extended random effects model:
% 
% \vspace{-4mm}
% \begin{align}
% \theta_i = \beta_0 + \beta_1(\sqrt{v_i + \tau^2}) + \epsilon_i(\sqrt{v_i + \tau^2}), \nonumber \\
% \epsilon_{i} \stackrel{\textrm{iid}}{\sim} N(0,1) \nonumber
% \end{align}
% 
% Use $\mathbb{E}(\theta_{i}) \rightarrow \beta_{0} + \beta_{1}\tau$ for $\sqrt{v_{i}} \rightarrow 0$
% as corrected treatment efffect.
% \end{frame}
% 
% 
% 
% \begin{frame}[fragile]{Limit Meta-Analysis}
% Funnel plot with effect with infinite precision:
% 
% \vspace{-1.1cm}
% <<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
% par(las = 1, mfrow = c(1,2))
% funnel(limitmeta(meta.example))
% funnel(limitmeta(meta.nexample))
% @
% \end{frame}
% 
% 
% 
% \begin{frame}{Selection model}
% \citet{Copas1}: model based on a bivariate normal distribution:
% 
% \vspace{-8mm}
% \begin{align}
% \hat{\theta_i} = \theta_{i} + \sigma_i\epsilon_i \label{population.model2} \\
% \theta_i \sim N(\theta, \tau^2) \label{population.model} \\
% z_i = a + b/\hat{s_i} + \delta_i \label{selection.model}
% \end{align}
% 
% \ref{population.model2},\ref{population.model} is called population model, \ref{selection.model} the selection model
% 
% $(\epsilon_i, \delta_i)$ are standard normal residuals with correlation $\rho = cor(y_i, z_i)$.
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Sensitivity Analysis}
% Model the selection process with different $a,b$
% 
% Test if small study effect is significant, by including \\ $\theta_i = \theta_i + \beta s_i + v_{i}\epsilon_i$
% 
% Estimation: Select $a, b$ such that $H0$ can not be rejected and estimated
% number of unpublished studies is minimal.
% \end{frame}
% 
% 
% \begin{frame}{Trim-and-Fill}
% Mirror studies that cause asymmetry:
% \vspace{-1.2cm}
% 
% <<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
% par(las = 1, mfrow = c(1, 2))
% funnel(meta.example)
% funnel(meta.nexample)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Results:}
% Difference between random and fixed effects meta-analysis estimate: $|\theta_f - \theta_{r}|$
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 3.6, message = FALSE, warning=FALSE>>=
% plot(hist.ranef)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Trim-and-fill}
% Absolute difference between adjusted and fixed effects meta-analysis estimate: $|{\theta_f} - {\theta_{\textrm{adjusted}}}|$
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 3.6, message = FALSE, warning=FALSE>>=
% plot(hist.trimfill1)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Copas}
% Absolute difference between adjusted and fixed effects meta-analysis estimate: $|{\theta_f} - {\theta_{\textrm{adjusted}}}|$
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 3.6, message = FALSE, warning=FALSE>>=
% plot(hist.copas1)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Regression}
% Absolute difference between adjusted and fixed effects meta-analysis estimate: $|{\theta_f} - {\theta_{\textrm{adjusted}}}|$
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 3.6, message = FALSE, warning=FALSE>>=
% plot(hist.reg1)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Results:}
% Random and fixed effects meta-analyses test statistics:
% 
% \vspace{-3mm} 
% <<echo = FALSE, fig.height = 3.8, message = FALSE, warning=FALSE>>=
% plot(ranef.fixef.sc.zval)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Trim-and-fill}
% Adjusted and meta-analysis test statistics:
% 
% \vspace{-3mm} 
% <<echo = FALSE, fig.height = 3.8, message = FALSE, warning=FALSE>>=
% grid.arrange(trimfill.fixef.sc.zval, trimfill.ranef.sc.zval, ncol = 2)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Copas}
% Adjusted and meta-analysis test statistics:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 3.8, message = FALSE, warning=FALSE>>=
% grid.arrange(copas.fixef.sc.zval, copas.ranef.sc.zval, ncol = 2)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Regression}
% Adjusted and meta-analysis test statistics:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 3.8, message = FALSE, warning=FALSE>>=
% grid.arrange(reg.fixef.sc.zval, reg.ranef.sc.zval, ncol = 2)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results}
% Missing study proportions:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 3.8, message = FALSE, warning=FALSE>>=
% grid.arrange(p.missing.copas, p.missing.trim, ncol = 2) 
% @
% \end{frame}
% 
% \begin{frame}[fragile]{Extreme Results}
% RR reduction by trimfill (-3.9), side effects
% 
% <<echo = FALSE, fig.height = 3.5>>=
% trimfill.id(165815)
% @
% \end{frame}
% 
% \begin{frame}[fragile]{Extreme Results}
% RR Reduction by copas selection model (-4), pain relief
% 
% <<echo = FALSE, fig.height = 3.5>>=
% funnel.id(73169)
% @
% \end{frame}
% 
% \begin{frame}[fragile]{Extreme Results}
% RR Amplification by regression (+14), side effects
% 
% <<echo = FALSE, fig.height = 3.5, warning=FALSE>>=
% limitmeta.id(119537)
% @
% \end{frame}
% 
% \begin{frame}[fragile]{Discussion}
% \begin{itemize}
% \item Proportion of positive tests is well above 10\%
% \item Effect sizes and evidence for treatment effect is diminishued
% \item Limitations: not only primary outcomes, adjustment methods known to
% perform poorly under the 0
% \end{itemize}
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Outlook}
% \begin{itemize}
% \item Connect results with different medical fields, look for differences
% \item Connect results with single studies and journals (?)
% \end{itemize}
% \end{frame}
% 
% 
% 
% 
% 
% 
% 
% 
% 
% 
% 
% 
% 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %Backup Slides%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% \begin{frame}[fragile]{Adjustment Results: Trim-and-fill}
% Treatment effect difference:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 4, message = FALSE, warning=FALSE>>=
% plot(hist.trimfill)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Copas}
% Treatment effect difference:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 4, message = FALSE, warning=FALSE>>=
% plot(hist.copas)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Regression}
% Treatment effect difference:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 4, message = FALSE, warning=FALSE>>=
% plot(hist.reg)
% @
% \end{frame}
% 
% \begin{frame}[fragile]{Results}
% log treatment effect estimates:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 4, message = FALSE, warning=FALSE>>=
% plot(ranef.fixef.sc.effect)
% @
% \end{frame}
% 
% \begin{frame}[fragile]{Adjustment Results: Trim-and-fill}
% log treatment effect estimates:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 4, message = FALSE, warning=FALSE>>=
% grid.arrange(trimfill.fixef.sc.effect, trimfill.ranef.sc.effect, ncol = 2)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Copas}
% log treatment effect estimates:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 4, message = FALSE, warning=FALSE>>=
% grid.arrange(copas.fixef.sc.effect, copas.ranef.sc.effect, ncol = 2)
% @
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Adjustment Results: Regression}
% log treatment effect estimates:
% 
% \vspace{-3mm}
% <<echo = FALSE, fig.height = 4, message = FALSE, warning=FALSE>>=
% grid.arrange(reg.fixef.sc.effect, reg.ranef.sc.effect, ncol = 2)
% @
% \end{frame}
% 
% 
% 
%

%\appendix
%% Possible backup slides...

%% chapter division is accomplished with:
%% \part{Appendix}

% \begin{frame}[fragile]{Test Results: Significance}
% 
% <<echo = FALSE, fig.height = 4>>=
% grid.arrange(p.bin, p.cont, ncol = 2)
% @
% \end{frame}

% \begin{frame}[fragile]{Pooling Studies - Meta-Analysis}
% Multiple results in a meta-analysis group can be pooled:
% <<echo=FALSE, results = 'asis'>>=
% print(xtable(cum.repr.trials.subg, label = "repr.groups", align = "lrrr", digits = 0), include.rownames = F, size = "footnotesize")
% @
% \end{frame}

% \begin{frame}{Dataset Structure}
% \begin{itemize}
% \item Comparison: What is compared, e.g. treatment vs. control
% \item Outcome: How it is compared
% \item Subgroup: Subgroup affiliation
% \item Meta-Analysis Group: Results from same comparison, outcome and subgroup
% \end{itemize}
% \end{frame}

% \begin{frame}[fragile]{Dataset Properties}
% Missing data:
% <<echo=FALSE, results = 'asis'>>=
% print(xtable(missing.table, align = "lr"), include.rownames = T, include.colnames = F)
% @
% \end{frame}

% \begin{frame}{Adjustment by regression}
% Similar to the tests, but with unnormalized effect $y_{i}$:
% 
% \begin{align}
% y_{i} = \beta_{0} + \beta_{1}x_{i}
% \end{align}
% 
% $\beta_{0}$	 corresponds to $y_{i}$ with $x_{i} = 0$
% \end{frame}
% 
% \begin{frame}[fragile]{Adjustment by regression}
% Linear regression method:
% 
% <<echo = FALSE, fig.height = 3>>=
% grid.arrange(reg.1, reg.2, ncol = 2)
% @
% \end{frame}

% \begin{frame}[fragile]{Thompson and Sharp's Tests}
% Generalised radial plot with new standard error $x_i = s_i + \tau$
% 
% \vspace{-1.1cm}
% <<message = FALSE, echo=FALSE, warning = FALSE, fig.height= 4>>=
% par(las = 1, mfrow = c(1,2))
% tau.ex <- meta.example$tau
% biased.rev$inv <- 1/sqrt((biased.rev$se^2 + tau.ex^2))
% m.ex <- lm(formula = (biased.rev$effect*biased.rev$inv) ~ (biased.rev$inv), weights = biased.rev$se)
% biased.rev$inv <- 1/biased.rev$se
% plot(y = (biased.rev$effect/biased.rev$se), x = (biased.rev$inv), xlim = c(0, 4), xlab = "inverse standard error", 
% 		 ylab = "mean diff. / std. error")
% abline(coef = coef(m.ex), lty = 2, col = 2)
% 
% tau.nex <- meta.nexample$tau
% unbiased.rev$inv <- 1/sqrt((unbiased.rev$se^2 + tau.nex^2))
% m.nex <- lm(formula = (unbiased.rev$effect*unbiased.rev$inv) ~ (unbiased.rev$inv), weights = unbiased.rev$se)
% plot(y = (unbiased.rev$effect/unbiased.rev$se), x = (unbiased.rev$inv), xlim = c(0, .6), 
% 		 ylim = c(-18, 0), xlab = "inverse standard error", ylab = "mean diff. / std. error")
% abline(coef = coef(m.nex), lty = 2, col = 2)
% @
% 
% \end{frame}
% 
% 
% \begin{frame}[fragile]{Limit Meta-Analysis}
% \begin{align}
% y_{M,i} &= \beta_{0} + \beta_{1}(\sqrt{v_{i}/M + \tau^2}) + \epsilon_{i}(\sqrt{s_{i}/M + \tau^2}) \nonumber
% \end{align}
% 
% Letting $M \rightarrow \infty$ and substituting for all parameters and the observed residual
% 
% \vspace{-8mm}
% \begin{align}
% y_{\infty,i} &= \beta_{0} + \sqrt{\frac{\tau^2}{v_{i} + \tau^2}}(y_i - \beta_0)
% \end{align}
% \end{frame}
% 
% \begin{frame}[fragile]{Limit Meta-Analysis}
% By plugging in $s_i$ for $v_i$, $\hat{tau}$ for $\tau$ and $\hat{\beta_0}$, get new study effects:
% 
% Three different treatment effect estimates:
% \begin{itemize}
% \item Expectated adjusted treatment effect with infinite precision: $\hat{\beta_{0}^\star} + \hat{\beta_{1}}\hat{\tau}$
% \item Fixed effect estimate based on $(y_{\infty,1},.. , y_{\infty,n}$
% \item Slope $\beta_{\textrm{lim}}$ of best-fitting regression line in radial plot with $(y_{\infty,i}, s_{i})$
% \end{itemize}
% \end{frame}

\end{document}